### 2022-07-14 cogsci-paper/cogsci-paper.Rmd
added files

---
title: "Does mere exposure to ideas encourage belief in them?"
bibliography: references.bib
csl: apa7.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Justin Mikell (jmikell@asu.edu)} \\ Arizona State University
    \AND {\large \bf Derek Powell (dmpowell@asu.edu)} \\ School of Social and Behavioral Sciences \\ Arizona State University}

abstract: >
    Numerous psychological findings have shown that mere exposure to ideas makes those ideas seem more true, a finding commonly referred to as the “illusory truth” effect [e.g. @hasher.etal1977]. In the presence of pervasive misinformation, this basic feature of cognition may undermine the functioning of a democratic society [@pennycook.etal2018]. However, genuine beliefs do not only produce judgments of truth, they also imply other beliefs and drive decision-making. Here, we sought to examine whether mere exposure to statements produces genuine beliefs by examining whether people draw inferences from statements after mere exposure. Surprisingly, and in contrast to familiarity-based accounts of the illusory truth effect [e.g. @dechene.etal2010], we found that exposure to “premise” statements affected participants’ truth ratings for novel “implied” statements. This “illusory implication” effect suggests that exposure to false statements has further-reaching impacts than previously thought and calls for a new mechanistic account of these effects.

keywords: >
    Illusory truth; Metacognition; Cognitive Psychology; Misinformation

output: 
  cogsci2016::cogsci_paper

final-submission: \cogscifinalcopy

# header-includes:
#   - \usepackage{float}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
```

```{r}
library(tidyverse)
library(patchwork)

df <- read_csv("../data/illusory-truth-main.csv") %>% 
  mutate(
    response = ordered(response, levels = c("Definitely False", "Probably False", "Maybe False", "Maybe True", "Probably True", "Definitely True"))
  ) %>% 
  mutate(resp_num = recode(response,
                           `Definitely True` = 2.5,
                           `Probably True` = 1.5,
                           `Maybe True` = .5,
                           `Maybe False` = -.5,
                           `Probably False` = -1.5,
                           `Definitely False` = -2.5)
         ) %>% 
  mutate(
    B_implied_true = ifelse(item_pair %in% c("war", "whitehouse", "bezos", "jackson",
                                          "indonesia", "eiffel", "newyork", "presidents",
                                          "tour", "africa", "disney", "avengers"), 
                            "F implied T", "T implied F"),
    B_implied_true = relevel(factor(B_implied_true), ref = "T implied F"),
      acc_num = case_when(
        item_type=="B" & B_implied_true == "F implied T" ~ -1*resp_num,
        item_type=="B" & B_implied_true != "F implied T" ~ resp_num,
        item_type=="A" ~ -1*resp_num,
        item_type=="Falsehood" ~ -1*resp_num,
        TRUE ~ resp_num
        ),
    acc_ordered = ordered(acc_num)
  )
```

```{r fit-models-exp1, include=F}
library(brms)

## Experiment 1

## Premises

fit_premise_fact <- brm(
  response ~ exposed + (1|subj_id) + (1|item_pair),
  data = df %>% 
    filter(phase=="test", item_type=="A", exposure_cond=="fact"),
  chains = 4,
  cores = 4,
  family=cumulative(),
  file="../local/fit_premise_fact"
)

fit_premise_quiz <- brm(
  response ~ exposed + (1|subj_id) + (1|item_pair),
  data = df %>% 
    filter(phase=="test", item_type=="A", exposure_cond=="quiz"),
  chains = 4,
  cores = 4,
  family=cumulative(),
  file="../local/fit_premise_quiz"
)

## within-subjs test

df_reg_within <- df %>% 
  filter(exposure_cond=="quiz") %>%
  filter(item_type == "A") %>% 
  ungroup() %>% 
  drop_na(response) %>% 
  select(subj_id, condition, exposed, exposure_cond, item_pair, phase, response)

## this is a bit needlessly complex but it is equivalent to testing the difference
fit_premise_within <- brm(
  response ~ phase + (1|subj_id:item_pair),
  data = df_reg_within,
  chains = 4,
  cores = 4,
  family=cumulative(),
  file="../local/fit_premise_within"
)

## Implications

fit_imp_fact <- brm(
  acc_ordered ~ exposed*B_implied_true + (1|subj_id) + (1|item_pair),
  data = df %>% 
    filter(phase=="test", item_type=="B", exposure_cond=="fact"),
  family = cumulative(),
  chains = 4,
  cores = 4,
  file="../local/fit_imp_fact"
)

# fit_imp_quiz <- brm(
#   acc_ordered ~ exposed + (1|subj_id) + (1|item_pair),
#   data = df %>% 
#     filter(phase=="test", item_type=="B", exposure_cond=="quiz"),
#   family = cumulative(),
#   chains = 4,
#   cores = 4,
#   file="../local/fit_imp_quiz"
# )

df_reg <- df %>% 
  mutate(
    familiarity = exposed,
    implication = case_when(
      B_implied_true == "F implied T" & exposed==1 ~ 1,
      B_implied_true == "T implied F" & exposed==1 ~ -1,
      TRUE ~ 0
    )
    # implied_true = if_else(B_implied_true == "F implied T" & exposure==1, 1, 0),
    # implied_false = if_else(B_implied_true != "F implied T" & exposure==1, 1, 0),
    ) %>% 
    filter(phase=="test", item_type=="B")

fit_imp_quiz_int <- brm(
  response ~ familiarity + implication + B_implied_true + 
    (1 + familiarity + implication|subj_id) + 
    (1 + familiarity + implication|item_pair),  # familiarity + implication + grouping (optional)
  data = df_reg %>% filter(exposure_cond=="quiz"),
  family = cumulative(),
  chains = 4,
  cores = 4,
  file="../local/fit_imp_quiz_int_max"
)

fit_imp_fact_int <- brm(
  response ~ familiarity + implication + B_implied_true + 
    (1 + familiarity + implication|subj_id) + 
    (1 + familiarity + implication|item_pair),  # familiarity + implication + grouping (optional)
  data = df_reg %>% filter(exposure_cond=="fact"),
  family = cumulative(),
  chains = 4,
  cores = 4,
  file="../local/fit_imp_fact_int_max"
)

```

```{r load-exp2, include=F}

df2 <- read_csv("../data/illusory-truth-main2.csv") %>% 
  mutate(
    response = ordered(response, levels = c("Definitely False", "Probably False", "Maybe False", "Maybe True", "Probably True", "Definitely True"))
  ) %>% 
  mutate(resp_num = recode(response,
                           `Definitely True` = 2.5,
                           `Probably True` = 1.5,
                           `Maybe True` = .5,
                           `Maybe False` = -.5,
                           `Probably False` = -1.5,
                           `Definitely False` = -2.5)
         ) %>% 
  mutate(
    B_implied_true = ifelse(item_pair %in% c("war", "whitehouse", "bezos", "jackson",
                                          "indonesia", "eiffel", "newyork", "presidents",
                                          "tour", "africa", "disney", "avengers"), 
                            "F implied T", "T implied F"),
    B_implied_true = relevel(factor(B_implied_true), ref = "T implied F"),
      acc_num = case_when(
        item_type=="B" & B_implied_true == "F implied T" ~ -1*resp_num,
        item_type=="B" & B_implied_true != "F implied T" ~ resp_num,
        item_type=="A" ~ -1*resp_num,
        item_type=="Falsehood" ~ -1*resp_num,
        TRUE ~ resp_num
        ),
    acc_ordered = ordered(acc_num)
  )
```

```{r fit-models-exp2, include=F}

## Experiment 2

df_reg2 <- df2 %>% 
  mutate(
    familiarity = exposed,
    implication = case_when(
      B_implied_true == "F implied T" & exposed==1 ~ 1,
      B_implied_true == "T implied F" & exposed==1 ~ -1,
      TRUE ~ 0
    )
    # implied_true = if_else(B_implied_true == "F implied T" & exposure==1, 1, 0),
    # implied_false = if_else(B_implied_true != "F implied T" & exposure==1, 1, 0),
    ) %>% 
    filter(phase=="test", item_type=="B", exposure_cond=="interest")

## Implications

fit_imp_exp2 <- brm(
  response ~ familiarity + implication + B_implied_true + 
    (1 + familiarity + implication|subj_id) + 
    (1 + familiarity + implication|item_pair),  # familiarity + implication + grouping (optional)
  data = df_reg2,
  family = cumulative(),
  prior = set_prior("lkj(5)", class = "cor"),
  chains = 4,
  cores = 4,
  iter = 3000,
  control = list(adapt_delta=.9),
  file="../local/fit_imp_exp2_max"
)

# Premises

fit_prem_exp2 <- brm(
  response ~ exposed + (1 + exposed|subj_id) + (1 + exposed|item_pair),
  data = df2 %>%
    filter(phase=="test", item_type=="A"),
  family = cumulative(),
  chains = 4,
  cores = 4,
  file="../local/fit_prem_exp2_max"
)
```

```{r make-tables, include=F}
make_table_df <- function(fit){
  fit_sum <- summary(fit)

  tbl_sum <- as_tibble(fit_sum$fixed, rownames="Term") %>%
    # bind_rows(
    #   as_tibble(fit_sum$random$item, rownames="Term") %>%
    #     mutate(Term = paste0("item_",Term))
    # ) %>%
    # bind_rows(
    #   as_tibble(fit_sum$random$subj_id, rownames="Term") %>%
    #     mutate(Term = paste0("subj_",Term))
    # ) %>%
    select(-Rhat, -Est.Error, -Bulk_ESS, -Tail_ESS)

  return(tbl_sum)
}

make_table <- function(table_df, caption=NULL){

  kableExtra::kbl(table_df, digits = 2, booktabs = T, caption=caption, escape = F, format="latex") %>%
    kableExtra::kable_classic(full_width=F)
}

mytable1 <- make_table_df(fit_imp_quiz_int) %>% 
  rename(`$CI_{2.5\\%}$` = `l-95% CI`,`$CI_{97.5\\%}$` = `u-95% CI`) %>%
  mutate(
    Term = case_when(
      Term == "B_implied_trueFimpliedT" ~ "implication type",
      # grepl("Intercept\\[", Term) ~ gsub("Intercept\\[(.*)\\]","$\\\\alpha_{\\1}$", Term),
      Term == "item_sd(Intercept)" ~ "$\\sigma_{item}$",
      Term == "subj_sd(Intercept)" ~ "$\\sigma_{subj}$",
      TRUE ~ Term
    )
  ) %>% 
  make_table("Population coefficients of Bayesian regression model for Implication effects in Experiment 1.")

mytable2 <- make_table_df(fit_imp_exp2) %>% 
  rename(`$CI_{2.5\\%}$` = `l-95% CI`,`$CI_{97.5\\%}$` = `u-95% CI`) %>%
  mutate(
    Term = case_when(
      Term == "B_implied_trueFimpliedT" ~ "implication type",
      # grepl("Intercept\\[", Term) ~ gsub("Intercept\\[(.*)\\]","$\\\\alpha_{\\1}$", Term),
      grepl("familiarity", Term) ~ "familiarity",
      # Term == "item_sd(Intercept)" ~ "$\\sigma_{item}$",
      # Term == "subj_sd(Intercept)" ~ "$\\sigma_{subj}$",
      TRUE ~ Term
    )
  ) %>% 
  make_table("Population coefficients of Bayesian regression model for Implication effects in Experiment 2.")

```

```{r helpers, include=F}
report_reg_coef <- function(model, coef_name){

  res_table <- summary(model)$fixed

  res <- res_table[coef_name,] %>% 
    mutate_if(is.numeric, ~round(.x, digits=3))

  res_text <- paste0(res[[1,1]], ", 95% CI [", res[1,3],", ",res[1,4], "]")
  return(res_text)
}

# report_reg_coef(fit_imp_exp2, "familiarityTRUE")
```

```{r demographics, include=F}
df_demo <- df %>% 
  group_by(subj_id) %>% 
  summarize(
    age = first(age),
    gender = first(gender)
  )

df2_demo <- df2 %>% 
  group_by(subj_id) %>% 
  summarize(
    age = first(age),
    gender = first(gender)
  )

exp1_n_f <- df_demo %>% 
  filter(gender=="Female") %>% 
  nrow()

exp2_n_f <- df2_demo %>% 
  filter(gender=="Female") %>% 
  nrow()

exp1_age <- median(df_demo$age, na.rm=TRUE)
exp2_age <- median(df2_demo$age, na.rm=TRUE)
```

Every day, people are faced with a barrage of unsupported claims, from pestering ad campaigns to blatant disinformation on social media. Altogether, we live within an information environment that is more connected and more saturated than ever before in human history [@bak-coleman.etal2021]. If we are sufficiently critical consumers of media, can we benefit from this rich access to information without being exploited by advertisers or misled by misinformation? Perhaps not. Numerous psychological findings indicate that mere exposure to ideas makes those ideas appear more true, a finding commonly referred to as the “illusory truth” effect [@dekeersmaecker.etal2020; @dechene.etal2010; @fazio.etal2015; @hasher.etal1977; @pennycook.etal2018]. This effect suggests that we cannot exist in an environment of mis- and disinformation without being affected by it---without exposure to these ideas distorting our sense of what is true.

Studies examining the illusory truth effect typically proceed in at least two phases. At exposure, participants are introduced to a set of false statements. Typically, the statements are part of a true/false quiz, or a cover story explains they are part of some other innocuous judgment task. Then, after some intervening time ranging from minutes to weeks, participants are asked to judge whether these statements are true. On average, participants rate the statements as more “true” when they have been exposed to them previously---an illusory truth effect.

Decades of research have demonstrated the consistency and robustness of the illusory truth effect [e.g. see @dechene.etal2010]. The effect has been demonstrated for frivolous trivia questions [e.g. @hasher.etal1977; @lacassagne.etal2021; @unkelbach2007; @wang.etal2016] as well as consequential fake news headlines [@pennycook.etal2018]. The illusory truth effect has also been shown to be robust across people with different levels of cognitive ability, need for cognitive closure, and cognitive styles [@dekeersmaecker.etal2020].

The effect is one of "illusory" truth because it occurs following "mere exposure" to statements. That is, it occurs when the statements are seen or heard in a non-communicative context, such as when they are read during a true/false quiz. Generally, being told something---even by a complete stranger of unknown trustworthiness---is prima facie reason for believing it [@grice1989]. But reading a statement on a true/false quiz is not reason for believing it---the statement is being shown only to test the quiz-taker's knowledge, and is just as likely to be false as to be true.

There is something unsettling about the illusory truth effect and the lack of agency it implies over our own beliefs. Even more worrisome, there could be dire societal implications if merely being exposed to an idea causes people to adopt it as a belief: Pennycook and colleagues [-@pennycook.etal2018] argue that the illusory truth effect, combined with an environment of pervasive misinformation, has important implications for the functioning of democratic society. The illusory truth effect suggests that mere exposure to misinformation could have impacts that cannot be stopped by fact-checking labels [@pennycook.etal2018], nor effectively curbed by retractions or corrective information [e.g. @ecker.etal2011; @lewandowsky.etal2012].

However, there is more to _belief_ than ratings of truth. To illustrate, philosophers advancing dispositionalist theories of beliefs have focused on a number of observable behaviors that are indicative of belief [@schwitzgebel2021]. Though a person’s assent to a proposition (whether they agree with or judge it to be truthful) is an important marker and convenient measure of belief, other features of belief are just as essential. For instance, beliefs imply other beliefs: for instance, believing that “it is sunny out” implies the belief that “it is daytime”. And believing misinformation, such as “Hillary Clinton runs a sex-trafficking ring out of a pizza parlor” [see @robb2017] might imply the belief that “Hillary Clinton would make a bad president.” Beliefs also inform action and decision-making, and whether or not someone holds a belief can be judged (at least partly) from their decisions. When truly believed, misinformation can have drastic impacts. To illustrate, a recent study found that people who marked just one piece of vaccine misinformation as accurate were nearly twice as likely to be unvaccinated against COVID-19 compared to those who did not endorse any vaccine misinformation [@ognyanova.etal2021].

Bearing these features of beliefs in mind, the current state of illusory truth research does not demonstrate that mere exposure to a statement truly engenders _belief_ in that statement. Consider the mechanisms thought to underlie the illusory truth effect: most explanations attribute illusory truth effects to participants’ reliance on familiarity or fluency during the judgment process [@unkelbach2007; @wang.etal2016; @fazio.etal2015]. Roughly, when asked whether something is true or false, people search their memory for knowledge pertaining to belief in the statement, but they also rely on a meta-cognitive sense of familiarity. Familiarity is a sign that we have “heard this somewhere before.” According to foundational theories of communication, a general assumption that communicators strive to make true utterances is a prerequisite for successful communication and social functioning [@grice1989]. Thus, if having “heard something before” suggests that someone said it, then ecologically this is a reasonable cue to truth. Psychological studies that elicit the illusory truth effect hijack this meta-cognitive heuristic: They provide this sense of familiarity, but from a communicative context that lacks any reasonable assumption of truth.

This predominant theoretical account suggests that illusory truth effects do not engender genuine belief and may not have the deleterious real-world effects that some authors have feared [c.f. @pennycook.etal2018]. Applying this mechanistic account to the real-world, we might summarize the most likely process as follows: When people are exposed to a false statement they form a memory impression of it, but they do not draw inferences from this statement to revise other beliefs, nor do they plan future actions on its basis. In a real-world context, there would be essentially no impact from the exposure. In the experimental context however, the statement is seen again and a truth-judgment is requested. It is only at this time that the memory impression of the statement produces a feeling of familiarity and affects the truth judgments rendered at that time. Under this theoretical account, the impact of prior exposure depends on a subsequent re-exposure, coupled with a judgment or action to be taken at that time. This would make the real-world consequences of the illusory truth effect far narrower than some have argued.

Nevertheless, much as we might be relieved to dismiss them, concerns like those raised by Pennycook and colleagues [-@pennycook.etal2018] loom large enough to warrant further consideration. Could mere exposure to statements truly impact beliefs beyond providing a sense of familiarity? Could illusory truth effects instead reflect genuine belief?

Here, we sought to examine whether the illusory truth effect reflects genuine beliefs in statements following their mere exposure. One way to test this is to examine whether people draw inferences from statements after exposure to those statements in a context without any presupposition of truth. We designed a study to examine whether mere exposure to one statement (the “premise”) could affect truth ratings for another, different statement that it would logically imply (the “implication”). Such an “illusory implication” effect would demonstrate that mere-exposure has genuine impacts on beliefs beyond those explained by familiarity. In contrast, if the illusory truth effect is caused entirely by feelings of familiarity during the judgment process, then there should be no effect of prior exposure for new unfamiliar statements.

# Experiment 1

Preregistration information for Experiment 1 can be found at https://osf.io/c4w8s/. Materials for both experiments can be found at https://osf.io/znq3y/.

## Methods

### Participants

A total of 400 Participants were recruited through Amazon Mechanical Turk. Participants who failed basic attention check questions were excluded from analyses. These questions asked participants to simply give a particular response. For example, "This item tests if you are paying attention, please select 'Definitely False'." At the end of the study, participants were also asked if they had searched online to check any of the claims during the experiment (after being reassured that they would receive compensation regardless of their answers). Participants who failed any attention checks or who indicated they searched online were excluded from analysis. This left a final sample of `r nrow(df_demo)` (`r exp1_n_f` female, median age `r exp1_age` years-old).

```{r fig1, fig.env="figure", fig.pos = "h", fig.align = "center", fig.height=2.5, fig.width=3, fig.cap="Average truth ratings for Implication statements in Experiment 1, broken down by exposure (exposed vs. unexposed), instructions (fact vs quiz conditions), implication type (true-implied-false vs false-implied-true). For visualization purposes, means were calcaulated by translating ordinal responses onto a scale from -2.5 to 2.5. Error bars indicate standard errors. "}
fig1 <- df %>% 
  filter(phase=="test") %>% 
  filter(item_type == "B") %>% 
  group_by(exposed, exposure_cond, B_implied_true) %>% 
  summarize(
    M = mean(resp_num),
    se = sd(resp_num)/sqrt(n()),
    ll = M - se,
    ul = M + se
  ) %>% 
  mutate(
    exposure_cond = if_else(exposure_cond=="fact", "Fact condition", "Quiz condition"),
    exposed = if_else(exposed==1, "Exposed", "Unexposed"),
    B_implied_true = as.character(B_implied_true)
    ) %>% 
  ggplot(
    aes(x=B_implied_true, color=factor(exposed), shape=factor(exposed), y = M, ymin = ll, ymax=ul)
  ) +
  geom_pointrange(position=position_dodge(width=.5)) +
  # coord_flip() +
  scale_color_manual(values = c("#da1e28", "#0f62fe")) +
  labs(y = "Truth rating", x = "Implication type", color="Exposure", shape="Exposure") +
  facet_wrap(~exposure_cond) +
  theme_bw(base_size=9.5) +
  theme(panel.grid=element_blank(), legend.position="bottom") +
  theme(axis.text.x = element_text(angle = 10, vjust = 0.5, hjust=.5))

fig1
```

### Materials

We created 24 pairs of statements presenting claims about history. Each pair consisted of a "premise" statement and a "implied" statement. Each “premise” statement was a falsehood that implied either the truth or falsity of the “implied” statement. Of the pairs, 12 had premise statements implying that a true statement was false (true-implied-false) and the other 12 had premise statements implying another false statement was true (false-implied-true). For instance, one "true-implied-false" pair was the premise statement, “No U.S. astronauts have died since the Challenger explosion in 1986” paired with the implied statement, “The space shuttle Columbia disintegrated over Texas in 2003” (true, but implied to be false by the "premise"). An example of a "false-implied-true" pair was the premise statement, "The Tour de France has been held every year since its inception in 1903" paired with the implied statement, "The Tour de France was still held during WWI and WWII" (false, but implied true by the "premise").

```{r, include=F}
# library(kableExtra)
# kableExtra::kable(
#   tibble(
#     Type = c("True implied false", "False implied true"),
#     Premise = c(
#       "The Tour de France has been held every year since its inception in 1903 (False)",
#       "Henry Ford Invented the first automobile with the Model T (False)"
#     ),
#     Implied = c(
#       "The Tour de France was still held during WWI and WWII (False, implied True)",
#       "Karl Benz patented and sold the first automobile in 1886 (True, implied false)"
#     )
#   ),
#   caption = "Examples of premise and implied statement pairs"
# ) %>% 
#   column_spec(1, width = "4em") %>% 
#   column_spec(2, width = "10em") %>% 
#   column_spec(3, width = "10em")

```

### Procedures

Participants first completed informed consent and a Captcha to limit bot participation.

The study proceeded in three phases: 1) an exposure phase, 2) a distraction phase, 3) and finally a testing phase. At the exposure phase, participants were presented with a subset of the "premise" statements (8 of the 24 total) as well as some control statements (12 true, 4 false). Then, after the distraction phase, the testing phase consisted of a true-or-false quiz. The items tested included both "premise" and "implied" statements from the 24 item-pairs.  For some of the tested "implied" statements, participants had previously seen the related "premise" statement (exposed implied test). For others, they had not seen the related statements (unexposed implied test). Similarly, they were tested on the "premise" statements they had previously seen during the exposure phase (exposed premise test) as well as new unseen premise statements (for which they had also not seen the related implied statements). Participants were randomly assigned into three counterbalancing conditions that varied which of the statement-pairs were assigned to each exposure/test combination.

At the beginning of the study, participants were randomly assigned to either the “fact” (_n_ = 100) or “quiz” (_n_ = 300) exposure condition and to one of three counterbalancing conditions. Participants assigned to the quiz and fact conditions received different instructions and performed different tasks in the exposure phase.

Participants in the fact condition were told the study’s main purpose was to learn more about how people learn and apply new knowledge. They were informed that the initial set of statements they would see were all facts, and were asked to rate how surprising each “fact” was to them. The true purpose of this condition was to test whether participants would successfully draw inferences from the "premise" to the "implied" statements when told that the premises were true.

Participants in the quiz condition were told that they were to be given a true/false quiz, so that some of the statements would be true and some false. These participants rated how confident they were that each of the presented statements were true or false. The purpose of the quiz condition was to provide "mere exposure" to these statements, to test for illusory truth and illusory implication effects. We tested for the illusory truth effect by comparing truth ratings for premise statements seen during exposure to those that had not been seen. Similarly, we tested for the illusory implication effect by comparing truth ratings for implication statements whose corresponding premise statements were and were not presented during the exposure phase. If mere exposure to "premise" statements affects participants' judgments of the "implied" statements, this would be evidence for an illusory implication effect.

After their initial exposure to the statements, participants provided basic demographic information and were then presented with the expanded 7 question version of the Cognitive Reflection Task [CRT, @frederick2005; @toplak.etal2014]. These tasks served to provide a period of distraction between exposure and test.

Then, participants advanced to the testing phase, where all participants were asked to rate their confidence that each statement was true or false. These ratings were made on a 6-point scale from "Definitely False" to "Definitely True". The test phase consisted of the 8 premise statements the participants previously saw, their respective matching 8 implication statements, 8 novel implication statements, 8 previously unseen premise statements, 12 true control statements, and 4 false control statements.

At the end of the study, participants were debriefed and presented with a list of the false statements they had seen.

## Results and discussion

Participants made their truth ratings in the test phase on a 6-point scale from “Definitely false” to “Definitely true”. To properly treat these Likert-style responses, the truth ratings were analyzed using multilevel Bayesian cumulative ordinal regression models [@burkner.vuorre2019] with random intercepts and slopes for participants and items. Cumulative ordinal regression models assume that participants’ discrete responses are driven by a continuous latent variable and a set of k-1 thresholds determining the range of the continuous variable corresponding to each of the ordinal response options. This model helps to account for potential differences in scale usage as well as the bounded nature of the response scale.

All models were fit using the brms R package [@burkner2017], with model posteriors estimated using the No-U-Turn Markov Chain Monte Carlo (MCMC) sampler implemented in Stan. Four MCMC chains were run for each model, with 2000 samples (1000 burn-in) drawn from each. Chains were assessed for convergence with $\hat{R}$ and the total estimated effective sample size was verified to be greater than 1000 for all parameters [@gelman.etal2014].

### Fact condition

First, we examined truth ratings for the premise and implications statements among participants who were told that the false statements were true at exposure (“fact” condition). As expected, exposure to the premise statements when presented as “facts” increased endorsement at test, $\beta =$ `r report_reg_coef(fit_premise_fact, "exposed")`. In addition, participants successfully drew inferences from these “facts”, resulting in decreased accuracy for the implication statements at test, $\beta =$ `r report_reg_coef(fit_imp_fact, "exposed")`.

### Quiz condition

It is quite appropriate that people should draw new inferences when they learn new facts. But would participants’ endorsements of the implication statements be affected by mere exposure to the premise statements in the “quiz” condition? To our surprise, it appears that they were. Figure 1 shows participants’ average truth ratings for the implication statements in the quiz condition. Participants gave higher truth ratings following exposure for false statements implied to be true, and somewhat lower truth ratings following exposure for true statements implied to be false.

```{r, showtable1, echo=F, message=F}
mytable1
```

```{r, showtable2, echo=F, message=F}
mytable2
```

An increase in truth ratings for the false-implied-true statements could potentially be explained by familiarity. This would be consistent with findings from Arkes and colleagues [-@arkes.etal1991], who found evidence for illusory truth for novel statements that were on the same topic as previously-seen statements. However, familiarity cannot explain the decrease in truth ratings for the true-implied-false items, as familiarity should encourage uniformly higher, not lower, truth ratings.

A regression model was used to tease apart the effects of familiarity and logical implication. The model includes a) a binary variable indicating the type of implication statement (true-implied-false or false-implied-true), b) a binary predictor for prior exposure to the related premise coded zero when the premise was not seen and 1 when it was seen (capturing potential effects of familiarity), c) and a variable capturing the effect of implications, coded 1 for implied truth, -1 for implied falsehood, and zero when the related premise statement was not seen at exposure. Thus the "implication" predictor accounts for the effect of exposure to the premise on truth judgments for the implied statements (either positive or negative), while the "familiarity" predictor accounts for any positive effect of familiarity. We also incorporated "maximal" random intercepts and slopes for all terms varying by subject and item [@barr.etal2013]. Expressed in the common “lme4” syntax [@bates.etal2015], the regression model:

\begin{align*}
\text{response} \sim 
 \text{item type} &+ \text{familiarity} + \text{implication} \\
&+ (1 + \text{familiarity} + \text{implication}|\text{subject}) \\
&+ (1 + \text{familiarity} + \text{implication}|\text{item})
\end{align*}

Table 1 presents a summary of the posterior distribution estimated for the population-level coefficients in this model. The results indicate that both familiarity and the implication of the premise statements affected participants’ truth ratings. The implication effect is very credibly greater than zero. To judge its overall magnitude, we compared the parameter estimate to the parameter estimate from the same regression model applied to participants’ responses in the “fact” condition ($\beta =$ `r report_reg_coef(fit_imp_fact_int, "implication")`). Although smaller than the effect observed in the “fact” condition, the effect in the “quiz” condition is of a similar order of magnitude.

Surprisingly, the effects of exposure on participants’ truth ratings for the premise statements were very subtle. Our primary preregistered test compared participants’ truth ratings for the exposed premise statements at test against their ratings for another set of 8 unexposed premise statements. This analysis did not find any evidence of an illusory truth effect, with a posterior estimate for the exposure parameter that credibly included zero. However, a secondary preregistered analysis comparing participants’ initial truth judgments for the premise-statements to their later truth judgments for those same statements at test did find a credible increase in endorsements, $\beta =$ `r report_reg_coef(fit_premise_within, "phasetest")`.

The small size of these illusory truth effects may owe to the use of a “quiz” judgment task at exposure as well as the relatively brief period of distraction between exposure and test used in Experiment 1. By first responding to the statements in a quiz, participants may then have felt pressure to maintain consistency in their responses, and may have still possessed some explicit memory of their responses for the premise statements. If so, this could have suppressed the influence of familiarity traditionally argued to produce the illusory truth effect.

More importantly, a drive to maintain consistency might also explain our surprising findings of an illusory implication effect. Participants showed a general bias toward judging statements “true” during the exposure phase, responding with some variation of “true” for 71.8% of responses. Attempting to maintain consistency with these prior responses could have thus influenced their responses to the implication statements, rendering the observed effect an artifact of the experimental context.

# Experiment 2

Experiment 2 was conducted to address the possibility that a pressure for internal consistency produced the effects on the implication-statement truth ratings observed in Experiment 1. Preregistration informationExperiment 2 can be found at https://osf.io/czkdh/.

```{r fig2, fig.env="figure", fig.pos = "t", fig.align = "center", fig.height=2.5, fig.width=2.5, fig.cap="Average truth ratings for Implication statements in Experiment 2, broken down by exposure (exposed vs. unexposed) and implication type (true-implied-false vs false-implied-true). For visualization purposes, means were calcaulated by translating ordinal responses onto a scale from -2.5 to 2.5. Error bars indicate standard errors."}

fig2 <- df2 %>% 
  filter(phase=="test") %>% 
  mutate(resp_num = recode(response,
                           `Definitely True` = 2.5,
                           `Probably True` = 1.5,
                           `Maybe True` = .5,
                           `Maybe False` = -.5,
                           `Probably False` = -1.5,
                           `Definitely False` = -2.5)
  ) %>% 
  mutate(
    B_implied_true = ifelse(item_pair %in% c("war", "whitehouse", "bezos", "jackson",
                                          "indonesia", "eiffel", "newyork", "presidents",
                                          "tour", "africa", "disney", "avengers"), "F implied T", "T implied F"),
      acc_num = ifelse(B_implied_true == "F implied T", -1*resp_num, resp_num)
  ) %>% 
  filter(item_type == "B") %>% 
  mutate(
    exposed = if_else(exposed, "Exposed", "Unexposed")
  ) %>% 
  group_by(exposed, exposure_cond, B_implied_true) %>% 
  summarize(
    M = mean(resp_num),
    se = sd(resp_num)/sqrt(n()),
    ll = M - se,
    ul = M + se
  ) %>% 
  ggplot(
    aes(x=B_implied_true, color=factor(exposed), shape=factor(exposed), y = M, ymin = ll, ymax=ul)
  ) +
  geom_pointrange(position=position_dodge(width=.5)) +
  scale_color_manual(values = c("#da1e28", "#0f62fe")) +
  theme_bw(base_size=10) +
  theme(legend.position="bottom", panel.grid=element_blank()) +
  # coord_flip() +
  labs(y = "Truth rating", x = "Implication type", color="Exposure", shape="Exposure")

fig2
```

## Methods

### Participants

A total of 300 participants were recruited from Amazon Mechanical Turk using procedures identical to Experiment 1. As before, participants who failed attention check questions or who indicated they had looked up answers were excluded from analyses, leaving a final sample of `r nrow(df2_demo)` (`r exp2_n_f` female, median age `r exp2_age` years-old).

### Materials and procedures

All items and procedures were identical to Experiment 1 except for three changes.

First, all participants were assigned to a new “interest” condition. At exposure, participants were told that they would see a set of statements, some true and some false, and they were instructed to rate how interesting each statement was. This change was made to avoid forcing participants to make a true/false judgment for the exposed premise-statements, which could have pushed them to try to make coherent or consistent responses to the related implication-statements. Importantly, as with a true/false quiz, the presentation of statements in this context should not warrant any inference as to the truth of those statements.

Secondly, to provide a longer period of distraction between exposure and test, the ordering of the items in the test phase was rearranged. Participants first rated 24 control items (12 true and 12 false) to extend the time between rating main test items.

Third, another change in test presentation order further prevented any consistency-pressure: Participants judged the truth of all of the implication statements before judging the premise statements.

## Results and discussion

Figure 2 shows participants’ average truth ratings for the implication statements with and without exposure to their corresponding premise statements in Experiment 2. The illusory implication effect was again observed. As shown in the figure, truth ratings were clearly affected by exposure: participants gave higher truth ratings following exposure for false statements implied to be true, and lower truth ratings following exposure for true statements implied to be false. Table 2 shows the posterior population-level estimates for an identical regression as was conducted for Experiment 1. Experiment 2 replicated the effects of Experiment 1 in a revised design that eliminated any consistency pressure or demands on participants. In fact, the magnitude of the effect for the implication statements was somewhat larger in Experiment 2 than in Experiment 1.

```{r fig3, fig.env="figure*", fig.pos = "!hbt", fig.align = "center", fig.width=6.5, fig.height=3, fig.cap="Average truth ratings for individual premise and implication statements in Experiment 2, broken down by exposure (exposed vs. unexposed) and implication type (true-implied-false vs false-implied-true). For visualization purposes, means were calcaulated by translating ordinal responses onto a scale from -2.5 to 2.5. Error bars indicate standard errors.", num.cols.cap=2}

df2_plotting <- df2 %>% 
  filter(phase=="test") %>% 
  mutate(resp_num = recode(response,
                           `Definitely True` = 2.5,
                           `Probably True` = 1.5,
                           `Maybe True` = .5,
                           `Maybe False` = -.5,
                           `Probably False` = -1.5,
                           `Definitely False` = -2.5)
  ) %>% 
  mutate(
    B_implied_true = ifelse(item_pair %in% c("war", "whitehouse", "bezos", "jackson",
                                          "indonesia", "eiffel", "newyork", "presidents",
                                          "tour", "africa", "disney", "avengers"), "F implied T", "T implied F")
  ) %>% 
  mutate(exposed = if_else(exposed, "Exposed", "Unexposed"))

df2_extra <- df2_plotting %>% 
  filter(phase=="test") %>% 
  filter(item_type == "A") %>% 
  group_by(exposed, exposure_cond, item_pair, B_implied_true) %>% 
  summarize(
    M = mean(resp_num),
    # se = sd(resp_num)/sqrt(n()),
    # ll = M - se,
    # ul = M + se
  ) %>% 
  spread(exposed, M) %>% 
  mutate(diff = abs(Exposed-Unexposed)) %>% 
  select(-Exposed, -Unexposed)

plt_A <- df2_plotting %>% 
  filter(phase=="test") %>% 
  filter(item_type == "A") %>% 
  group_by(exposed, exposure_cond, item_pair, B_implied_true) %>% 
  summarize(
    M = mean(resp_num),
    se = sd(resp_num)/sqrt(n()),
    ll = M - se,
    ul = M + se
  ) %>% 
  left_join(df2_extra) %>% 
  ggplot(
    aes(x=reorder(item_pair, diff), color=factor(exposed), shape=factor(exposed), y = M, ymin = ll, ymax=ul)
  ) +
  geom_pointrange(position=position_dodge(width=.5)) +
  scale_color_manual(values = c("#da1e28", "#0f62fe")) +
  coord_flip() +
  labs(title = "Premise-statements", y = "Truth rating", x= "Item", color="Exposure", shape="Exposure") +
  theme_bw(base_size=9) +
  theme(legend.position="right", legend.direction="vertical",panel.grid=element_blank()) +
  facet_wrap(~B_implied_true, scales="free", ncol=1)

# plt_A

plt_B <- df2_plotting %>%  
  filter(phase=="test") %>% 
  filter(item_type == "B") %>% 
  group_by(exposed, exposure_cond, item_pair, B_implied_true) %>% 
  summarize(
    M = mean(resp_num),
    se = sd(resp_num)/sqrt(n()),
    ll = M - se,
    ul = M + se
  ) %>% 
  left_join(df2_extra) %>% 
  ggplot(
    aes(x=reorder(item_pair, diff), color=factor(exposed), shape=factor(exposed),  y = M, ymin = ll, ymax=ul)
  ) +
  geom_pointrange(position=position_dodge(width=.5)) +
  scale_color_manual(values = c("#da1e28", "#0f62fe")) +
  coord_flip() +
  labs(title = "Implication-statements", y = "Truth rating", x= "Item", color="Exposure", shape="Exposure") +
  theme_bw(base_size=9) +
  theme(legend.position="right", legend.direction="vertical", panel.grid=element_blank()) +
  facet_wrap(~B_implied_true, scales="free", ncol=1)

fig3 <- (plt_A + plt_B + guide_area()) + plot_layout(widths=c(4,4,1.5), heights=5, guides="collect")

fig3 # fix legend
```

In addition, participant’s truth ratings for the premise statements in Experiment 2 revealed the classic illusory truth effect: premise statements were rated as more true when participants had been exposed to them previously $\beta =$ `r report_reg_coef(fit_prem_exp2, "exposedTRUE")`.

Interestingly, though somewhat smaller, the illusory implication effect for the implication statements was generally similar in magnitude to the illusory truth effect observed among the premise statements.

One limitation of these studies relative to prior work on the illusory truth effect is the relatively limited number of statements participants saw and were tested on. Generating diverse pairs of statements with a clear implication relation is relatively more challenging than simply generating false statements. Fortunately as shown in Figure 3, the illusory implication effect was fairly robust across the majority of the 24 different individual item pairs in Experiment 2.

# Discussion

We observed an “illusory implication” effect across two preregistered experiments: mere exposure to premises (e.g. that “The Tour de France has been held every year since its inception in 1903”) influenced participant’s truth ratings for examples of their logical implications (e.g. that “The Tour de France was still held during World War I and World War II”). Unlike the illusory truth effect, this novel effect cannot be easily explained by the effects of familiarity or fluency on the judgment process [cf.  @fazio.etal2015; @unkelbach2007; @wang.etal2016]: whereas familiarity would be expected to uniformly increase endorsements, participants’ truth ratings for implication statements were both increased and decreased according to the implications of the corresponding premises. Thus, it appears that mere exposure has a genuine impression on people’s beliefs beyond simply creating a sense of familiarity.

If not familiarity, then what cognitive mechanism could explain these findings? One potential explanation is that mere exposure to the premise statements essentially leads participants to at least partially adopt the premise statements as beliefs. Simply entertaining the idea may create such an impression despite awareness of its ambiguous truth-value. Or, consistent with a general overriding expectation of truth in testimony [e.g. @grice1989; @levine2014], participants may at least partially accept the premise statements as true despite the experimental context making clear they should not. This explanation would be consistent with the general truth-bias observed throughout both studies. Future research might vary the plausibility of the premise statements to evaluate this explanation or identify its boundary conditions [@fazio.etal2019].

Another potential explanation would attribute the illusory implication effect to processes that play out in the testing phase. Rather than familiarity, the illusory implication effect may instead reflect explicit memory and source misattribution: Participants may (at least partially) remember the premise statements, but fail to properly attribute their source to the experimental context. Future research might explore how explicit memory for the premise statements correlates with the strength of the illusory implication effects. In particular, research might disentangle memory effects by exploring factors that are known to affect memory, but that should not affect other reasoning processes, such as retention intervals or serial-position effects [e.g. @bjork.whitten1974].

Finally, given the surprising nature of our findings, we must raise the possibility of potential deflationary explanations: that the finding could be some kind of experimental demand characteristic. For instance, participants’ might imagine there is some kind of trick so that they are expected to reason forward from the initial exposure to the items. Or, some proportion might have been sufficiently confused so as to imagine they were meant to believe the statements on their initial exposure (though it is hard to imagine this in Experiment 1). Further work should explore participants' perceptions of the purposes of these studies and the potential for different response strategies. However, it should be noted that similar concerns could be levied at essentially all prior research on the illusory truth effect.

Whatever the cognitive mechanism behind the illusory implication effect we observed, these findings are further cause for concern about the spread of misinformation on and offline. The spread of misinformation and fake news may outpace the spread of factual news [@vosoughi.etal2018] and creates massive challenges for online platforms seeking to remove, label, and correct misinformation [@sharma.etal2019]. These challenges may be further compounded by basic features of human cognition: mere exposure to false statements can make them appear true. Further, as our findings are the first to indicate, these impacts generalize to related statements and are potentially indicative of genuine belief following mere exposure.

# References

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent

### 2023-08-04 cogsci-paper/cogsci-paper.tex
for submission

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{apacite}

% KM added 1/4/18 to allow control of blind submission
\cogscifinalcopy

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{Does mere exposure to ideas encourage belief in them?}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\author{{\large \bf Justin Mikell (jmikell@asu.edu)} \\ Arizona State University \AND {\large \bf Derek Powell (dmpowell@asu.edu)} \\ School of Social and Behavioral Sciences \\ Arizona State University}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{CSLReferences}%
  {}%
  {\par}

\begin{document}

\maketitle

\begin{abstract}
Numerous psychological findings have shown that mere exposure to ideas
makes those ideas seem more true, a finding commonly referred to as the
``illusory truth'' effect (e.g. Hasher et al., 1977). In the presence of
pervasive misinformation, this basic feature of cognition may undermine
the functioning of a democratic society (Pennycook et al., 2018).
However, genuine beliefs do not only produce judgments of truth, they
also imply other beliefs and drive decision-making. Here, we sought to
examine whether mere exposure to statements produces genuine beliefs by
examining whether people draw inferences from statements after mere
exposure. Surprisingly, and in contrast to familiarity-based accounts of
the illusory truth effect (e.g. Dechêne et al., 2010), we found that
exposure to ``premise'' statements affected participants' truth ratings
for novel ``implied'' statements. This ``illusory implication'' effect
suggests that exposure to false statements has further-reaching impacts
than previously thought and calls for a new mechanistic account of these
effects.

\textbf{Keywords:}
Illusory truth; Metacognition; Cognitive Psychology; Misinformation
\end{abstract}

Every day, people are faced with a barrage of unsupported claims, from
pestering ad campaigns to blatant disinformation on social media.
Altogether, we live within an information environment that is more
connected and more saturated than ever before in human history
(Bak-Coleman et al., 2021). If we are sufficiently critical consumers of
media, can we benefit from this rich access to information without being
exploited by advertisers or misled by misinformation? Perhaps not.
Numerous psychological findings indicate that mere exposure to ideas
makes those ideas appear more true, a finding commonly referred to as
the ``illusory truth'' effect (De keersmaecker et al., 2020; Dechêne et
al., 2010; Fazio et al., 2015; Hasher et al., 1977; Pennycook et al.,
2018). This effect suggests that we cannot exist in an environment of
mis- and disinformation without being affected by it---without exposure
to these ideas distorting our sense of what is true.

Studies examining the illusory truth effect typically proceed in at
least two phases. At exposure, participants are introduced to a set of
false statements. Typically, the statements are part of a true/false
quiz, or a cover story explains they are part of some other innocuous
judgment task. Then, after some intervening time ranging from minutes to
weeks, participants are asked to judge whether these statements are
true. On average, participants rate the statements as more ``true'' when
they have been exposed to them previously---an illusory truth effect.

Decades of research have demonstrated the consistency and robustness of
the illusory truth effect (e.g.~see Dechêne et al., 2010). The effect
has been demonstrated for frivolous trivia questions (e.g. Hasher et
al., 1977; Lacassagne et al., 2021; Unkelbach, 2007; Wang et al., 2016)
as well as consequential fake news headlines (Pennycook et al., 2018).
The illusory truth effect has also been shown to be robust across people
with different levels of cognitive ability, need for cognitive closure,
and cognitive styles (De keersmaecker et al., 2020).

The effect is one of ``illusory'' truth because it occurs following
``mere exposure'' to statements. That is, it occurs when the statements
are seen or heard in a non-communicative context, such as when they are
read during a true/false quiz. Generally, being told something---even by
a complete stranger of unknown trustworthiness---is prima facie reason
for believing it (Grice, 1989). But reading a statement on a true/false
quiz is not reason for believing it---the statement is being shown only
to test the quiz-taker's knowledge, and is just as likely to be false as
to be true.

There is something unsettling about the illusory truth effect and the
lack of agency it implies over our own beliefs. Even more worrisome,
there could be dire societal implications if merely being exposed to an
idea causes people to adopt it as a belief: Pennycook and colleagues
(2018) argue that the illusory truth effect, combined with an
environment of pervasive misinformation, has important implications for
the functioning of democratic society. The illusory truth effect
suggests that mere exposure to misinformation could have impacts that
cannot be stopped by fact-checking labels (Pennycook et al., 2018), nor
effectively curbed by retractions or corrective information (e.g. Ecker
et al., 2011; Lewandowsky et al., 2012).

However, there is more to \emph{belief} than ratings of truth. To
illustrate, philosophers advancing dispositionalist theories of beliefs
have focused on a number of observable behaviors that are indicative of
belief (Schwitzgebel, 2021). Though a person's assent to a proposition
(whether they agree with or judge it to be truthful) is an important
marker and convenient measure of belief, other features of belief are
just as essential. For instance, beliefs imply other beliefs: for
instance, believing that ``it is sunny out'' implies the belief that
``it is daytime.'' And believing misinformation, such as ``Hillary
Clinton runs a sex-trafficking ring out of a pizza parlor'' (see Robb,
2017) might imply the belief that ``Hillary Clinton would make a bad
president.'' Beliefs also inform action and decision-making, and whether
or not someone holds a belief can be judged (at least partly) from their
decisions. When truly believed, misinformation can have drastic impacts.
To illustrate, a recent study found that people who marked just one
piece of vaccine misinformation as accurate were nearly twice as likely
to be unvaccinated against COVID-19 compared to those who did not
endorse any vaccine misinformation (Ognyanova et al., 2021).

Bearing these features of beliefs in mind, the current state of illusory
truth research does not demonstrate that mere exposure to a statement
truly engenders \emph{belief} in that statement. Consider the mechanisms
thought to underlie the illusory truth effect: most explanations
attribute illusory truth effects to participants' reliance on
familiarity or fluency during the judgment process (Fazio et al., 2015;
Unkelbach, 2007; Wang et al., 2016). Roughly, when asked whether
something is true or false, people search their memory for knowledge
pertaining to belief in the statement, but they also rely on a
meta-cognitive sense of familiarity. Familiarity is a sign that we have
``heard this somewhere before.'' According to foundational theories of
communication, a general assumption that communicators strive to make
true utterances is a prerequisite for successful communication and
social functioning (Grice, 1989). Thus, if having ``heard something
before'' suggests that someone said it, then ecologically this is a
reasonable cue to truth. Psychological studies that elicit the illusory
truth effect hijack this meta-cognitive heuristic: They provide this
sense of familiarity, but from a communicative context that lacks any
reasonable assumption of truth.

This predominant theoretical account suggests that illusory truth
effects do not engender genuine belief and may not have the deleterious
real-world effects that some authors have feared (c.f. Pennycook et al.,
2018). Applying this mechanistic account to the real-world, we might
summarize the most likely process as follows: When people are exposed to
a false statement they form a memory impression of it, but they do not
draw inferences from this statement to revise other beliefs, nor do they
plan future actions on its basis. In a real-world context, there would
be essentially no impact from the exposure. In the experimental context
however, the statement is seen again and a truth-judgment is requested.
It is only at this time that the memory impression of the statement
produces a feeling of familiarity and affects the truth judgments
rendered at that time. Under this theoretical account, the impact of
prior exposure depends on a subsequent re-exposure, coupled with a
judgment or action to be taken at that time. This would make the
real-world consequences of the illusory truth effect far narrower than
some have argued.

Nevertheless, much as we might be relieved to dismiss them, concerns
like those raised by Pennycook and colleagues (2018) loom large enough
to warrant further consideration. Could mere exposure to statements
truly impact beliefs beyond providing a sense of familiarity? Could
illusory truth effects instead reflect genuine belief?

Here, we sought to examine whether the illusory truth effect reflects
genuine beliefs in statements following their mere exposure. One way to
test this is to examine whether people draw inferences from statements
after exposure to those statements in a context without any
presupposition of truth. We designed a study to examine whether mere
exposure to one statement (the ``premise'') could affect truth ratings
for another, different statement that it would logically imply (the
``implication''). Such an ``illusory implication'' effect would
demonstrate that mere-exposure has genuine impacts on beliefs beyond
those explained by familiarity. In contrast, if the illusory truth
effect is caused entirely by feelings of familiarity during the judgment
process, then there should be no effect of prior exposure for new
unfamiliar statements.

\hypertarget{experiment-1}{%
\section{Experiment 1}\label{experiment-1}}

Preregistration information for Experiment 1 can be found at
\url{https://osf.io/c4w8s/}. Materials for both experiments can be found
at \url{https://osf.io/znq3y/}.

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

\hypertarget{participants}{%
\subsubsection{Participants}\label{participants}}

A total of 400 Participants were recruited through Amazon Mechanical
Turk. Participants who failed basic attention check questions were
excluded from analyses. These questions asked participants to simply
give a particular response. For example, ``This item tests if you are
paying attention, please select `Definitely False.'\,'' At the end of
the study, participants were also asked if they had searched online to
check any of the claims during the experiment (after being reassured
that they would receive compensation regardless of their answers).
Participants who failed any attention checks or who indicated they
searched online were excluded from analysis. This left a final sample of
378 (157 female, median age 36 years-old).

\begin{CodeChunk}
\begin{figure}[h]

{\centering \includegraphics{figs/fig1-1}

}

\caption[Average truth ratings for Implication statements in Experiment 1, broken down by exposure (exposed vs]{Average truth ratings for Implication statements in Experiment 1, broken down by exposure (exposed vs. unexposed), instructions (fact vs quiz conditions), implication type (true-implied-false vs false-implied-true). For visualization purposes, means were calcaulated by translating ordinal responses onto a scale from -2.5 to 2.5. Error bars indicate standard errors. }\label{fig:fig1}
\end{figure}
\end{CodeChunk}

\hypertarget{materials}{%
\subsubsection{Materials}\label{materials}}

We created 24 pairs of statements presenting claims about history. Each
pair consisted of a ``premise'' statement and a ``implied'' statement.
Each ``premise'' statement was a falsehood that implied either the truth
or falsity of the ``implied'' statement. Of the pairs, 12 had premise
statements implying that a true statement was false (true-implied-false)
and the other 12 had premise statements implying another false statement
was true (false-implied-true). For instance, one ``true-implied-false''
pair was the premise statement, ``No U.S. astronauts have died since the
Challenger explosion in 1986'' paired with the implied statement, ``The
space shuttle Columbia disintegrated over Texas in 2003'' (true, but
implied to be false by the ``premise''). An example of a
``false-implied-true'' pair was the premise statement, ``The Tour de
France has been held every year since its inception in 1903'' paired
with the implied statement, ``The Tour de France was still held during
WWI and WWII'' (false, but implied true by the ``premise'').

\hypertarget{procedures}{%
\subsubsection{Procedures}\label{procedures}}

Participants first completed informed consent and a Captcha to limit bot
participation.

The study proceeded in three phases: 1) an exposure phase, 2) a
distraction phase, 3) and finally a testing phase. At the exposure
phase, participants were presented with a subset of the ``premise''
statements (8 of the 24 total) as well as some control statements (12
true, 4 false). Then, after the distraction phase, the testing phase
consisted of a true-or-false quiz. The items tested included both
``premise'' and ``implied'' statements from the 24 item-pairs. For some
of the tested ``implied'' statements, participants had previously seen
the related ``premise'' statement (exposed implied test). For others,
they had not seen the related statements (unexposed implied test).
Similarly, they were tested on the ``premise'' statements they had
previously seen during the exposure phase (exposed premise test) as well
as new unseen premise statements (for which they had also not seen the
related implied statements). Participants were randomly assigned into
three counterbalancing conditions that varied which of the
statement-pairs were assigned to each exposure/test combination.

At the beginning of the study, participants were randomly assigned to
either the ``fact'' (\emph{n} = 100) or ``quiz'' (\emph{n} = 300)
exposure condition and to one of three counterbalancing conditions.
Participants assigned to the quiz and fact conditions received different
instructions and performed different tasks in the exposure phase.

Participants in the fact condition were told the study's main purpose
was to learn more about how people learn and apply new knowledge. They
were informed that the initial set of statements they would see were all
facts, and were asked to rate how surprising each ``fact'' was to them.
The true purpose of this condition was to test whether participants
would successfully draw inferences from the ``premise'' to the
``implied'' statements when told that the premises were true.

Participants in the quiz condition were told that they were to be given
a true/false quiz, so that some of the statements would be true and some
false. These participants rated how confident they were that each of the
presented statements were true or false. The purpose of the quiz
condition was to provide ``mere exposure'' to these statements, to test
for illusory truth and illusory implication effects. We tested for the
illusory truth effect by comparing truth ratings for premise statements
seen during exposure to those that had not been seen. Similarly, we
tested for the illusory implication effect by comparing truth ratings
for implication statements whose corresponding premise statements were
and were not presented during the exposure phase. If mere exposure to
``premise'' statements affects participants' judgments of the
``implied'' statements, this would be evidence for an illusory
implication effect.

After their initial exposure to the statements, participants provided
basic demographic information and were then presented with the expanded
7 question version of the Cognitive Reflection Task (CRT, Frederick,
2005; Toplak et al., 2014). These tasks served to provide a period of
distraction between exposure and test.

Then, participants advanced to the testing phase, where all participants
were asked to rate their confidence that each statement was true or
false. These ratings were made on a 6-point scale from ``Definitely
False'' to ``Definitely True.'' The test phase consisted of the 8
premise statements the participants previously saw, their respective
matching 8 implication statements, 8 novel implication statements, 8
previously unseen premise statements, 12 true control statements, and 4
false control statements.

At the end of the study, participants were debriefed and presented with
a list of the false statements they had seen.

\hypertarget{results-and-discussion}{%
\subsection{Results and discussion}\label{results-and-discussion}}

Participants made their truth ratings in the test phase on a 6-point
scale from ``Definitely false'' to ``Definitely true.'' To properly
treat these Likert-style responses, the truth ratings were analyzed
using multilevel Bayesian cumulative ordinal regression models (Bürkner
\& Vuorre, 2019) with random intercepts and slopes for participants and
items. Cumulative ordinal regression models assume that participants'
discrete responses are driven by a continuous latent variable and a set
of k-1 thresholds determining the range of the continuous variable
corresponding to each of the ordinal response options. This model helps
to account for potential differences in scale usage as well as the
bounded nature of the response scale.

All models were fit using the brms R package (Bürkner, 2017), with model
posteriors estimated using the No-U-Turn Markov Chain Monte Carlo (MCMC)
sampler implemented in Stan. Four MCMC chains were run for each model,
with 2000 samples (1000 burn-in) drawn from each. Chains were assessed
for convergence with \(\hat{R}\) and the total estimated effective
sample size was verified to be greater than 1000 for all parameters
(Gelman et al., 2014).

\hypertarget{fact-condition}{%
\subsubsection{Fact condition}\label{fact-condition}}

First, we examined truth ratings for the premise and implications
statements among participants who were told that the false statements
were true at exposure (``fact'' condition). As expected, exposure to the
premise statements when presented as ``facts'' increased endorsement at
test, \(\beta =\) 1.193, 95\% CI {[}0.987, 1.388{]}. In addition,
participants successfully drew inferences from these ``facts,''
resulting in decreased accuracy for the implication statements at test,
\(\beta =\) -0.345, 95\% CI {[}-0.601, -0.098{]}.

\hypertarget{quiz-condition}{%
\subsubsection{Quiz condition}\label{quiz-condition}}

It is quite appropriate that people should draw new inferences when they
learn new facts. But would participants' endorsements of the implication
statements be affected by mere exposure to the premise statements in the
``quiz'' condition? To our surprise, it appears that they were. Figure 1
shows participants' average truth ratings for the implication statements
in the quiz condition. Participants gave higher truth ratings following
exposure for false statements implied to be true, and somewhat lower
truth ratings following exposure for true statements implied to be
false.

\begin{CodeChunk}
\begin{table}

\caption{\label{tab:make-tables}Population coefficients of Bayesian regression model for Implication effects in Experiment 1.}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Term & Estimate & $CI_{2.5\%}$ & $CI_{97.5\%}$\\
\midrule
Intercept[1] & -3.38 & -3.86 & -2.92\\
Intercept[2] & -2.26 & -2.74 & -1.80\\
Intercept[3] & -1.36 & -1.83 & -0.91\\
Intercept[4] & 0.27 & -0.20 & 0.72\\
Intercept[5] & 1.80 & 1.34 & 2.27\\
\addlinespace
familiarity & 0.07 & -0.09 & 0.23\\
implication & 0.17 & 0.00 & 0.35\\
implication type & -0.84 & -1.47 & -0.19\\
\bottomrule
\end{tabular}
\end{table}

\end{CodeChunk}

\begin{CodeChunk}
\begin{table}

\caption{\label{tab:make-tables}Population coefficients of Bayesian regression model for Implication effects in Experiment 2.}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Term & Estimate & $CI_{2.5\%}$ & $CI_{97.5\%}$\\
\midrule
Intercept[1] & -3.17 & -3.58 & -2.73\\
Intercept[2] & -2.04 & -2.44 & -1.61\\
Intercept[3] & -1.18 & -1.58 & -0.75\\
Intercept[4] & 0.36 & -0.03 & 0.80\\
Intercept[5] & 1.93 & 1.52 & 2.36\\
\addlinespace
familiarity & 0.13 & -0.03 & 0.29\\
implication & 0.38 & 0.21 & 0.55\\
implication type & -0.79 & -1.36 & -0.21\\
\bottomrule
\end{tabular}
\end{table}

\end{CodeChunk}

An increase in truth ratings for the false-implied-true statements could
potentially be explained by familiarity. This would be consistent with
findings from Arkes and colleagues (1991), who found evidence for
illusory truth for novel statements that were on the same topic as
previously-seen statements. However, familiarity cannot explain the
decrease in truth ratings for the true-implied-false items, as
familiarity should encourage uniformly higher, not lower, truth ratings.

A regression model was used to tease apart the effects of familiarity
and logical implication. The model includes a) a binary variable
indicating the type of implication statement (true-implied-false or
false-implied-true), b) a binary predictor for prior exposure to the
related premise coded zero when the premise was not seen and 1 when it
was seen (capturing potential effects of familiarity), c) and a variable
capturing the effect of implications, coded 1 for implied truth, -1 for
implied falsehood, and zero when the related premise statement was not
seen at exposure. Thus the ``implication'' predictor accounts for the
effect of exposure to the premise on truth judgments for the implied
statements (either positive or negative), while the ``familiarity''
predictor accounts for any positive effect of familiarity. We also
incorporated ``maximal'' random intercepts and slopes for all terms
varying by subject and item (Barr et al., 2013). Expressed in the common
``lme4'' syntax (Bates et al., 2015), the regression model:

\begin{align*}
\text{response} \sim 
 \text{item type} &+ \text{familiarity} + \text{implication} \\
&+ (1 + \text{familiarity} + \text{implication}|\text{subject}) \\
&+ (1 + \text{familiarity} + \text{implication}|\text{item})
\end{align*}

Table 1 presents a summary of the posterior distribution estimated for
the population-level coefficients in this model. The results indicate
that both familiarity and the implication of the premise statements
affected participants' truth ratings. The implication effect is very
credibly greater than zero. To judge its overall magnitude, we compared
the parameter estimate to the parameter estimate from the same
regression model applied to participants' responses in the ``fact''
condition (\(\beta =\) 0.532, 95\% CI {[}0.215, 0.848{]}). Although
smaller than the effect observed in the ``fact'' condition, the effect
in the ``quiz'' condition is of a similar order of magnitude.

Surprisingly, the effects of exposure on participants' truth ratings for
the premise statements were very subtle. Our primary preregistered test
compared participants' truth ratings for the exposed premise statements
at test against their ratings for another set of 8 unexposed premise
statements. This analysis did not find any evidence of an illusory truth
effect, with a posterior estimate for the exposure parameter that
credibly included zero. However, a secondary preregistered analysis
comparing participants' initial truth judgments for the
premise-statements to their later truth judgments for those same
statements at test did find a credible increase in endorsements,
\(\beta =\) 0.213, 95\% CI {[}0.078, 0.345{]}.

The small size of these illusory truth effects may owe to the use of a
``quiz'' judgment task at exposure as well as the relatively brief
period of distraction between exposure and test used in Experiment 1. By
first responding to the statements in a quiz, participants may then have
felt pressure to maintain consistency in their responses, and may have
still possessed some explicit memory of their responses for the premise
statements. If so, this could have suppressed the influence of
familiarity traditionally argued to produce the illusory truth effect.

More importantly, a drive to maintain consistency might also explain our
surprising findings of an illusory implication effect. Participants
showed a general bias toward judging statements ``true'' during the
exposure phase, responding with some variation of ``true'' for 71.8\% of
responses. Attempting to maintain consistency with these prior responses
could have thus influenced their responses to the implication
statements, rendering the observed effect an artifact of the
experimental context.

\hypertarget{experiment-2}{%
\section{Experiment 2}\label{experiment-2}}

Experiment 2 was conducted to address the possibility that a pressure
for internal consistency produced the effects on the
implication-statement truth ratings observed in Experiment 1.
Preregistration informationExperiment 2 can be found at
\url{https://osf.io/czkdh/}.

\begin{CodeChunk}
\begin{figure}[t]

{\centering \includegraphics{figs/fig2-1}

}

\caption[Average truth ratings for Implication statements in Experiment 2, broken down by exposure (exposed vs]{Average truth ratings for Implication statements in Experiment 2, broken down by exposure (exposed vs. unexposed) and implication type (true-implied-false vs false-implied-true). For visualization purposes, means were calcaulated by translating ordinal responses onto a scale from -2.5 to 2.5. Error bars indicate standard errors.}\label{fig:fig2}
\end{figure}
\end{CodeChunk}

\hypertarget{methods-1}{%
\subsection{Methods}\label{methods-1}}

\hypertarget{participants-1}{%
\subsubsection{Participants}\label{participants-1}}

A total of 300 participants were recruited from Amazon Mechanical Turk
using procedures identical to Experiment 1. As before, participants who
failed attention check questions or who indicated they had looked up
answers were excluded from analyses, leaving a final sample of 286 (157
female, median age 38 years-old).

\hypertarget{materials-and-procedures}{%
\subsubsection{Materials and
procedures}\label{materials-and-procedures}}

All items and procedures were identical to Experiment 1 except for three
changes.

First, all participants were assigned to a new ``interest'' condition.
At exposure, participants were told that they would see a set of
statements, some true and some false, and they were instructed to rate
how interesting each statement was. This change was made to avoid
forcing participants to make a true/false judgment for the exposed
premise-statements, which could have pushed them to try to make coherent
or consistent responses to the related implication-statements.
Importantly, as with a true/false quiz, the presentation of statements
in this context should not warrant any inference as to the truth of
those statements.

Secondly, to provide a longer period of distraction between exposure and
test, the ordering of the items in the test phase was rearranged.
Participants first rated 24 control items (12 true and 12 false) to
extend the time between rating main test items.

Third, another change in test presentation order further prevented any
consistency-pressure: Participants judged the truth of all of the
implication statements before judging the premise statements.

\hypertarget{results-and-discussion-1}{%
\subsection{Results and discussion}\label{results-and-discussion-1}}

Figure 2 shows participants' average truth ratings for the implication
statements with and without exposure to their corresponding premise
statements in Experiment 2. The illusory implication effect was again
observed. As shown in the figure, truth ratings were clearly affected by
exposure: participants gave higher truth ratings following exposure for
false statements implied to be true, and lower truth ratings following
exposure for true statements implied to be false. Table 2 shows the
posterior population-level estimates for an identical regression as was
conducted for Experiment 1. Experiment 2 replicated the effects of
Experiment 1 in a revised design that eliminated any consistency
pressure or demands on participants. In fact, the magnitude of the
effect for the implication statements was somewhat larger in Experiment
2 than in Experiment 1.

\begin{CodeChunk}
\begin{figure*}[!hbt]

{\centering \includegraphics{figs/fig3-1}

}

\caption[Average truth ratings for individual premise and implication statements in Experiment 2, broken down by exposure (exposed vs]{Average truth ratings for individual premise and implication statements in Experiment 2, broken down by exposure (exposed vs. unexposed) and implication type (true-implied-false vs false-implied-true). For visualization purposes, means were calcaulated by translating ordinal responses onto a scale from -2.5 to 2.5. Error bars indicate standard errors.}\label{fig:fig3}
\end{figure*}
\end{CodeChunk}

In addition, participant's truth ratings for the premise statements in
Experiment 2 revealed the classic illusory truth effect: premise
statements were rated as more true when participants had been exposed to
them previously \(\beta =\) 0.525, 95\% CI {[}0.333, 0.719{]}.

Interestingly, though somewhat smaller, the illusory implication effect
for the implication statements was generally similar in magnitude to the
illusory truth effect observed among the premise statements.

One limitation of these studies relative to prior work on the illusory
truth effect is the relatively limited number of statements participants
saw and were tested on. Generating diverse pairs of statements with a
clear implication relation is relatively more challenging than simply
generating false statements. Fortunately as shown in Figure 3, the
illusory implication effect was fairly robust across the majority of the
24 different individual item pairs in Experiment 2.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

We observed an ``illusory implication'' effect across two preregistered
experiments: mere exposure to premises (e.g.~that ``The Tour de France
has been held every year since its inception in 1903'') influenced
participant's truth ratings for examples of their logical implications
(e.g.~that ``The Tour de France was still held during World War I and
World War II''). Unlike the illusory truth effect, this novel effect
cannot be easily explained by the effects of familiarity or fluency on
the judgment process (cf. Fazio et al., 2015; Unkelbach, 2007; Wang et
al., 2016): whereas familiarity would be expected to uniformly increase
endorsements, participants' truth ratings for implication statements
were both increased and decreased according to the implications of the
corresponding premises. Thus, it appears that mere exposure has a
genuine impression on people's beliefs beyond simply creating a sense of
familiarity.

If not familiarity, then what cognitive mechanism could explain these
findings? One potential explanation is that mere exposure to the premise
statements essentially leads participants to at least partially adopt
the premise statements as beliefs. Simply entertaining the idea may
create such an impression despite awareness of its ambiguous
truth-value. Or, consistent with a general overriding expectation of
truth in testimony (e.g. Grice, 1989; Levine, 2014), participants may at
least partially accept the premise statements as true despite the
experimental context making clear they should not. This explanation
would be consistent with the general truth-bias observed throughout both
studies. Future research might vary the plausibility of the premise
statements to evaluate this explanation or identify its boundary
conditions (Fazio et al., 2019).

Another potential explanation would attribute the illusory implication
effect to processes that play out in the testing phase. Rather than
familiarity, the illusory implication effect may instead reflect
explicit memory and source misattribution: Participants may (at least
partially) remember the premise statements, but fail to properly
attribute their source to the experimental context. Future research
might explore how explicit memory for the premise statements correlates
with the strength of the illusory implication effects. In particular,
research might disentangle memory effects by exploring factors that are
known to affect memory, but that should not affect other reasoning
processes, such as retention intervals or serial-position effects (e.g.
Bjork \& Whitten, 1974).

Finally, given the surprising nature of our findings, we must raise the
possibility of potential deflationary explanations: that the finding
could be some kind of experimental demand characteristic. For instance,
participants' might imagine there is some kind of trick so that they are
expected to reason forward from the initial exposure to the items. Or,
some proportion might have been sufficiently confused so as to imagine
they were meant to believe the statements on their initial exposure
(though it is hard to imagine this in Experiment 1). Further work should
explore participants' perceptions of the purposes of these studies and
the potential for different response strategies. However, it should be
noted that similar concerns could be levied at essentially all prior
research on the illusory truth effect.

Whatever the cognitive mechanism behind the illusory implication effect
we observed, these findings are further cause for concern about the
spread of misinformation on and offline. The spread of misinformation
and fake news may outpace the spread of factual news (Vosoughi et al.,
2018) and creates massive challenges for online platforms seeking to
remove, label, and correct misinformation (Sharma et al., 2019). These
challenges may be further compounded by basic features of human
cognition: mere exposure to false statements can make them appear true.
Further, as our findings are the first to indicate, these impacts
generalize to related statements and are potentially indicative of
genuine belief following mere exposure.

\hypertarget{references}{%
\section{References}\label{references}}

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}

\noindent

\hypertarget{refs}{}
\begin{CSLReferences}
\leavevmode\hypertarget{ref-arkes.etal1991}{}%
Arkes, H. R., Boehm, L. E., \& Xu, G. (1991). Determinants of judged
validity. \emph{Journal of Experimental Social Psychology},
\emph{27}(6), 576--605.
\url{https://doi.org/10.1016/0022-1031(91)90026-3}

\leavevmode\hypertarget{ref-bak-coleman.etal2021}{}%
Bak-Coleman, J. B., Alfano, M., Barfuss, W., Bergstrom, C. T., Centeno,
M. A., Couzin, I. D., Donges, J. F., Galesic, M., Gersick, A. S.,
Jacquet, J., Kao, A. B., Moran, R. E., Romanczuk, P., Rubenstein, D. I.,
Tombak, K. J., Bavel, J. J. V., \& Weber, E. U. (2021). Stewardship of
global collective behavior. \emph{Proceedings of the National Academy of
Sciences}, \emph{118}(27). \url{https://doi.org/10.1073/pnas.2025764118}

\leavevmode\hypertarget{ref-barr.etal2013}{}%
Barr, D. J., Levy, R., Scheepers, C., \& Tily, H. J. (2013). Random
effects structure for confirmatory hypothesis testing: {Keep} it
maximal. \emph{Journal of Memory and Language}, \emph{68}(3),
10.1016/j.jml.2012.11.001.
\url{https://doi.org/10.1016/j.jml.2012.11.001}

\leavevmode\hypertarget{ref-bates.etal2015}{}%
Bates, D., Mächler, M., Bolker, B., \& Walker, S. (2015). Fitting
{Linear} {Mixed}-{Effects} {Models} {Using} \textbf{lme4}. \emph{Journal
of Statistical Software}, \emph{67}(1).
\url{https://doi.org/10.18637/jss.v067.i01}

\leavevmode\hypertarget{ref-bjork.whitten1974}{}%
Bjork, R. A., \& Whitten, W. B. (1974). Recency-sensitive retrieval
processes in long-term free recall. \emph{Cognitive Psychology},
\emph{6}(2), 173--189.
\url{https://doi.org/10.1016/0010-0285(74)90009-7}

\leavevmode\hypertarget{ref-burkner2017}{}%
Bürkner, P.-C. (2017). \textbf{Brms} : {An} \emph{r} {Package} for
{Bayesian} {Multilevel} {Models} {Using} \emph{stan}. \emph{Journal of
Statistical Software}, \emph{80}(1).
\url{https://doi.org/10.18637/jss.v080.i01}

\leavevmode\hypertarget{ref-burkner.vuorre2019}{}%
Bürkner, P.-C., \& Vuorre, M. (2019). Ordinal {Regression} {Models} in
{Psychology}: {A} {Tutorial}. \emph{Advances in Methods and Practices in
Psychological Science}, \emph{2}(1), 77--101.
\url{https://doi.org/10.1177/2515245918823199}

\leavevmode\hypertarget{ref-dekeersmaecker.etal2020}{}%
De keersmaecker, J., Dunning, D., Pennycook, G., Rand, D. G., Sanchez,
C., Unkelbach, C., \& Roets, A. (2020). Investigating the {Robustness}
of the {Illusory} {Truth} {Effect} {Across} {Individual} {Differences}
in {Cognitive} {Ability}, {Need} for {Cognitive} {Closure}, and
{Cognitive} {Style}. \emph{Personality and Social Psychology Bulletin},
\emph{46}(2), 204--215. \url{https://doi.org/10.1177/0146167219853844}

\leavevmode\hypertarget{ref-dechene.etal2010}{}%
Dechêne, A., Stahl, C., Hansen, J., \& Wänke, M. (2010). The {Truth}
{About} the {Truth}: {A} {Meta}-{Analytic} {Review} of the {Truth}
{Effect}. \emph{Personality and Social Psychology Review}, \emph{14}(2),
238--257. \url{https://doi.org/10.1177/1088868309352251}

\leavevmode\hypertarget{ref-ecker.etal2011}{}%
Ecker, U. K. H., Lewandowsky, S., Swire, B., \& Chang, D. (2011).
Correcting false information in memory: {Manipulating} the strength of
misinformation encoding and its retraction. \emph{Psychonomic Bulletin
\& Review}, \emph{18}(3), 570--578.
\url{https://doi.org/10.3758/s13423-011-0065-1}

\leavevmode\hypertarget{ref-fazio.etal2015}{}%
Fazio, L. K., Brashier, N. M., Payne, B. K., \& Marsh, E. J. (2015).
Knowledge does not protect against illusory truth. \emph{Journal of
Experimental Psychology: General}, \emph{144}(5), 993--1002.
\url{https://doi.org/10.1037/xge0000098}

\leavevmode\hypertarget{ref-fazio.etal2019}{}%
Fazio, L. K., Rand, D. G., \& Pennycook, G. (2019). Repetition increases
perceived truth equally for plausible and implausible statements.
\emph{Psychonomic Bulletin \& Review}, \emph{26}(5), 1705--1710.
\url{https://doi.org/10.3758/s13423-019-01651-4}

\leavevmode\hypertarget{ref-frederick2005}{}%
Frederick, S. (2005). Cognitive {Reflection} and {Decision} {Making}.
\emph{Journal of Economic Perspectives}, \emph{19}(4), 25--42.
\url{https://doi.org/10.1257/089533005775196732}

\leavevmode\hypertarget{ref-gelman.etal2014}{}%
Gelman, A., Carlin, J. B., Stern, H. S., Dunson, D. B., Vehtari, A., \&
Rubin, D. B. (2014). \emph{Bayesian data analysis} (Third edition). CRC
Press.

\leavevmode\hypertarget{ref-grice1989}{}%
Grice, H. P. (1989). \emph{Studies in the way of words}. Harvard
university press.

\leavevmode\hypertarget{ref-hasher.etal1977}{}%
Hasher, L., Goldstein, D., \& Toppino, T. (1977). Frequency and the
conference of referential validity. \emph{Journal of Verbal Learning and
Verbal Behavior}, \emph{16}(1), 107--112.
\url{https://doi.org/10.1016/S0022-5371(77)80012-1}

\leavevmode\hypertarget{ref-lacassagne.etal2021}{}%
Lacassagne, D., Béna, J., \& Corneille, O. (2021). \emph{Is {Earth} a
{Perfect} {Square}? {Repetition} {Increases} the {Perceived} {Truth} of
{Highly} {Implausible} {Statements}}. PsyArXiv.
\url{https://doi.org/10.31234/osf.io/fce8z}

\leavevmode\hypertarget{ref-levine2014}{}%
Levine, T. R. (2014). Truth-{Default} {Theory} ({TDT}): {A} {Theory} of
{Human} {Deception} and {Deception} {Detection}. \emph{Journal of
Language and Social Psychology}, \emph{33}(4), 378--392.
\url{https://doi.org/10.1177/0261927X14535916}

\leavevmode\hypertarget{ref-lewandowsky.etal2012}{}%
Lewandowsky, S., Ecker, U. K. H., Seifert, C. M., Schwarz, N., \& Cook,
J. (2012). Misinformation and {Its} {Correction}: {Continued}
{Influence} and {Successful} {Debiasing}. \emph{Psychological Science in
the Public Interest}, \emph{13}(3), 106--131.
\url{https://doi.org/10.1177/1529100612451018}

\leavevmode\hypertarget{ref-ognyanova.etal2021}{}%
Ognyanova, K., Lazer, D., Baum, M., Druckman, J., Green, J., Perlis, R.
H., Santillana, M., Simonson, M. D., Lin, J., \& Uslu, A. (2021).
\emph{The {COVID} {States} {Project} \#60: {COVID}-19 vaccine
misinformation: {From} uncertainty to resistance}. OSF Preprints.
\url{https://doi.org/10.31219/osf.io/xtjad}

\leavevmode\hypertarget{ref-pennycook.etal2018}{}%
Pennycook, G., Cannon, T. D., \& Rand, D. G. (2018). Prior exposure
increases perceived accuracy of fake news. \emph{Journal of Experimental
Psychology: General}, \emph{147}(12), 1865--1880.
https://doi.org/\url{http://dx.doi.org/10.1037/xge0000465}

\leavevmode\hypertarget{ref-robb2017}{}%
Robb, A. (2017). Pizzagate: {Anatomy} of a {Fake} {News} {Scandal}. In
\emph{Rolling Stone}.
\url{https://www.rollingstone.com/feature/anatomy-of-a-fake-news-scandal-125877/}

\leavevmode\hypertarget{ref-schwitzgebel2021}{}%
Schwitzgebel, E. (2021). Belief. In E. N. Zalta (Ed.), \emph{The
{Stanford} {Encyclopedia} of {Philosophy}} (Winter 2021). Metaphysics
Research Lab, Stanford University.
\url{https://plato.stanford.edu/archives/win2021/entries/belief/}

\leavevmode\hypertarget{ref-sharma.etal2019}{}%
Sharma, K., Qian, F., Jiang, H., Ruchansky, N., Zhang, M., \& Liu, Y.
(2019). Combating {Fake} {News}: {A} {Survey} on {Identification} and
{Mitigation} {Techniques}. \emph{ACM Transactions on Intelligent Systems
and Technology}, \emph{10}(3), 1--42.
\url{https://doi.org/10.1145/3305260}

\leavevmode\hypertarget{ref-toplak.etal2014}{}%
Toplak, M. E., West, R. F., \& Stanovich, K. E. (2014). Assessing
miserly information processing: {An} expansion of the {Cognitive}
{Reflection} {Test}. \emph{Thinking \& Reasoning}, \emph{20}(2),
147--168. \url{https://doi.org/10.1080/13546783.2013.844729}

\leavevmode\hypertarget{ref-unkelbach2007}{}%
Unkelbach, C. (2007). Reversing the {Truth} {Effect}: {Learning} the
{Interpretation} of {Processing} {Fluency} in {Judgments} of {Truth}.
\emph{Journal of Experimental Psychology. Learning, Memory, and
Cognition}, \emph{33}, 219--230.
\url{https://doi.org/10.1037/0278-7393.33.1.219}

\leavevmode\hypertarget{ref-vosoughi.etal2018}{}%
Vosoughi, S., Roy, D., \& Aral, S. (2018). The spread of true and false
news online. \emph{Science}.
\url{https://doi.org/10.1126/science.aap9559}

\leavevmode\hypertarget{ref-wang.etal2016}{}%
Wang, W.-C., Brashier, N. M., Wing, E. A., Marsh, E. J., \& Cabeza, R.
(2016). On {Known} {Unknowns}: {Fluency} and the {Neural} {Mechanisms}
of {Illusory} {Truth}. \emph{Journal of Cognitive Neuroscience},
\emph{28}(5), 739--746. \url{https://doi.org/10.1162/jocn_a_00923}

\end{CSLReferences}

\bibliographystyle{apacite}

\end{document}

### 2023-08-04 paper/paper-papaja.Rmd
for submission

--- 2022-07-14 cogsci-paper/cogsci-paper.Rmd
@@
 ---
-title: "Does mere exposure to ideas encourage belief in them?"
-bibliography: references.bib
-csl: apa7.csl
-document-params: "10pt, letterpaper"
+title             : "The title"
+shorttitle        : "Title"
 
-author-information: > 
-    \author{{\large \bf Justin Mikell (jmikell@asu.edu)} \\ Arizona State University
-    \AND {\large \bf Derek Powell (dmpowell@asu.edu)} \\ School of Social and Behavioral Sciences \\ Arizona State University}
+author: 
+  - name          : "Justin Mikell"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    address       : "Arizona State University - West Campus, 4701 W. Thunderbird Rd., Glendale, AZ 85306"
+    email         : "jmikell@asu.edu"
+    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
+      - "Investigation"
+      - "Methodology"
+      - "Visualization"
+      - "Writing - Original Draft Preparation"
+      - "Writing - Review & Editing"
+  - name          : "Derek Powell"
+    affiliation   : "1"
+    role:
+      - "Conceptualization"
+      - "Formal Analysis"
+      - "Methodology"
+      - "Writing - Original Draft Preparation"
+      - "Writing - Review & Editing"
+      - "Validation"
+      - "Supervision"
 
-abstract: >
+affiliation:
+  - id            : "1"
+    institution   : "Arizona State University"
+
+authornote: |
+  Arizona State University, School of Social and Behavioral Sciences
+
+  Materials for all experiments can be found at https://osf.io/znq3y/.
+Preregistration information for Experiments can be found at https://osf.io/c4w8s/ (Exp.
+1), https://osf.io/czkdh/ (Exp.
+2), and https://osf.io/5ewa6 (Exp.
+3).
+
+abstract: |
   Numerous psychological findings have shown that mere exposure to ideas makes those ideas seem more true, a finding commonly referred to as the “illusory truth” effect [e.g.
@@
 In the presence of pervasive misinformation, this basic feature of cognition may undermine the functioning of a democratic society [@pennycook.etal2018].
-However, genuine beliefs do not only produce judgments of truth, they also imply other beliefs and drive decision-making.
+However, genuine beliefs do not only affect judgments about individual statements, they also imply other beliefs and drive decision-making.
 Here, we sought to examine whether mere exposure to statements produces genuine beliefs by examining whether people draw inferences from statements after mere exposure.
-Surprisingly, and in contrast to familiarity-based accounts of the illusory truth effect [e.g.
+Surprisingly, and in contrast to fluency or familiarity-based accounts of the illusory truth effect [e.g.
 @dechene.etal2010], we found that exposure to “premise” statements affected participants’ truth ratings for novel “implied” statements.
-This “illusory implication” effect suggests that exposure to false statements has further-reaching impacts than previously thought and calls for a new mechanistic account of these effects.
+This “illusory implication” effect suggests that the consequences of exposure to false statements reach further than previously thought and calls for a new mechanistic account of these effects.
 
-keywords: >
-    Illusory truth; Metacognition; Cognitive Psychology; Misinformation
+  <!-- https://tinyurl.com/ybremelq -->
 
-output: 
-  cogsci2016::cogsci_paper
+keywords          : "Illusory truth; Metacognition; Cognitive Psychology; Misinformation"
+wordcount         : "X"
 
-final-submission: \cogscifinalcopy
+bibliography      : "references.bib"
 
-# header-includes:
-#   - \usepackage{float}
+floatsintext      : yes
+linenumbers       : yes
+draft             : no
+mask              : no
+
+figurelist        : no
+tablelist         : no
+footnotelist      : no
+
+classoption       : "man"
+output            : papaja::apa6_pdf
 ---
+
+```{r setup, include = FALSE}
+library("papaja")
+r_refs("r-references.bib")
+```
+
+```{r analysis-preferences}
+# Seed for random number generation
+set.seed(42)
+# knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+```
 
@@
 library(patchwork)
+library(textstem)
 
@@
 
+```{r fit-models-exp3, include=F}
+df_all <- read_csv("../data/illusory-truth-main-2.csv")
+
+df_passing <- df_all %>% 
+  mutate(
+    fail_checks = case_when(
+      (item == "att_exp_titanic") & (response != "Statement B") ~ 1,
+      (item == "att_exp_old") & (response != "Statement A") ~ 1,
+      (item == "att_exp_wright") & (response != "Statement B") ~ 1,
+      (item == "att_test_taxes") & (grepl("True", response)) ~ 1,
+      (item == "att_test_pop") & (grepl("False", response)) ~ 1,
+      (item == "att_test_tall") & (grepl("True", response)) ~ 1,
+      instructions_check!="Mix of true and false" ~ 1,
+      TRUE ~ 0
+    )
+  ) %>%
+  group_by(subj_id) %>%
+  filter(sum(fail_checks) < 1) %>% # remove anyone failing checks (none fail)
+  filter(!grepl("att_", item)) %>% # (remove check items)
+  select(-fail_checks) %>% 
+  ungroup() %>% 
+  mutate(
+    suspicious = if_else(susp_check1 == "Yes", 1, 0)
+  )
+
+df3 <- df_passing  %>% 
+  filter(phase=="test") %>% 
+  mutate(
+    response = ordered(response, levels = c("Definitely False", "Probably False", "Maybe False", "Maybe True", "Probably True", "Definitely True"))
+  ) %>% 
+  mutate(resp_num = recode(response,
+                           `Definitely True` = 2.5,
+                           `Probably True` = 1.5,
+                           `Maybe True` = .5,
+                           `Maybe False` = -.5,
+                           `Probably False` = -1.5,
+                           `Definitely False` = -2.5)
+         ) %>% 
+  mutate(
+    B_implied_true = ifelse(item %in% c("war", "whitehouse", "gates", "jackson",
+                                          "indonesia", "eiffel", "newyork", "presidents",
+                                          "tour", "africa", "disney", "avengers"), 
+                            "F implied T", "T implied F"),
+    B_implied_true = relevel(factor(B_implied_true), ref = "T implied F"),
+      acc_num = case_when(
+        item_type=="B" & B_implied_true == "F implied T" ~ -1*resp_num,
+        item_type=="B" & B_implied_true != "F implied T" ~ resp_num,
+        item_type=="A" ~ -1*resp_num,
+        item_type=="Falsehood" ~ -1*resp_num,
+        TRUE ~ resp_num
+        ),
+    acc_ordered = ordered(acc_num)
+  )
+
+df_open_coded <- read_csv("../data/illusory-truth-main-2-free-responses-coded.csv") %>%
+  select(subj_id, flagged)
+
+df3 <- df3 %>% 
+  left_join(df_open_coded, by = "subj_id")
+
+## Main analyses
+
+df_reg3 <- df3 %>% 
+  mutate(
+    familiarity = exposed,
+    implication = case_when(
+      B_implied_true == "F implied T" & exposed==1 ~ 1,
+      B_implied_true == "T implied F" & exposed==1 ~ -1,
+      TRUE ~ 0
+    )
+    # implied_true = if_else(B_implied_true == "F implied T" & exposure==1, 1, 0),
+    # implied_false = if_else(B_implied_true != "F implied T" & exposure==1, 1, 0),
+    ) %>% 
+    filter(phase=="test", item_type=="B")
+
+fit_imp_maximal <- brm(
+  response ~ familiarity + implication + B_implied_true + (1 + familiarity + implication + B_implied_true |subj_id) + (1 + familiarity + implication|item),  # familiarity + implication + grouping (optional)
+  data = df_reg3,
+  family = cumulative(),
+  chains = 4,
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta = .95),
+  file="../local/fit_imp_rep_maximal"
+)
+
+# fit_imp <- brm(
+#   response ~ familiarity + implication + B_implied_true + (1 |subj_id) + (1 |item),  # familiarity + implication + grouping (optional)
+#   data = df_reg,
+#   family = cumulative(),
+#   chains = 4,
+#   cores = 4,
+#   file="local/fit_imp_rep"
+# )
+
+### excluding suspicious
+
+df_reg3_nosus <- df_reg3 %>% 
+    filter(suspicious==0)
+
+fit_imp_maximal2 <- brm(
+  response ~ familiarity + implication + B_implied_true + (1 + familiarity + implication + B_implied_true |subj_id) + (1 + familiarity + implication|item),  # familiarity + implication + grouping (optional)
+  data = df_reg3_nosus,
+  family = cumulative(),
+  chains = 4,
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta = .95),
+  file="../local/fit_imp_rep_maximal2"
+)
+
+df_reg3_noflag <- df_reg3 %>% 
+    filter(suspicious==0, flagged==0)
+
+fit_imp_maximal3 <- brm(
+  response ~ familiarity + implication + B_implied_true + (1 + familiarity + implication + B_implied_true |subj_id) + (1 + familiarity + implication|item),  # familiarity + implication + grouping (optional)
+  data = df_reg3_noflag,
+  family = cumulative(),
+  chains = 4,
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta = .95),
+  file="../local/fit_imp_rep_maximal3"
+)
+
+```
+
+```{r, item-fluency-analysis, include=F}
+statements_raw <- read_csv("../items.csv")
+
+clean_words <- function(x){
+  lower <- str_to_lower(x)
+  word_list <- str_split(lower, " ")
+  lemmas <- lemmatize_strings(word_list[[1]])
+  # no_stops <- lemmas[ !(lemmas %in% lemmatize_words(lexicon::sw_lucene)) ]
+  lemma_stems <- stem_words(lemmas)
+
+  out <- gsub("('|,)", "", lemma_stems)
+
+  return(out)
+
+}
+
+statements <- statements_raw %>%
+  mutate(
+    prem_lst = map(premise, clean_words),
+    imp_lst = map(implication, clean_words),
+    overlap = map2_dbl(prem_lst, imp_lst, ~sum(.y %in% .x)), # how many words in implication were in premise?
+    overlap_prop = overlap/length(imp_lst)
+  )
+
+statement_overlap <- statements %>%
+  group_by(B_implied_true) %>%
+  summarize(overlap = median(overlap_prop))
+
+item_fluency_fig <- statements %>%
+  mutate(B_implied_true = if_else(B_implied_true==1, "F implied T", "T implied F")) %>% 
+  ggplot(aes(x = factor(B_implied_true), y = overlap_prop)) +
+  stat_summary(fun.data=mean_se, fun.args = list(mult=2), 
+                 geom="pointrange", color="red", shape=3, position=position_dodge()) +
+  geom_jitter(width=.1, size=2) +
+  # geom_dotplot(binaxis="y", stackdir="center", ,
+  #              stackratio=1.5, dotsize=.6) +
+  theme_bw(base_size = 12) +
+  theme(panel.grid=element_blank()) +
+  ylim(0,.5) +
+  labs(y = "Prop. shared lemmas", x = "Item Group")
+```
+
 ```{r make-tables, include=F}
@@
 
+mytable3 <- make_table_df(fit_imp_maximal) %>% 
+  rename(`$CI_{2.5\\%}$` = `l-95% CI`,`$CI_{97.5\\%}$` = `u-95% CI`) %>%
+  mutate(
+    Term = case_when(
+      Term == "B_implied_trueFimpliedT" ~ "implication type",
+      # grepl("Intercept\\[", Term) ~ gsub("Intercept\\[(.*)\\]","$\\\\alpha_{\\1}$", Term),
+      grepl("familiarity", Term) ~ "familiarity",
+      # Term == "item_sd(Intercept)" ~ "$\\sigma_{item}$",
+      # Term == "subj_sd(Intercept)" ~ "$\\sigma_{subj}$",
+      TRUE ~ Term
+    )
+  ) %>% 
+  make_table("Population coefficients of Bayesian regression model for Implication effects in Experiment 3.")
+
 ```
@@
 
-  res_table <- summary(model)$fixed
+  res_df <- broom.mixed::tidy(model, effects="fixed")
+  res <- res_df %>% 
+    filter(term == coef_name) %>% 
+    mutate_if(is.numeric, ~round(., 3))
 
-  res <- res_table[coef_name,] %>% 
-    mutate_if(is.numeric, ~round(.x, digits=3))
-
-  res_text <- paste0(res[[1,1]], ", 95% CI [", res[1,3],", ",res[1,4], "]")
+  res_text <- paste0(res[[1,4]], ", 95% CI [", res[[1,6]],", ",res[[1,7]], "]")
   return(res_text)
@@
 # report_reg_coef(fit_imp_exp2, "familiarityTRUE")
+# report_reg_coef(fit_imp_fact, "exposed")
 ```
@@
 
+df3_demo <- df3 %>% 
+  group_by(subj_id) %>% 
+  summarize(
+    age = first(age),
+    gender = first(gender)
+  )
+
 exp1_n_f <- df_demo %>% 
@@
 
+exp3_n_f <- df3_demo %>% 
+  filter(gender=="Female") %>% 
+  nrow()
+
 exp1_age <- median(df_demo$age, na.rm=TRUE)
 exp2_age <- median(df2_demo$age, na.rm=TRUE)
+exp3_age <- median(df3_demo$age, na.rm=TRUE)
+
+exp3_susp_count <- df3 %>% 
+  filter(suspicious==1) %>% 
+  distinct(subj_id) %>% 
+  nrow()
+
+exp3_flagged_count <- df3 %>% 
+  filter(suspicious==0) %>%
+  distinct(subj_id, .keep_all = TRUE) %>% 
+  filter(flagged==1) %>% 
+  nrow()
+
 ```
+
+## Relevance Statement
+
+Three pre-registered experiments examined whether mere exposure to ideas might lead them to be adopted as beliefs.
+It is well-established that mere exposure, such as in a true-or-false quiz, can lead people to later rate statements as more true (the “illusory truth” effect).
+If people's beliefs are affected by mere exposure, then the pervasiveness of misinformation would be even more harmful than currently recognized.
+However, leading theories attribute illusory truth effects to feelings of familiarity or fluency rather than changes in beliefs.
+These metacognitive heuristics are unlikely to produce further inferences or drive decision-making, as would beliefs.
+In contrast, we demonstrate that mere exposure to statements affected participants' responses to other novel statements (what we term an “illusory implication” effect), suggesting that mere exposure can lead to genuine belief.
+Our findings illustrate new dangers in exposure to misinformation, showing that mere exposure may have more wide-ranging influences than previously understood.
+
+\newpage
 
@@
 That is, it occurs when the statements are seen or heard in a non-communicative context, such as when they are read during a true/false quiz.
-Generally, being told something---even by a complete stranger of unknown trustworthiness---is prima facie reason for believing it [@grice1989].
-But reading a statement on a true/false quiz is not reason for believing it---the statement is being shown only to test the quiz-taker's knowledge, and is just as likely to be false as to be true.
+Generally, being told something---even by a source of unknown trustworthiness---is prima facie reason for believing it [@grice1989].
+But reading a statement on a true/false quiz is not reason for believing it; the statement is just as likely to be false as to be true.
+
+Why should mere exposure affect later ratings of truth?
+Predominant theoretical accounts attribute illusory truth effects to metacognitive experiences of fluency or familiarity [@dechene.etal2010].
+Research on metacognition has shown that people's behavior and judgments are not only influenced by the content of their thoughts, but also by the phenomenological experience of processing those thoughts [For a review, see @schwarz2015].
+
+Metacognitive processing "fluency" is the experienced ease of processing or thinking about information [@alter.oppenheimer2009; @wanke.hansen2015].
+When thinking is easy and fluent, people tend to judge the targets of that thinking more favorably than if thinking is challenging and disfluent [@wanke.etal1997].
+A number of findings have shown that processing fluency contributes to the illusory truth effect: Mere prior exposure to statements makes later processing of those statements more fluent, and this metacognitive experience of fluency leads people to rate those statements as more true [@difonzo.etal2016; @fazio.etal2015; @unkelbach2007; @wang.etal2016].
+
+Other researchers have attributed the illusory truth effect to familiarity [e.g.
+@arkes.etal1991; @mitchell.etal2006].
+Prior exposure to statements engenders a feeling of familiarity upon re-exposure to those statements.
+This metacognitive feeling of familiarity provides a cue that leads people to infer the statement is more likely to be true.
+
+Both theoretical accounts provide a similar overall view of the illusory truth effect: Roughly, when asked whether something is true or false, people search their memory for knowledge pertaining to belief in the statement, but they also rely on their metacognitive experiences of familiarity and fluency.
+Familiarity and fluency are both signs that we have “heard this somewhere before.” According to foundational theories of communication, a general assumption that communicators strive to make true utterances is a prerequisite for successful communication and social functioning [@grice1989].
+Thus, if having “heard something before” suggests that someone said it, then ecologically this is a reasonable cue to truth.
+Psychological studies that elicit the illusory truth effect hijack these metacognitive heuristics: They provide this sense of familiarity or fluency, but from a communicative context that lacks any reasonable assumption of truth.
 
@@
 Even more worrisome, there could be dire societal implications if merely being exposed to an idea causes people to adopt it as a belief: Pennycook and colleagues [-@pennycook.etal2018] argue that the illusory truth effect, combined with an environment of pervasive misinformation, has important implications for the functioning of democratic society.
-The illusory truth effect suggests that mere exposure to misinformation could have impacts that cannot be stopped by fact-checking labels [@pennycook.etal2018], nor effectively curbed by retractions or corrective information [e.g.
+The illusory truth effect suggests that mere exposure to misinformation could have impacts that cannot be stopped by fact-checking labels [@pennycook.etal2018], nor effectively curbed by retractions or corrective information [c.f.
 @ecker.etal2011; @lewandowsky.etal2012].
 
-However, there is more to _belief_ than ratings of truth.
-To illustrate, philosophers advancing dispositionalist theories of beliefs have focused on a number of observable behaviors that are indicative of belief [@schwitzgebel2021].
+However, there are more to _beliefs_ than ratings of truth.
 Though a person’s assent to a proposition (whether they agree with or judge it to be truthful) is an important marker and convenient measure of belief, other features of belief are just as essential.
-For instance, beliefs imply other beliefs: for instance, believing that “it is sunny out” implies the belief that “it is daytime”.
-And believing misinformation, such as “Hillary Clinton runs a sex-trafficking ring out of a pizza parlor” [see @robb2017] might imply the belief that “Hillary Clinton would make a bad president.” Beliefs also inform action and decision-making, and whether or not someone holds a belief can be judged (at least partly) from their decisions.
-When truly believed, misinformation can have drastic impacts.
-To illustrate, a recent study found that people who marked just one piece of vaccine misinformation as accurate were nearly twice as likely to be unvaccinated against COVID-19 compared to those who did not endorse any vaccine misinformation [@ognyanova.etal2021].
+For one, beliefs imply other beliefs: for instance, believing that “it is sunny out” implies the belief that “it is daytime”.
+And believing misinformation, such as “Dominion voting machines were rigged in the 2020 election” might imply the belief that “Donald Trump actually won the 2020 election.” Beliefs also inform action and decision-making, and whether or not someone holds a belief can be judged (at least partly) from their decisions.
+When truly believed, misinformation can have serious consequences.
+For instance, a recent study found that people who marked just one piece of vaccine misinformation as accurate were nearly twice as likely to be unvaccinated against COVID-19 compared to those who did not endorse any vaccine misinformation [@ognyanova.etal2021].
 
-Bearing these features of beliefs in mind, the current state of illusory truth research does not demonstrate that mere exposure to a statement truly engenders _belief_ in that statement.
-Consider the mechanisms thought to underlie the illusory truth effect: most explanations attribute illusory truth effects to participants’ reliance on familiarity or fluency during the judgment process [@unkelbach2007; @wang.etal2016; @fazio.etal2015].
-Roughly, when asked whether something is true or false, people search their memory for knowledge pertaining to belief in the statement, but they also rely on a meta-cognitive sense of familiarity.
-Familiarity is a sign that we have “heard this somewhere before.” According to foundational theories of communication, a general assumption that communicators strive to make true utterances is a prerequisite for successful communication and social functioning [@grice1989].
-Thus, if having “heard something before” suggests that someone said it, then ecologically this is a reasonable cue to truth.
-Psychological studies that elicit the illusory truth effect hijack this meta-cognitive heuristic: They provide this sense of familiarity, but from a communicative context that lacks any reasonable assumption of truth.
+Generally, it is these wider consequences of belief (beyond truth ratings for a specific proposition) that should be most concerning.
+If illusory truth effects operate only through metacognitive influences, then these effects may not have much impact outside the lab.
+Under these theoretical accounts, the impact of prior exposure depends on a subsequent re-exposure to a now fluent or familiar statement.
+This would make the real-world consequences of the illusory truth effect far narrower than one might fear: These metacognitive mechanisms are liable to apply only in response to the exact statements previously seen, but are not as likely to generalize to other related ideas, or to drive actions or decision-making in novel contexts.
 
-This predominant theoretical account suggests that illusory truth effects do not engender genuine belief and may not have the deleterious real-world effects that some authors have feared [c.f.
-@pennycook.etal2018].
-Applying this mechanistic account to the real-world, we might summarize the most likely process as follows: When people are exposed to a false statement they form a memory impression of it, but they do not draw inferences from this statement to revise other beliefs, nor do they plan future actions on its basis.
-In a real-world context, there would be essentially no impact from the exposure.
-In the experimental context however, the statement is seen again and a truth-judgment is requested.
-It is only at this time that the memory impression of the statement produces a feeling of familiarity and affects the truth judgments rendered at that time.
-Under this theoretical account, the impact of prior exposure depends on a subsequent re-exposure, coupled with a judgment or action to be taken at that time.
-This would make the real-world consequences of the illusory truth effect far narrower than some have argued.
+Nevertheless, concerns about how the influence of mere exposure might operate in contexts of pervasive misinformation [@difonzo.etal2016; @pennycook.etal2018] loom large enough to warrant further consideration.
+Could mere exposure to statements truly change beliefs?
 
-Nevertheless, much as we might be relieved to dismiss them, concerns like those raised by Pennycook and colleagues [-@pennycook.etal2018] loom large enough to warrant further consideration.
-Could mere exposure to statements truly impact beliefs beyond providing a sense of familiarity?
-Could illusory truth effects instead reflect genuine belief?
+A crucial feature of belief is that beliefs are connected with other related beliefs.
+Thus, one way to test whether mere exposure affects beliefs is to examine whether mere exposure to statements can affect truth ratings for other related statements.
 
-Here, we sought to examine whether the illusory truth effect reflects genuine beliefs in statements following their mere exposure.
-One way to test this is to examine whether people draw inferences from statements after exposure to those statements in a context without any presupposition of truth.
-We designed a study to examine whether mere exposure to one statement (the “premise”) could affect truth ratings for another, different statement that it would logically imply (the “implication”).
-Such an “illusory implication” effect would demonstrate that mere-exposure has genuine impacts on beliefs beyond those explained by familiarity.
-In contrast, if the illusory truth effect is caused entirely by feelings of familiarity during the judgment process, then there should be no effect of prior exposure for new unfamiliar statements.
+A handful of studies are suggestive in this regard: Arkes and collagues [-@arkes.etal1991] found that mere exposure to one set of statements also affected judgments about novel related statements on the same topic (China).
+However, these findings could also be explained by a sense of familiary engendered by the shared topic of old and new statements [@arkes.etal1991].
+
+Bacon [-@bacon1979] found that mere exposure to one statement ($P$) could lead to lower truth ratings for a new, contradictory statements($\neg P$).
+For instance, mere exposure to "crocodiles sleep with their eyes open" produced lower truth ratings for "crocodiles sleep with their eyes closed".
+It seems clear at least that familiarity cannot explain this pattern of results.
+However, other findings under very similar paradigms have partly contradicted these results [@garcia-marques.etal2015] and it is somewhat unclear whether fluency could account for these findings, i.e.
+whether contradictory statements can be expected to lead to fluent or disfluent processing.
+
+Here, we sought to decisively examine whether mere exposure to statements could influence people's beliefs.
+To do so, we examined whether people draw inferences from statements after exposure in a context without any presupposition of truth.
+We designed a study to examine whether mere exposure to one statement (the “premise”) could affect truth ratings for a  different statement the premise would logically entail (the “implication”).
+Such an “illusory implication” effect would indicate that mere exposure has genuine impacts on beliefs extending beyond those attributable to metacognitive cues.
 
 # Experiment 1
-
-Preregistration information for Experiment 1 can be found at https://osf.io/c4w8s/. Materials for both experiments can be found at https://osf.io/znq3y/.
 
@@
 
-A total of 400 Participants were recruited through Amazon Mechanical Turk.
-Participants who failed basic attention check questions were excluded from analyses.
+A total of 400 Participants were recruited through CloudResearch.
+Basic attention check questions were used to screen inattentive participants.
 These questions asked participants to simply give a particular response.
-For example, "This item tests if you are paying attention, please select 'Definitely False'." At the end of the study, participants were also asked if they had searched online to check any of the claims during the experiment (after being reassured that they would receive compensation regardless of their answers).
-Participants who failed any attention checks or who indicated they searched online were excluded from analysis.
+Participants who failed any attention checks or who indicated during debrief that they had searched online for question answers were excluded from analysis.
 This left a final sample of `r nrow(df_demo)` (`r exp1_n_f` female, median age `r exp1_age` years-old).
 
-```{r fig1, fig.env="figure", fig.pos = "h", fig.align = "center", fig.height=2.5, fig.width=3, fig.cap="Average truth ratings for Implication statements in Experiment 1, broken down by exposure (exposed vs.
+```{r fig1, fig.env="figure", fig.pos = "h", fig.align = "center", fig.cap="Average truth ratings for Implication statements in Experiment 1, broken down by exposure (exposed vs.
 unexposed), instructions (fact vs quiz conditions), implication type (true-implied-false vs false-implied-true).
@@
 
-```{r, include=F}
-# library(kableExtra)
-# kableExtra::kable(
-#   tibble(
-#     Type = c("True implied false", "False implied true"),
-#     Premise = c(
-#       "The Tour de France has been held every year since its inception in 1903 (False)",
-#       "Henry Ford Invented the first automobile with the Model T (False)"
-#     ),
-#     Implied = c(
-#       "The Tour de France was still held during WWI and WWII (False, implied True)",
-#       "Karl Benz patented and sold the first automobile in 1886 (True, implied false)"
-#     )
-#   ),
-#   caption = "Examples of premise and implied statement pairs"
-# ) %>% 
-#   column_spec(1, width = "4em") %>% 
-#   column_spec(2, width = "10em") %>% 
-#   column_spec(3, width = "10em")
-
-```
-
 ### Procedures
 
-Participants first completed informed consent and a Captcha to limit bot participation.
-
-The study proceeded in three phases: 1) an exposure phase, 2) a distraction phase, 3) and finally a testing phase.
-At the exposure phase, participants were presented with a subset of the "premise" statements (8 of the 24 total) as well as some control statements (12 true, 4 false).
-Then, after the distraction phase, the testing phase consisted of a true-or-false quiz.
-The items tested included both "premise" and "implied" statements from the 24 item-pairs.
-For some of the tested "implied" statements, participants had previously seen the related "premise" statement (exposed implied test).
-For others, they had not seen the related statements (unexposed implied test).
-Similarly, they were tested on the "premise" statements they had previously seen during the exposure phase (exposed premise test) as well as new unseen premise statements (for which they had also not seen the related implied statements).
-Participants were randomly assigned into three counterbalancing conditions that varied which of the statement-pairs were assigned to each exposure/test combination.
+The study proceeded in three phases: 1) an exposure phase, 2) a distraction phase, and 3) a final testing phase.
 
 At the beginning of the study, participants were randomly assigned to either the “fact” (_n_ = 100) or “quiz” (_n_ = 300) exposure condition and to one of three counterbalancing conditions.
+At the exposure phase, participants were presented with a subset of the "premise" statements (8 of the 24 total, counterbalanaced across participants) as well a set of control statements (12 true, 4 false).
 Participants assigned to the quiz and fact conditions received different instructions and performed different tasks in the exposure phase.
@@
 The purpose of the quiz condition was to provide "mere exposure" to these statements, to test for illusory truth and illusory implication effects.
+
+After their initial exposure to the statements, participants provided basic demographic information and were then presented with the expanded 7 question version of the Cognitive Reflection Task [CRT, @frederick2005; @toplak.etal2014].
+These tasks served to provide a period of distraction between exposure and test.
+
+Finally, the testing phase for all participants consisted of a true-or-false quiz, where participants were asked to rate their confidence that each statement was true or false.
+These ratings were made on a 6-point scale from "Definitely False" to "Definitely True".
+The items tested included both "premise" and "implied" statements from the 24 item-pairs.
+Participants were quizzed on the 8 previously-exposed premise statements, as well as their corresponding "implied" statements.
+For comparison, they were also quizzed on 8 new premise statements and 8 unrelated implied statements.
+Participants were randomly assigned into three counterbalancing conditions that varied which of the statement-pairs were assigned to each exposure/test combination.
+
 We tested for the illusory truth effect by comparing truth ratings for premise statements seen during exposure to those that had not been seen.
@@
 
-After their initial exposure to the statements, participants provided basic demographic information and were then presented with the expanded 7 question version of the Cognitive Reflection Task [CRT, @frederick2005; @toplak.etal2014].
-These tasks served to provide a period of distraction between exposure and test.
-
-Then, participants advanced to the testing phase, where all participants were asked to rate their confidence that each statement was true or false.
-These ratings were made on a 6-point scale from "Definitely False" to "Definitely True".
-The test phase consisted of the 8 premise statements the participants previously saw, their respective matching 8 implication statements, 8 novel implication statements, 8 previously unseen premise statements, 12 true control statements, and 4 false control statements.
-
 At the end of the study, participants were debriefed and presented with a list of the false statements they had seen.
 
-## Results and discussion
+## Results
 
@@
 To properly treat these Likert-style responses, the truth ratings were analyzed using multilevel Bayesian cumulative ordinal regression models [@burkner.vuorre2019] with random intercepts and slopes for participants and items.
-Cumulative ordinal regression models assume that participants’ discrete responses are driven by a continuous latent variable and a set of k-1 thresholds determining the range of the continuous variable corresponding to each of the ordinal response options.
+Cumulative ordinal regression models assume that participants’ discrete responses are driven by a continuous latent variable and a set of $k-1$ thresholds determining the range of the continuous variable corresponding to each of the ordinal response options.
 This model helps to account for potential differences in scale usage as well as the bounded nature of the response scale.
 
-All models were fit using the brms R package [@burkner2017], with model posteriors estimated using the No-U-Turn Markov Chain Monte Carlo (MCMC) sampler implemented in Stan.
+All models were fit using the `brms` R package [@burkner2017], with model posteriors estimated using the No-U-Turn Markov Chain Monte Carlo (MCMC) sampler implemented in Stan.
 Four MCMC chains were run for each model, with 2000 samples (1000 burn-in) drawn from each.
@@
 
-It is quite appropriate that people should draw new inferences when they learn new facts.
-But would participants’ endorsements of the implication statements be affected by mere exposure to the premise statements in the “quiz” condition?
-To our surprise, it appears that they were.
 Figure 1 shows participants’ average truth ratings for the implication statements in the quiz condition.
@@
 
-An increase in truth ratings for the false-implied-true statements could potentially be explained by familiarity.
+An increase in truth ratings for the false-implied-true statements could potentially be explained by increased familiarity.
 This would be consistent with findings from Arkes and colleagues [-@arkes.etal1991], who found evidence for illusory truth for novel statements that were on the same topic as previously-seen statements.
@@
 We also incorporated "maximal" random intercepts and slopes for all terms varying by subject and item [@barr.etal2013].
-Expressed in the common “lme4” syntax [@bates.etal2015], the regression model:
+Expressed in the common “lme4” syntax [@bates.etal2015], the regression model was:
 
@@
 The results indicate that both familiarity and the implication of the premise statements affected participants’ truth ratings.
-The implication effect is very credibly greater than zero.
-To judge its overall magnitude, we compared the parameter estimate to the parameter estimate from the same regression model applied to participants’ responses in the “fact” condition ($\beta =$ `r report_reg_coef(fit_imp_fact_int, "implication")`).
-Although smaller than the effect observed in the “fact” condition, the effect in the “quiz” condition is of a similar order of magnitude.
+The implication effect is credibly greater than zero.
 
-Surprisingly, the effects of exposure on participants’ truth ratings for the premise statements were very subtle.
+Somewhat surprisingly, the effects of exposure on participants’ truth ratings for the premise statements were very subtle.
 Our primary preregistered test compared participants’ truth ratings for the exposed premise statements at test against their ratings for another set of 8 unexposed premise statements.
@@
 
-The small size of these illusory truth effects may owe to the use of a “quiz” judgment task at exposure as well as the relatively brief period of distraction between exposure and test used in Experiment 1.
-By first responding to the statements in a quiz, participants may then have felt pressure to maintain consistency in their responses, and may have still possessed some explicit memory of their responses for the premise statements.
-If so, this could have suppressed the influence of familiarity traditionally argued to produce the illusory truth effect.
-
-More importantly, a drive to maintain consistency might also explain our surprising findings of an illusory implication effect.
+Although the "quiz" presentation used in Experiment 1 is traditionally seen as providing "mere exposure" to statements, asking participants to judge the truth of these items may have lead them to feel pressured to maintain consistency in their responses.
 Participants showed a general bias toward judging statements “true” during the exposure phase, responding with some variation of “true” for 71.8% of responses.
-Attempting to maintain consistency with these prior responses could have thus influenced their responses to the implication statements, rendering the observed effect an artifact of the experimental context.
+Attempting to maintain consistency with these prior responses could have thus influenced their responses to the implication statements and thus the observed effect may be an artifact of the experimental context.
 
@@
 Experiment 2 was conducted to address the possibility that a pressure for internal consistency produced the effects on the implication-statement truth ratings observed in Experiment 1.
-Preregistration informationExperiment 2 can be found at https://osf.io/czkdh/.
 
-```{r fig2, fig.env="figure", fig.pos = "t", fig.align = "center", fig.height=2.5, fig.width=2.5, fig.cap="Average truth ratings for Implication statements in Experiment 2, broken down by exposure (exposed vs.
+```{r fig2, fig.env="figure", fig.pos = "t", fig.align = "center", fig.cap="Average truth ratings for Implication statements in Experiment 2, broken down by exposure (exposed vs.
 unexposed) and implication type (true-implied-false vs false-implied-true).
@@
 
-A total of 300 participants were recruited from Amazon Mechanical Turk using procedures identical to Experiment 1.
+A total of 300 participants were recruited from CloudResearch using procedures identical to Experiment 1.
 As before, participants who failed attention check questions or who indicated they had looked up answers were excluded from analyses, leaving a final sample of `r nrow(df2_demo)` (`r exp2_n_f` female, median age `r exp2_age` years-old).
@@
 Secondly, to provide a longer period of distraction between exposure and test, the ordering of the items in the test phase was rearranged.
-Participants first rated 24 control items (12 true and 12 false) to extend the time between rating main test items.
+Participants first judged the truth of 24 control items (12 true and 12 false) to extend the time between viewing the main test items.
 
@@
 
-## Results and discussion
+## Results
 
@@
 
-fig3 # fix legend
+# fig3 # fix legend
 ```
@@
 
-Interestingly, though somewhat smaller, the illusory implication effect for the implication statements was generally similar in magnitude to the illusory truth effect observed among the premise statements.
+The illusory implication effect for the implication statements was generally similar in magnitude to the illusory truth effect observed among the premise statements.
 
-One limitation of these studies relative to prior work on the illusory truth effect is the relatively limited number of statements participants saw and were tested on.
-Generating diverse pairs of statements with a clear implication relation is relatively more challenging than simply generating false statements.
-Fortunately as shown in Figure 3, the illusory implication effect was fairly robust across the majority of the 24 different individual item pairs in Experiment 2.
+# Experiment 3
+
+Finally, we sought to replicate these findings once again and to rule out the possibility that the illusory implication effect is a product of experimental idiosyncrasies, such as misapprehension of instructions or potential task demands.
+In Experiment 3, we ensured participants were attending carefully and understood the instructions through additional attention and comprehension checks.
+In addition, we evaluated the potential for unanticipated task demands by excluding from analysis any participants who had suspicions about the study or who guessed its true purpose.
+
+## Method
+
+### Participants
+
+We recruited 540 participants from CloudResearch's Connect service and excluded participants who failed attention or instruction comprehension checks, leaving a final sample of `r nrow(df3_demo)` (`r exp3_n_f` female, median age `r exp3_age` years-old).
+In a secondary set of pre-registered analyses, we also excluded participants who suspected that the study had an undisclosed purpose (`r exp3_susp_count`) and additionally those who guessed the true purpose of the study in an open-ended response (`r exp3_flagged_count`), leaving `r nrow(df3_demo) - exp3_susp_count - exp3_flagged_count` participants for these analyses.
+
+### Materials and procedures
+
+All item pairs were identical to Experiment 2.
+However, several changes were made to the study's procedures limit the potential that the illusory implication effect could be produced by idiosyncratic features of the experimental task.
+
+First, we sought to further ensure that participants would not mistake the exposure context for a communicative context, while also not inducing any consistency pressure at test.
+To this end, we created a new "believability" task for the exposure phase.
+Participants were instructed to carefully read two statements and select the one that they thought would be more believable to the average person.
+This task should provide a pragmatic context for mere exposure where it was explicitly and pragmatically clear that the statements could be either true or false, but without inducing the consistency pressure that might come with a true or false “quiz”.
+To further ensure this task provided mere exposure, participants were reminded that statements could be true or false each time they judged their believability and comprehension check questions were included before and after the exposure phase to ensure participants remembered this aspect of the task.
+
+To improve our ability to measure the implication effect, in Experiment 3 participants were exposed to 12 of the premise statements at exposure rather than eight.
+As a consequence, we did not examine illusory truth effects.
+The 24 premise statements were randomly sorted in to 12 comparison questions, and each participant saw 6 of these pairs at exposure, along with five control pairs.
+
+Following the test phase, two new questions assessed the potential influence of task demands.
+Participants were asked if they ever suspected the study had an undisclosed purpose (yes/no), and then regardless of their answer asked to explained their suspicions in an open-ended response.
+Responses were coded to identify participants who guessed that the true purpose concerned how their responses to the test phase would be affected by the exposure phase or by seeing misinformation.
+These participants were excluded from certain analyses following our pre-registered analysis plan.
+
+Finally, six attention check questions were more tightly integrated into the task to better ensure attendant responding.
+Following the "infrequency-items" approach of [@zorowitz.etal2023], exposure phase attention checks consisted of a very well-known fact and one wholly unbelievable fiction.
+Test phase attention checks consisted of very easy trivia items that all participants should know.
+Participants who answered any of the six attention checks incorrectly were excluded from analyses.
+
+## Results
+
+```{r fig-exp3, fig.align='center', fig.cap="Average truth ratings for Implication statements in Experiment 3, broken down by exposure (exposed vs.
+unexposed) and implication type (true-implied-false vs false-implied-true).
+For visualization purposes, means were calcaulated by translating ordinal responses onto a scale from -2.5 to 2.5.
+Error bars indicate standard errors."}
+
+df3 %>% 
+  filter(phase=="test") %>% 
+  filter(item_type == "B") %>% 
+  mutate(
+    exposed = if_else(exposed, "Exposed", "Unexposed")
+  ) %>% 
+  group_by(exposed, B_implied_true) %>% 
+  summarize(
+    M = mean(resp_num),
+    se = sd(resp_num)/sqrt(n()),
+    ll = M - se,
+    ul = M + se
+  ) %>% 
+  ggplot(
+    aes(x=B_implied_true, color=factor(exposed), shape=factor(exposed), y = M, ymin = ll, ymax=ul)
+  ) +
+  geom_pointrange(position=position_dodge(width=.5))  +
+  scale_color_manual(values = c("#da1e28", "#0f62fe")) +
+  theme_bw(base_size=10) +
+  theme(legend.position="bottom", panel.grid=element_blank()) +
+  # coord_flip() +
+  labs(y = "Truth rating", x = "Implication type", color="Exposure", shape="Exposure")
+```
+
+### Preregistered analyses
+
+First, in accordance with our preregistered analysis plan, we replicate the primary "maximal" random-effects regression analysis of Experiments 1 and 2.
+Table 3 presents the results of this regression.
+The illusory implication effect was replicated, with a magnitude of the implication coefficient very similar to that observed in Experiment 1.
+
+Next, to explore the possibility that these results might be influenced by task demands, we also carried out a parallel analysis excluding participants whose responses may have been impacted by their suspicions as to the true purpose of the study.
+We excluded both participants who explicitly stated they were suspicious when asked directly, and those who forced to guess in an open-ended response, guessed that the true purpose had to do with the first phase somehow affecting their responses in the second.
+Removing these participants did not substantially change the estimate of the implication effect (stats).
+
+```{r, showtable3, echo=F, message=F}
+mytable3
+```
+
+### Secondary and exploratory analyses
+
+In Experiments 1 and 2, the effect of familiarity was weaker than the measured implication effect, so that average truth ratings were higher after exposure for false-implied-true items and lower for true-implied-false items (Figures 1 and 2).
+In Experiment 3 however, the estimated effect of familiarity and implication were similar, and truth ratings were similar for exposed and unexposed true-implied-false items (Figure 3).
+This may suggest a possible alternate explanation of our findings: that the true-implied-false items' premise statements simply do not engender as much familiarity or fluency for their implication as do the false-implied-true items, so that the effects differ between these groups.
+Put another way, the identification of the "implication" effect coefficient rests on the assumption that familiarity and fluency effects are not accidentally confounded with item groups.
+
+To assess the potential for confounding in the familiarity or fluency that exposure to the premise statements would likely engender for the implication statements, we examined the degree of surface-level overlap within each item pair.
+We compared the proportion of shared words (after lemmatization) between premise and implications within each item pair.
+As shown in Figure 4, there is no evidence for confounding between surface-level similarity and item types (median overlap = .19 in false-implied-true items, .21 in true-implied-false items) that would suggest differential effects of fluency or familiarity across types.
+
+Taken in concert with findings from Experiments 1 and 2, where the effect of exposure differed in sign across the item groups, we interpret our regressions as identifying and estimating an implication effect.
+
+```{r, item-fluency-print-fig, fig.align = "center", fig.cap = "Word-level overlap between premise and implication statements (after lemmatization)."}
+item_fluency_fig
+```
 
@@
 
-We observed an “illusory implication” effect across two preregistered experiments: mere exposure to premises (e.g.
+We observed an “illusory implication” effect across three preregistered experiments: mere exposure to premises (e.g.
 that “The Tour de France has been held every year since its inception in 1903”) influenced participant’s truth ratings for examples of their logical implications (e.g.
@@
 Unlike the illusory truth effect, this novel effect cannot be easily explained by the effects of familiarity or fluency on the judgment process [cf.
-@fazio.etal2015; @unkelbach2007; @wang.etal2016]: whereas familiarity would be expected to uniformly increase endorsements, participants’ truth ratings for implication statements were both increased and decreased according to the implications of the corresponding premises.
-Thus, it appears that mere exposure has a genuine impression on people’s beliefs beyond simply creating a sense of familiarity.
+@arkes.etal1991; @fazio.etal2015; @unkelbach2007; @wang.etal2016]: whereas these metacognitive cues would be expected to uniformly increase endorsements, participants’ truth ratings for implication statements were both increased and decreased according to the implications of the corresponding premises.
+Thus, it appears that mere exposure has a genuine effect on people’s beliefs beyond simply creating a sense of familiarity.
 
-If not familiarity, then what cognitive mechanism could explain these findings?
-One potential explanation is that mere exposure to the premise statements essentially leads participants to at least partially adopt the premise statements as beliefs.
-Simply entertaining the idea may create such an impression despite awareness of its ambiguous truth-value.
-Or, consistent with a general overriding expectation of truth in testimony [e.g.
-@grice1989; @levine2014], participants may at least partially accept the premise statements as true despite the experimental context making clear they should not.
-This explanation would be consistent with the general truth-bias observed throughout both studies.
-Future research might vary the plausibility of the premise statements to evaluate this explanation or identify its boundary conditions [@fazio.etal2019].
+What cognitive mechanism could explain these findings?
+One potential explanation is that mere exposure to the premise statements leads participants to at least partially adopt them as beliefs.
+Simply entertaining the idea may create such an impression despite awareness of its ambiguous truth-value [@gilbert.etal1990; @gilbert1991].
+If so, this apparently occurs despite participant's awareness that the statements may be false, as ensured by constant reminders and comprehension checks in Experiment 3.
 
-Another potential explanation would attribute the illusory implication effect to processes that play out in the testing phase.
-Rather than familiarity, the illusory implication effect may instead reflect explicit memory and source misattribution: Participants may (at least partially) remember the premise statements, but fail to properly attribute their source to the experimental context.
+Alternately, the illusory implication effect may reflect explicit memory and source misattribution during testing: Participants may have some explicit memory for the premise statements, but fail to properly attribute their source to the experimental context.
 Future research might explore how explicit memory for the premise statements correlates with the strength of the illusory implication effects.
-In particular, research might disentangle memory effects by exploring factors that are known to affect memory, but that should not affect other reasoning processes, such as retention intervals or serial-position effects [e.g.
-@bjork.whitten1974].
 
-Finally, given the surprising nature of our findings, we must raise the possibility of potential deflationary explanations: that the finding could be some kind of experimental demand characteristic.
-For instance, participants’ might imagine there is some kind of trick so that they are expected to reason forward from the initial exposure to the items.
-Or, some proportion might have been sufficiently confused so as to imagine they were meant to believe the statements on their initial exposure (though it is hard to imagine this in Experiment 1).
-Further work should explore participants' perceptions of the purposes of these studies and the potential for different response strategies.
-However, it should be noted that similar concerns could be levied at essentially all prior research on the illusory truth effect.
+Whatever the cognitive mechanism behind the illusory implication effect we observed, these findings are further cause for concern about the spread and influence of misinformation.
+The spread of misinformation and fake news appears to outpace the spread of factual news [@vosoughi.etal2018] and creates massive challenges for online platforms seeking to remove, label, and correct misinformation [@sharma.etal2019].
+These challenges may be further compounded by basic features of human cognition  that our studies have revealed: mere exposure to false statements can lead them to be _believed_.
 
-Whatever the cognitive mechanism behind the illusory implication effect we observed, these findings are further cause for concern about the spread of misinformation on and offline.
-The spread of misinformation and fake news may outpace the spread of factual news [@vosoughi.etal2018] and creates massive challenges for online platforms seeking to remove, label, and correct misinformation [@sharma.etal2019].
-These challenges may be further compounded by basic features of human cognition: mere exposure to false statements can make them appear true.
-Further, as our findings are the first to indicate, these impacts generalize to related statements and are potentially indicative of genuine belief following mere exposure.
+\newpage
 
@@
 
-```{r}
-# References will be generated automatically by Pandoc and included here.
-# The following code is some latex to format the bibliography. Do not remove it.
-```
-
-\setlength{\parindent}{-0.1in} 
-\setlength{\leftskip}{0.125in}
-\noindent
+::: {#refs custom-style="Bibliography"}
+:::

### 2023-08-04 cogsci-paper/random-slopes-models.R

@@ -1,0 +1,150 @@
+fit_prem_exp2_rs <- brm(
+  response ~ exposed + (1 + exposed|subj_id) + (1|item_pair),
+  data = df2 %>%
+    filter(phase=="test", item_type=="A"),
+  family = cumulative(),
+  chains = 4,
+  cores = 4,
+  file="../local/fit_prem_exp2_rs"
+)
+
+summary(fit_prem_exp2_rs)
+
+glimpse(posterior_samples(fit_prem_exp2_rs))
+
+library(tidybayes)
+
+fit_prem_exp2_rs %>% 
+  spread_draws(r_subj_id[subj_id,exposedTRUE],  b_exposedTRUE) %>% 
+  mutate(beta = r_subj_id + b_exposedTRUE) %>% 
+  group_by(subj_id) %>% 
+  summarize(
+    M = mean(beta),
+    ul = quantile(beta, .75),
+    ll = quantile(beta, .25),
+  ) %>% 
+  ggplot(aes(x=reorder(subj_id,M), y = M, ymin = ll, ymax=ul)) +
+  geom_pointrange()
+
+fit_prem_exp2_rs %>% 
+  spread_draws(r_subj_id[subj_id,exposedTRUE],  b_exposedTRUE) %>% 
+  mutate(beta = r_subj_id + b_exposedTRUE) %>% 
+  group_by(subj_id) %>% 
+  summarize(
+    M = mean(beta),
+    ul = quantile(beta, .75),
+    ll = quantile(beta, .25),
+  ) %>% 
+  ggplot(aes(x=M)) +
+  geom_histogram()
+
+## -- implications exp 2
+
+fit_imp_exp2_rs <- brm(
+  response ~ familiarity + implication + B_implied_true + (1 + familiarity + implication|subj_id) + (1+ familiarity + implication|item_pair),  # familiarity + implication + grouping (optional)
+  data = df_reg2,
+  family = cumulative(),
+  chains = 4,
+  cores = 4,
+  file="../local/fit_imp_exp2_rs"
+)
+
+summary(fit_imp_exp2_rs)
+
+glimpse(posterior_samples(fit_imp_exp2_rs))
+
+# fit_imp_exp2_rs %>% 
+#   spread_draws(r_subj_id[subj_id,implication],  b_implication) %>% 
+#   mutate(beta = r_subj_id + b_implication) %>% 
+#   group_by(subj_id) %>% 
+#   summarize(
[…]

### 2023-08-18 paper/paper-papaja.Rmd
small pre submission edits

@@
 ---
-title             : "The title"
-shorttitle        : "Title"
+title             : "Illusory Implications: Evidence that mere exposure to ideas can induce beliefs"
+shorttitle        : "Illusory implications"
 
@@
 
-Next, to explore the possibility that these results might be influenced by task demands, we also carried out a parallel analysis excluding participants whose responses may have been impacted by their suspicions as to the true purpose of the study.
+Next, to explore the possibility that these results might be influenced by task demands, we also carried out a parallel analysis excluding participants whose responses may have been impacted by their suspicions as to the true purpose of the study [@coles.frank2023].
 We excluded both participants who explicitly stated they were suspicious when asked directly, and those who forced to guess in an open-ended response, guessed that the true purpose had to do with the first phase somehow affecting their responses in the second.
-Removing these participants did not substantially change the estimate of the implication effect (stats).
+Removing these participants did not substantially change the estimate of the implication effect, $\beta =$ `r report_reg_coef(fit_imp_maximal3, "implication")`.
 

### 2023-08-18 paper/paper-papaja.Rmd
word count added

@@
 ---
-title             : "Illusory Implications: Evidence that mere exposure to ideas can induce beliefs"
+title             : "Illusory Implications: Mere exposure to ideas can induce beliefs"
 shorttitle        : "Illusory implications"
@@
 keywords          : "Illusory truth; Metacognition; Cognitive Psychology; Misinformation"
-wordcount         : "X"
+wordcount         : "4929"
 

### 2023-08-22 paper/paper-papaja.Rmd
fixed typos

@@
 keywords          : "Illusory truth; Metacognition; Cognitive Psychology; Misinformation"
-wordcount         : "4929"
+wordcount         : "4930"
 
@@
 ```{r fig1, fig.env="figure", fig.pos = "h", fig.align = "center", fig.cap="Average truth ratings for Implication statements in Experiment 1, broken down by exposure (exposed vs.
-unexposed), instructions (fact vs quiz conditions), implication type (true-implied-false vs false-implied-true).
-For visualization purposes, means were calcaulated by translating ordinal responses onto a scale from -2.5 to 2.5.
+unexposed), instructions (fact vs quiz conditions), implication type (false-implied-true vs true-implied-false).
+For visualization purposes, means were calculated by translating ordinal responses onto a scale from -2.5 to 2.5.
 Error bars indicate standard errors.
@@
 ```{r fig2, fig.env="figure", fig.pos = "t", fig.align = "center", fig.cap="Average truth ratings for Implication statements in Experiment 2, broken down by exposure (exposed vs.
-unexposed) and implication type (true-implied-false vs false-implied-true).
-For visualization purposes, means were calcaulated by translating ordinal responses onto a scale from -2.5 to 2.5.
+unexposed) and implication type (false-implied-true vs true-implied-false).
+For visualization purposes, means were calculated by translating ordinal responses onto a scale from -2.5 to 2.5.
 Error bars indicate standard errors."}
@@
 ```{r fig3, fig.env="figure*", fig.pos = "!hbt", fig.align = "center", fig.width=6.5, fig.height=3, fig.cap="Average truth ratings for individual premise and implication statements in Experiment 2, broken down by exposure (exposed vs.
-unexposed) and implication type (true-implied-false vs false-implied-true).
-For visualization purposes, means were calcaulated by translating ordinal responses onto a scale from -2.5 to 2.5.
+unexposed) and implication type (false-implied-true vs true-implied-false).
+For visualization purposes, means were calculated by translating ordinal responses onto a scale from -2.5 to 2.5.
 Error bars indicate standard errors.", num.cols.cap=2}
@@
 All item pairs were identical to Experiment 2.
-However, several changes were made to the study's procedures limit the potential that the illusory implication effect could be produced by idiosyncratic features of the experimental task.
+However, several changes were made to the study's procedures to limit the potential that the illusory implication effect could be produced by idiosyncratic features of the experimental task.
 
@@
 As a consequence, we did not examine illusory truth effects.
-The 24 premise statements were randomly sorted in to 12 comparison questions, and each participant saw 6 of these pairs at exposure, along with five control pairs.
+The 24 premise statements were randomly sorted into 12 comparison questions, and each participant saw 6 of these pairs at exposure, along with five control pairs.
 
 Following the test phase, two new questions assessed the potential influence of task demands.
-Participants were asked if they ever suspected the study had an undisclosed purpose (yes/no), and then regardless of their answer asked to explained their suspicions in an open-ended response.
+Participants were asked if they ever suspected the study had an undisclosed purpose (yes/no), and then regardless of their answer asked to explain their suspicions in an open-ended response.
 Responses were coded to identify participants who guessed that the true purpose concerned how their responses to the test phase would be affected by the exposure phase or by seeing misinformation.
@@
 ```{r fig-exp3, fig.align='center', fig.cap="Average truth ratings for Implication statements in Experiment 3, broken down by exposure (exposed vs.
-unexposed) and implication type (true-implied-false vs false-implied-true).
-For visualization purposes, means were calcaulated by translating ordinal responses onto a scale from -2.5 to 2.5.
+unexposed) and implication type (false-implied-true vs true-implied-false).
+For visualization purposes, means were calculated by translating ordinal responses onto a scale from -2.5 to 2.5.
 Error bars indicate standard errors."}
@@
   ) %>% 
+  mutate(B_implied_true = ordered(B_implied_true, levels = c("F implied T", "T implied F"))) %>% 
   ggplot(
@@
 Next, to explore the possibility that these results might be influenced by task demands, we also carried out a parallel analysis excluding participants whose responses may have been impacted by their suspicions as to the true purpose of the study [@coles.frank2023].
-We excluded both participants who explicitly stated they were suspicious when asked directly, and those who forced to guess in an open-ended response, guessed that the true purpose had to do with the first phase somehow affecting their responses in the second.
+We excluded both participants who explicitly stated they were suspicious when asked directly, and those who, when forced to guess in an open-ended response, guessed that the true purpose had to do with the first phase somehow affecting their responses in the second.
 Removing these participants did not substantially change the estimate of the implication effect, $\beta =$ `r report_reg_coef(fit_imp_maximal3, "implication")`.

### 2023-10-25 paper/paper-papaja.Rmd
minor revisions and prep for submission

@@
 
-  Materials for all experiments can be found at https://osf.io/znq3y/.
+  Materials, data, and analysis code for all experiments can be found at https://osf.io/znq3y/.
 Preregistration information for Experiments can be found at https://osf.io/c4w8s/ (Exp.
@@
 3).
+Experiments 1 and 2 were first reported in a six-page paper presented at the _43rd Annual Meeting of the Cognitive Science Society_ and appearing in the conference's _Proceedings_.
+This non-archival conference paper included similar narrative motivations for the studies as presented here, with a more limited literature review.
+The narrative interpretations and discussion have changed in light of the findings of the novel third experiment and additional analyses reported here.
 
 abstract: |
-  Numerous psychological findings have shown that mere exposure to ideas makes those ideas seem more true, a finding commonly referred to as the “illusory truth” effect [e.g.
-@hasher.etal1977].
-In the presence of pervasive misinformation, this basic feature of cognition may undermine the functioning of a democratic society [@pennycook.etal2018].
-However, genuine beliefs do not only affect judgments about individual statements, they also imply other beliefs and drive decision-making.
-Here, we sought to examine whether mere exposure to statements produces genuine beliefs by examining whether people draw inferences from statements after mere exposure.
-Surprisingly, and in contrast to fluency or familiarity-based accounts of the illusory truth effect [e.g.
-@dechene.etal2010], we found that exposure to “premise” statements affected participants’ truth ratings for novel “implied” statements.
-This “illusory implication” effect suggests that the consequences of exposure to false statements reach further than previously thought and calls for a new mechanistic account of these effects.
+  Numerous psychological findings have shown that mere exposure to ideas makes those ideas seem more true, a finding commonly referred to as the “illusory truth” effect.
+This basic feature of cognition may hamper efforts to curb misinformation: critical media consumption and fact-checking seem poised to fail if mere exposure can persuade people to adopt ideas.
+Under predominant accounts of the illusory truth effect, initial exposure to a statement provides a metacognitive feeling of “fluency” or familiarity that, upon subsequent exposure, leads people to infer the statement is more likely to be true.
+However, genuine beliefs do not only affect truth judgments about individual statements, they also imply other beliefs and drive decision-making.
+Here, we conducted three pre-registered experiments to examine whether mere exposure to statements produces genuine beliefs by examining whether people draw inferences from statements after mere exposure.
+Surprisingly, and in contrast to fluency or familiarity-based accounts of the illusory truth effect, we found that exposure to “premise” statements affected participants’ truth ratings for novel “implied” statements.
+This “illusory implication” effect suggests that the consequences of exposure to mis- and disinformation reach further than previously thought and calls for a new mechanistic account of these effects.
 
@@
 keywords          : "Illusory truth; Metacognition; Cognitive Psychology; Misinformation"
-wordcount         : "4930"
+wordcount         : "5099"
 
@@
 
-## Relevance Statement
-
-Three pre-registered experiments examined whether mere exposure to ideas might lead them to be adopted as beliefs.
-It is well-established that mere exposure, such as in a true-or-false quiz, can lead people to later rate statements as more true (the “illusory truth” effect).
-If people's beliefs are affected by mere exposure, then the pervasiveness of misinformation would be even more harmful than currently recognized.
-However, leading theories attribute illusory truth effects to feelings of familiarity or fluency rather than changes in beliefs.
-These metacognitive heuristics are unlikely to produce further inferences or drive decision-making, as would beliefs.
-In contrast, we demonstrate that mere exposure to statements affected participants' responses to other novel statements (what we term an “illusory implication” effect), suggesting that mere exposure can lead to genuine belief.
-Our findings illustrate new dangers in exposure to misinformation, showing that mere exposure may have more wide-ranging influences than previously understood.
+## Public Significance Statement
+
+It is well-established that mere exposure to statements, such as in a true-or-false quiz, can lead people to later rate those same statements as more true (the “illusory truth” effect).
+On most accounts, the illusory truth effect is thought to be caused by fluency or familiarity with a specific statement, and therefore should not more widely affect people's beliefs or decisions in most cases.
+In contrast to these prior accounts, however, we found that mere exposure to statements affected participants' responses to other novel statements (what we term an “illusory implication” effect), suggesting that mere exposure can instead lead to genuine belief.
+Our findings are further cause for concern about the spread and influence of misinformation, as they indicate that any exposure to misinformation may cause harm by inducing false beliefs.
 
@@
 There is something unsettling about the illusory truth effect and the lack of agency it implies over our own beliefs.
-Even more worrisome, there could be dire societal implications if merely being exposed to an idea causes people to adopt it as a belief: Pennycook and colleagues [-@pennycook.etal2018] argue that the illusory truth effect, combined with an environment of pervasive misinformation, has important implications for the functioning of democratic society.
+Even more worrisome, there could be important societal implications if merely being exposed to an idea causes people to adopt it as a belief: Pennycook and colleagues [-@pennycook.etal2018] argue that the illusory truth effect, combined with an environment of pervasive misinformation, has important consequences for the functioning of democratic society.
 The illusory truth effect suggests that mere exposure to misinformation could have impacts that cannot be stopped by fact-checking labels [@pennycook.etal2018], nor effectively curbed by retractions or corrective information [c.f.
@@
 
-A handful of studies are suggestive in this regard: Arkes and collagues [-@arkes.etal1991] found that mere exposure to one set of statements also affected judgments about novel related statements on the same topic (China).
-However, these findings could also be explained by a sense of familiary engendered by the shared topic of old and new statements [@arkes.etal1991].
+A handful of studies are suggestive in this regard: Unkelbach and Rom [-@unkelbach.rom2017] found that mere exposure to one statement could increase endorsements of implied statements.
+For instance, being more likely to endorse the statement "Most accidents occur close to weekends" after exposure to the statement "Most accidents occur on Mondays".
+Similarly, Arkes and colleagues [-@arkes.etal1991] found that mere exposure to one set of statements also affected judgments about novel related statements on the same topic (China).
+However, in both cases these findings can also be easily explained by fluency or familiarity engendered by the similarity between the presented and tested statements [@arkes.etal1991; @unkelbach.rom2017].
 
@@
 Such an “illusory implication” effect would indicate that mere exposure has genuine impacts on beliefs extending beyond those attributable to metacognitive cues.
+
+# Transparency and Openness
+
+All experiments were preregistered prior to data collection and all materials, data, and analysis code have been made publicly available on the Open Science Foundation website (https://osf.io/znq3y/).
+We report our approach to determining sample sizes, all data exclusions, all manipulations, and all measures in each study.
+Where appropriate, we explicitly differentiate between preregistered analyses and non-preregistered analyses in our results.
+All statistical models were fit using the `brms` R package [@burkner2017].
+Our study was reviewed and approved by the Arizona State University Institutional Review Board under project identifier `STUDY00013322`.
+Participants provided informed consent before beginning the study.
 
@@
 
+## Constraints on Generality
+
+The studies reported here targeted U.S.
+adults through an online survey provider (CloudResearch).
+Study materials were designed accordingly and may produce different results if tested on populations with different backgrounds and prior knowledge.
+
 \newpage

### 2024-04-03 paper/paper-papaja.Rmd
paper updates

@@
 
-## Public Significance Statement
-
-It is well-established that mere exposure to statements, such as in a true-or-false quiz, can lead people to later rate those same statements as more true (the “illusory truth” effect).
-On most accounts, the illusory truth effect is thought to be caused by fluency or familiarity with a specific statement, and therefore should not more widely affect people's beliefs or decisions in most cases.
-In contrast to these prior accounts, however, we found that mere exposure to statements affected participants' responses to other novel statements (what we term an “illusory implication” effect), suggesting that mere exposure can instead lead to genuine belief.
-Our findings are further cause for concern about the spread and influence of misinformation, as they indicate that any exposure to misinformation may cause harm by inducing false beliefs.
-
-\newpage
-
 Every day, people are faced with a barrage of unsupported claims, from pestering ad campaigns to blatant disinformation on social media.
@@
 
+## Theories of the Illusory Truth Effect
+
 Why should mere exposure affect later ratings of truth?
@@
 
+## Mere Exposure and Belief
+
 There is something unsettling about the illusory truth effect and the lack of agency it implies over our own beliefs.
@@
 For one, beliefs imply other beliefs: for instance, believing that “it is sunny out” implies the belief that “it is daytime”.
-And believing misinformation, such as “Dominion voting machines were rigged in the 2020 election” might imply the belief that “Donald Trump actually won the 2020 election.” Beliefs also inform action and decision-making, and whether or not someone holds a belief can be judged (at least partly) from their decisions.
+And believing misinformation, such as “Dominion voting machines were rigged in the 2020 U.S.
+election” might imply the belief that “Donald Trump actually won the 2020 U.S.
+election.” Beliefs also inform action and decision-making, and whether or not someone holds a belief can be judged (at least partly) from their decisions.
 When truly believed, misinformation can have serious consequences.
@@
 Under these theoretical accounts, the impact of prior exposure depends on a subsequent re-exposure to a now fluent or familiar statement.
-This would make the real-world consequences of the illusory truth effect far narrower than one might fear: These metacognitive mechanisms are liable to apply only in response to the exact statements previously seen, but are not as likely to generalize to other related ideas, or to drive actions or decision-making in novel contexts.
+This would make the real-world consequences of the illusory truth effect far narrower than one might otherwise fear: These metacognitive mechanisms are liable to apply only in response to the exact statements previously seen, but are not as likely to generalize to other related ideas, or to drive actions or decision-making in novel contexts.
 
@@
 
+## Is there an "Illusory Implication" Effect?
+
 A crucial feature of belief is that beliefs are connected with other related beliefs.
 Thus, one way to test whether mere exposure affects beliefs is to examine whether mere exposure to statements can affect truth ratings for other related statements.
+We refer to this hypothetical impact of mere exposure on implied statements as the "Illusory Implication" Effect.
 
 A handful of studies are suggestive in this regard: Unkelbach and Rom [-@unkelbach.rom2017] found that mere exposure to one statement could increase endorsements of implied statements.
-For instance, being more likely to endorse the statement "Most accidents occur close to weekends" after exposure to the statement "Most accidents occur on Mondays".
+For instance, people were more likely to endorse the statement "Most accidents occur close to weekends" after exposure to the statement "Most accidents occur on Mondays".
 Similarly, Arkes and colleagues [-@arkes.etal1991] found that mere exposure to one set of statements also affected judgments about novel related statements on the same topic (China).
@@
 
-Bacon [-@bacon1979] found that mere exposure to one statement ($P$) could lead to lower truth ratings for a new, contradictory statements($\neg P$).
+A case where mere exposure lead to a lower truth rating for a related statement (i.e., one implied to be false) would more clearly implicate an effect of mere exposure on beliefs.
+In our review of the literature, we find three reported experiments that offer preliminary evidence of this effect: Bacon [-@bacon1979] found that mere exposure to one statement ($P$) could lead to lower truth ratings for a new, contradictory statements($\neg P$) after a three-week delay.
 For instance, mere exposure to "crocodiles sleep with their eyes open" produced lower truth ratings for "crocodiles sleep with their eyes closed".
-It seems clear at least that familiarity cannot explain this pattern of results.
-However, other findings under very similar paradigms have partly contradicted these results [@garcia-marques.etal2015] and it is somewhat unclear whether fluency could account for these findings, i.e.
-whether contradictory statements can be expected to lead to fluent or disfluent processing.
-
-Here, we sought to decisively examine whether mere exposure to statements could influence people's beliefs.
+Findings from an experiment by Garcia-Marques and colleagues [-@garcia-marques.etal2015] partially contradicted these results (using a slightly different exposure task), finding lower truth ratings for contradictory statements when participants were queried immediately, but finding the opposite pattern of results at the delay used by the original Bacon [-@bacon1979] study.
+A later series of experiments by Unkelbach and Rom [-@unkelbach.rom2017] replicated these new findings [@garcia-marques.etal2015], although the effects were not observed consistently in each experiment.
+
+## The present studies
+
+Here, we sought to expand upon prior experiments to decisively examine whether mere exposure to statements could influence people's beliefs by testing for an "illusory implication" effect.
 To do so, we examined whether people draw inferences from statements after exposure in a context without any presupposition of truth.
-We designed a study to examine whether mere exposure to one statement (the “premise”) could affect truth ratings for a  different statement the premise would logically entail (the “implication”).
-Such an “illusory implication” effect would indicate that mere exposure has genuine impacts on beliefs extending beyond those attributable to metacognitive cues.
+We designed a series of studies to examine whether mere exposure to one statement (the “premise”) could affect truth ratings for a  different statement the premise would logically entail (the “implication”).
+Crucially, we test this across cases where premises entail either the truth or falsehood of another related statement, ruling out purely metacognitive explanations.
+
+In addition, we make several additional contributions toward definitively demonstrating the illusory implication effect beyond existing studies by 1) developing a new set of materials, 2) recruiting larger samples (present studies $N = 1020$, total prior cited studies $N = 452$), 3) utilizing modern hierarchical statistical modeling methods, and 4) taking great pains to ensure that participants' truth ratings are not influenced by task demands (Experiments 2 and 3).
+Across our studies, we find new compelling evidence for this “illusory implication” effect, indicating that mere exposure has genuine impacts on beliefs extending beyond those attributable to metacognitive cues.
 
@@
 In Experiment 3 however, the estimated effect of familiarity and implication were similar, and truth ratings were similar for exposed and unexposed true-implied-false items (Figure 3).
-This may suggest a possible alternate explanation of our findings: that the true-implied-false items' premise statements simply do not engender as much familiarity or fluency for their implication as do the false-implied-true items, so that the effects differ between these groups.
+Although the estimate for the implication effect in our statistical model was clearly greater than zero, this may nevertheless allow a possible alternate explanation of our findings: that the true-implied-false items' premise statements simply do not engender as much familiarity or fluency for their implication as do the false-implied-true items, so that the effects differ between these groups.
 Put another way, the identification of the "implication" effect coefficient rests on the assumption that familiarity and fluency effects are not accidentally confounded with item groups.
@@
 One potential explanation is that mere exposure to the premise statements leads participants to at least partially adopt them as beliefs.
-Simply entertaining the idea may create such an impression despite awareness of its ambiguous truth-value [@gilbert.etal1990; @gilbert1991].
+Simply entertaining the idea may create such an impression despite awareness of its ambiguous truth-value [@gilbert.etal1990; @gilbert1991; @garcia-marques.etal2015].
 If so, this apparently occurs despite participant's awareness that the statements may be false, as ensured by constant reminders and comprehension checks in Experiment 3.

### 2024-04-03 paper/paper-papaja.Rmd
ethics + wordcount

@@
 
+  **Ethics Statement**: This research was conducted under the approval of the Institutional Review Board at Arizona State University, IRB protocol ID STUDY00013322.
+Informed consent was obtained from all participants.
+
 abstract: |
@@
 keywords          : "Illusory truth; Metacognition; Cognitive Psychology; Misinformation"
-wordcount         : "5099"
+wordcount         : "5209"
 
@@
 A case where mere exposure lead to a lower truth rating for a related statement (i.e., one implied to be false) would more clearly implicate an effect of mere exposure on beliefs.
-In our review of the literature, we find three reported experiments that offer preliminary evidence of this effect: Bacon [-@bacon1979] found that mere exposure to one statement ($P$) could lead to lower truth ratings for a new, contradictory statements($\neg P$) after a three-week delay.
+In our review of the literature, we find three reported studies that offer preliminary evidence of this effect: Bacon [-@bacon1979] found that mere exposure to one statement ($P$) could lead to lower truth ratings for a new, contradictory statements($\neg P$) after a three-week delay.
 For instance, mere exposure to "crocodiles sleep with their eyes open" produced lower truth ratings for "crocodiles sleep with their eyes closed".
 Findings from an experiment by Garcia-Marques and colleagues [-@garcia-marques.etal2015] partially contradicted these results (using a slightly different exposure task), finding lower truth ratings for contradictory statements when participants were queried immediately, but finding the opposite pattern of results at the delay used by the original Bacon [-@bacon1979] study.
-A later series of experiments by Unkelbach and Rom [-@unkelbach.rom2017] replicated these new findings [@garcia-marques.etal2015], although the effects were not observed consistently in each experiment.
+A later series of experiments by Unkelbach and Rom [-@unkelbach.rom2017] generally replicated the newer findings [@garcia-marques.etal2015], although the effects were not observed consistently in each experiment.
 

### 2024-04-09 paper/paper-papaja.Rmd
rm constraints on generality
revisionsand concept fluency stuff

@@
   Numerous psychological findings have shown that mere exposure to ideas makes those ideas seem more true, a finding commonly referred to as the “illusory truth” effect.
-This basic feature of cognition may hamper efforts to curb misinformation: critical media consumption and fact-checking seem poised to fail if mere exposure can persuade people to adopt ideas.
 Under predominant accounts of the illusory truth effect, initial exposure to a statement provides a metacognitive feeling of “fluency” or familiarity that, upon subsequent exposure, leads people to infer the statement is more likely to be true.
 However, genuine beliefs do not only affect truth judgments about individual statements, they also imply other beliefs and drive decision-making.
-Here, we conducted three pre-registered experiments to examine whether mere exposure to statements produces genuine beliefs by examining whether people draw inferences from statements after mere exposure.
-Surprisingly, and in contrast to fluency or familiarity-based accounts of the illusory truth effect, we found that exposure to “premise” statements affected participants’ truth ratings for novel “implied” statements.
-This “illusory implication” effect suggests that the consequences of exposure to mis- and disinformation reach further than previously thought and calls for a new mechanistic account of these effects.
+Here, we consider whether exposure to “premise” statements affects people's truth ratings for novel “implied” statements, a pattern of findings a pattern of findings we call the "illusory implication" effect.
+We argue these effects would constitute evidence for genuine belief change from mere exposure and  identify a handful of existing findings that offer preliminary support for this claim.
+Building upon these, we conduct three new pre-registered experiments to further test this hypothesis, finding additional evidence that exposure to “premise” statements affected participants’ truth ratings for novel “implied” statements, including for considerably more distant implications than those previously explored.
+Our findings suggest that the effects of mere exposure reach further than previously thought, with potentially consequential implications for concerns around mis- and disinformation.
 
@@
 keywords          : "Illusory truth; Metacognition; Cognitive Psychology; Misinformation"
-wordcount         : "5209"
+wordcount         : "5636"
 
@@
 That is, it occurs when the statements are seen or heard in a non-communicative context, such as when they are read during a true/false quiz.
-Generally, being told something---even by a source of unknown trustworthiness---is prima facie reason for believing it [@grice1989].
+Generally, being told something---even by a source of unknown trustworthiness---is prima facie reason for believing it [@grice1989; also see @corneille.etal2020].
 But reading a statement on a true/false quiz is not reason for believing it; the statement is just as likely to be false as to be true.
@@
 Why should mere exposure affect later ratings of truth?
-Predominant theoretical accounts attribute illusory truth effects to metacognitive experiences of fluency or familiarity [@dechene.etal2010].
-Research on metacognition has shown that people's behavior and judgments are not only influenced by the content of their thoughts, but also by the phenomenological experience of processing those thoughts [For a review, see @schwarz2015].
+Gilbert [-@gilbert1991] proposed that any idea entertained is by default first believed and must later be "unbelieved." However, predominant theoretical accounts attribute illusory truth effects not to belief but rather to metacognitive experiences of fluency or familiarity [@dechene.etal2010].
+Research on metacognition has shown that people's behavior and judgments are not only influenced by the content of their thoughts, but also by the phenomenological experience of processing those thoughts [@alter.oppenheimer2009; @schwarz2015].
 
@@
 A number of findings have shown that processing fluency contributes to the illusory truth effect: Mere prior exposure to statements makes later processing of those statements more fluent, and this metacognitive experience of fluency leads people to rate those statements as more true [@difonzo.etal2016; @fazio.etal2015; @unkelbach2007; @wang.etal2016].
+
+Metacognitive fluency can derive from different sources, and several different types have been shown to influence ratings of truth.
+Fluency can be *perceptual*: statements printed more clearly or in easier-to-read fonts are rated as more true [@ReberSchwarz1999Effects].
+Fluency can be influenced by *memory retrieval*, such that statements that are easier to retreive from memory are rated as more true [@BeggEtAl1992dissociation].
+And fluency can also be semantic or *conceptual* in nature: concepts that are semantically primed are seen as better answers to trivia questions [@KelleyLindsay1993Remembering], and statements evoking previously-activated concepts are rated as more true [@silva.etal2017].
 
@@
 Both theoretical accounts provide a similar overall view of the illusory truth effect: Roughly, when asked whether something is true or false, people search their memory for knowledge pertaining to belief in the statement, but they also rely on their metacognitive experiences of familiarity and fluency.
-Familiarity and fluency are both signs that we have “heard this somewhere before.” According to foundational theories of communication, a general assumption that communicators strive to make true utterances is a prerequisite for successful communication and social functioning [@grice1989].
+Reliance on metacognitive experiences may be, at least to some extent, adaptive: Familiarity and fluency are both signs that we have “heard this somewhere before.” According to foundational theories of communication, a general assumption that communicators strive to make true utterances is a prerequisite for successful communication and social functioning [@grice1989].
 Thus, if having “heard something before” suggests that someone said it, then ecologically this is a reasonable cue to truth.
@@
 
-There is something unsettling about the illusory truth effect and the lack of agency it implies over our own beliefs.
-Even more worrisome, there could be important societal implications if merely being exposed to an idea causes people to adopt it as a belief: Pennycook and colleagues [-@pennycook.etal2018] argue that the illusory truth effect, combined with an environment of pervasive misinformation, has important consequences for the functioning of democratic society.
+There is something unsettling about the illusory truth effect and the lack of agency seems to imply about our own thinking.
+Even more worrisome, there could be important societal implications if merely being exposed to an idea causes people to adopt it as a belief: Pennycook and colleagues [-@pennycook.etal2018] argue that the illusory truth effect, combined with an environment of pervasive misinformation, has important consequences for the functioning of democratic society [also see @UnkelbachEtAl2019Truth].
 The illusory truth effect suggests that mere exposure to misinformation could have impacts that cannot be stopped by fact-checking labels [@pennycook.etal2018], nor effectively curbed by retractions or corrective information [c.f.
@@
 
-Generally, it is these wider consequences of belief (beyond truth ratings for a specific proposition) that should be most concerning.
-If illusory truth effects operate only through metacognitive influences, then these effects may not have much impact outside the lab.
-Under these theoretical accounts, the impact of prior exposure depends on a subsequent re-exposure to a now fluent or familiar statement.
-This would make the real-world consequences of the illusory truth effect far narrower than one might otherwise fear: These metacognitive mechanisms are liable to apply only in response to the exact statements previously seen, but are not as likely to generalize to other related ideas, or to drive actions or decision-making in novel contexts.
-
-Nevertheless, concerns about how the influence of mere exposure might operate in contexts of pervasive misinformation [@difonzo.etal2016; @pennycook.etal2018] loom large enough to warrant further consideration.
-Could mere exposure to statements truly change beliefs?
+The extent to which the illusory truth effect operates entirely through metacognitive processes stands to determine the extent of its real-world consequences.
+Generally, it is the wider consequences of belief (beyond truth ratings for a specific proposition) that should be most concerning.
+If illusory truth effects operate only through metacognitive influences, then the impact of prior exposure depends on a subsequent re-exposure to a now fluent or familiar situation or statement (which might include one activating similar concepts).
+To be sure, this might have important consequences, such as increasing the tendency to share information online after having been exposed previously [@vellani.etal2023].
+Still, the impact of metacognitive effects seem less likely to drive inferences or decision-making in more novel situations.
+Instead, their influences would be constrained to operating within contexts similar to the initial exposure.
 
@@
 A crucial feature of belief is that beliefs are connected with other related beliefs.
-Thus, one way to test whether mere exposure affects beliefs is to examine whether mere exposure to statements can affect truth ratings for other related statements.
+Thus, one way to test whether mere exposure affects beliefs is to examine whether mere exposure to statements can affect truth ratings for other related or implied statements.
 We refer to this hypothetical impact of mere exposure on implied statements as the "Illusory Implication" Effect.
 
-A handful of studies are suggestive in this regard: Unkelbach and Rom [-@unkelbach.rom2017] found that mere exposure to one statement could increase endorsements of implied statements.
-For instance, people were more likely to endorse the statement "Most accidents occur close to weekends" after exposure to the statement "Most accidents occur on Mondays".
-Similarly, Arkes and colleagues [-@arkes.etal1991] found that mere exposure to one set of statements also affected judgments about novel related statements on the same topic (China).
-However, in both cases these findings can also be easily explained by fluency or familiarity engendered by the similarity between the presented and tested statements [@arkes.etal1991; @unkelbach.rom2017].
-
-A case where mere exposure lead to a lower truth rating for a related statement (i.e., one implied to be false) would more clearly implicate an effect of mere exposure on beliefs.
-In our review of the literature, we find three reported studies that offer preliminary evidence of this effect: Bacon [-@bacon1979] found that mere exposure to one statement ($P$) could lead to lower truth ratings for a new, contradictory statements($\neg P$) after a three-week delay.
-For instance, mere exposure to "crocodiles sleep with their eyes open" produced lower truth ratings for "crocodiles sleep with their eyes closed".
-Findings from an experiment by Garcia-Marques and colleagues [-@garcia-marques.etal2015] partially contradicted these results (using a slightly different exposure task), finding lower truth ratings for contradictory statements when participants were queried immediately, but finding the opposite pattern of results at the delay used by the original Bacon [-@bacon1979] study.
-A later series of experiments by Unkelbach and Rom [-@unkelbach.rom2017] generally replicated the newer findings [@garcia-marques.etal2015], although the effects were not observed consistently in each experiment.
+We believe a number of existing findings provide at least preliminary evidence for this illusory implication effect.
+A number of researchers have found that mere exposure affects truth ratings for paraphrases and direct implications or contradictions [@arkes.etal1991; @bacon1979; @garcia-marques.etal2015; @silva.etal2017; @unkelbach.rom2017].
+A crucial question is how distantly inferences are drawn following mere exposure, and whether or not metacognitive accounts can explain these findings.
+
+Silva and colleagues [-@silva.etal2017] found that mere exposure to statements increased truth ratings for paraphrases of those statements that expressed the same idea while sharing only minimal surface-level linguistic features.
+Conversely, exposure to statements *decreased* truth ratings for rephrased contradictions of those statements.
+Silva and colleagues [-@silva.etal2017] explain their findings as resulting from conceptual fluency and disfluency produced by the initial mere exposure to statements: prior exposure "activates" the relevant concepts, easing or interfering with their retrieval and use in processing subsequent statements and thereby producing different metacognitive experiences.
+
+However, findings from Unkelbach and Rom [-@unkelbach.rom2017] cast doubt on a purely fluency-based account of these types of findings.
+These researchers also found that exposure to statements decreased truth judgments for contradictions of those statements.
+Crucially however, they also found that responses were faster for contradictions of previously-seen statements than for novel statements, indicating that processing of these statements was *more* rather than less fluent.
+Thus "conceptual disfluency" is not likely to be an appropriate account of these findings [@unkelbach.rom2017].
+Instead, they and other authors theorize that people tend to believe statements so long as they are generally coherent with respect to existing knowledge, even if they are presented in a context of mere exposure [@garcia-marques.etal2015; @gilbert1991; @unkelbach.rom2017].
 
@@
 
-Here, we sought to expand upon prior experiments to decisively examine whether mere exposure to statements could influence people's beliefs by testing for an "illusory implication" effect.
-To do so, we examined whether people draw inferences from statements after exposure in a context without any presupposition of truth.
-We designed a series of studies to examine whether mere exposure to one statement (the “premise”) could affect truth ratings for a  different statement the premise would logically entail (the “implication”).
-Crucially, we test this across cases where premises entail either the truth or falsehood of another related statement, ruling out purely metacognitive explanations.
-
-In addition, we make several additional contributions toward definitively demonstrating the illusory implication effect beyond existing studies by 1) developing a new set of materials, 2) recruiting larger samples (present studies $N = 1020$, total prior cited studies $N = 452$), 3) utilizing modern hierarchical statistical modeling methods, and 4) taking great pains to ensure that participants' truth ratings are not influenced by task demands (Experiments 2 and 3).
-Across our studies, we find new compelling evidence for this “illusory implication” effect, indicating that mere exposure has genuine impacts on beliefs extending beyond those attributable to metacognitive cues.
+Prior work tested the effect of mere exposure on implied statements using only direct contradictions of the original statement (or paraphrases of direct contradictions).
+Here, we sought to further test for an "illusory implication effect" by testing the effect of mere exposure to statements on their more distant logical implications.
+To do so, we designed a series of studies to examine whether mere exposure to one statement (the “premise”) could affect truth ratings for another statement this premise would logically entail to be either true or false (the “implication”).
+
+We developed materials involving more distant inferences that do not afford direct one-to-one mappings between shared concepts in the premise and implication statements.
+For instance, one statement pair was the premise statement, “No U.S.
+astronauts have died since the Challenger explosion in 1986” paired with the implied statement, “The space shuttle Columbia disintegrated over Texas in 2003” (true, but implied to be false by the "premise").
+In contrast with prior materials using "contradictory" statements, there is no direct one-to-one mapping of concepts between the two statements.
+Instead, a chain of inferences must developed for a conclusion to be drawn: that the space shuttle Columbia contained astronauts, that its disintegration would have killed them, and that 2003 is later than 1986.
+
+In addition, we make several additional contributions toward definitively demonstrating the illusory implication effect beyond existing studies by 1) testing the effect with large samples of participants (present studies $N = 1020$), 2) utilizing modern hierarchical statistical modeling methods, and 3) taking great pains to ensure that participants' truth ratings are not influenced by task demands (Experiments 2 and 3).
+Across our studies, we find further compelling evidence for this “illusory implication” effect, indicating that mere exposure has genuine impacts on beliefs extending beyond those attributable to metacognitive cues.
 
@@
 
-In Experiments 1 and 2, the effect of familiarity was weaker than the measured implication effect, so that average truth ratings were higher after exposure for false-implied-true items and lower for true-implied-false items (Figures 1 and 2).
-In Experiment 3 however, the estimated effect of familiarity and implication were similar, and truth ratings were similar for exposed and unexposed true-implied-false items (Figure 3).
-Although the estimate for the implication effect in our statistical model was clearly greater than zero, this may nevertheless allow a possible alternate explanation of our findings: that the true-implied-false items' premise statements simply do not engender as much familiarity or fluency for their implication as do the false-implied-true items, so that the effects differ between these groups.
+In Experiments 1 and 2, the effect of exposure was weaker than the measured implication effect, so that average truth ratings were higher after exposure for false-implied-true items and lower for true-implied-false items (Figures 1 and 2).
+In Experiment 3 however, the estimated effect of exposure and implication were similar, and truth ratings were similar for exposed and unexposed true-implied-false items (Figure 3).
+Although the estimate for the implication effect in our statistical model was clearly greater than zero, this may nevertheless allow a possible alternate explanation of our findings: that the true-implied-false items' premise statements simply do not engender as much familiarity or fluency for their implications as do the false-implied-true items, so that the effects differ between these groups.
 Put another way, the identification of the "implication" effect coefficient rests on the assumption that familiarity and fluency effects are not accidentally confounded with item groups.
@@
 One potential explanation is that mere exposure to the premise statements leads participants to at least partially adopt them as beliefs.
-Simply entertaining the idea may create such an impression despite awareness of its ambiguous truth-value [@gilbert.etal1990; @gilbert1991; @garcia-marques.etal2015].
+Simply entertaining the idea may create such an impression despite awareness of its ambiguous truth-value [@gilbert.etal1990; @gilbert1991; @garcia-marques.etal2015; @unkelbach.rom2017].
 If so, this apparently occurs despite participant's awareness that the statements may be false, as ensured by constant reminders and comprehension checks in Experiment 3.
 
-Alternately, the illusory implication effect may reflect explicit memory and source misattribution during testing: Participants may have some explicit memory for the premise statements, but fail to properly attribute their source to the experimental context.
-Future research might explore how explicit memory for the premise statements correlates with the strength of the illusory implication effects.
+Modern theorists do not conceive of beliefs as lists of propositiosn, but instead as subserved by structured mental models or intuitive theories of how the world works [e.g.
+@gelman.legare2011; @gerstenberg.tenenbaum2017; @powell.etal2023; @johnson-laird2001; @thagard1989].
+Thus, coming to believe a statement entails updating one's mental models with the new information conveyed by that statement.
+A key feature of these mental models is their coherence: updating them is expected to have wide-ranging and coherent implications [@gerstenberg.tenenbaum2017; @powell.etal2023; @thagard1989].
+In the context of mere exposure effects, Unkelbach and Rom [-@unkelbach.rom2017] have proposed an account based on parallel constraint satisfaction within semantic networks, which might plausibly be elaborated to account for the present findings.
+In any case, we take the coherence of inferences following exposure to be evidence for a meaningful influence on people's mental models, and the key to our claim that mere exposure influences beliefs.
+
+Finally, it is worth clarifying that our findings and wider claims about the illusory implication effect are broadly consistent with existing metacognitive explanations of the illusory truth effect: we hypothesize that the illusory implication effect is a distinct and somewhat orthogonal effect, based in appreciation of the meaning of statements rather than solely in the metacognitive experiences they produce.
+Nevertheless, it remains clear that fluency plays a crucial role in the classic illusory truth effect, especially as evidenced by neuroscientific findings [@wang.etal2016] and by the findings indicating the effect can be reversed when the pragmatics of the exposure context are changed [@corneille.etal2020].
+
+## Conclusions
 
@@
 The spread of misinformation and fake news appears to outpace the spread of factual news [@vosoughi.etal2018] and creates massive challenges for online platforms seeking to remove, label, and correct misinformation [@sharma.etal2019].
-These challenges may be further compounded by basic features of human cognition  that our studies have revealed: mere exposure to false statements can lead them to be _believed_.
-
-## Constraints on Generality
-
-The studies reported here targeted U.S.
-adults through an online survey provider (CloudResearch).
-Study materials were designed accordingly and may produce different results if tested on populations with different backgrounds and prior knowledge.
+These challenges may be further compounded by apparently basic features of human cognition: that mere exposure to false statements can lead them to be _believed_.
 

### 2024-04-25 paper/paper-papaja.Rmd
revised methods to contradictory and entailed statements

@@
 item_fluency_fig <- statements %>%
-  mutate(B_implied_true = if_else(B_implied_true==1, "F implied T", "T implied F")) %>% 
+  mutate(
+    B_implied_true = if_else(B_implied_true==1, "F implied T", "T implied F"),
+    B_implied_true = if_else(B_implied_true == "T implied F", "Contradicted", "Entailed") 
+    ) %>% 
   ggplot(aes(x = factor(B_implied_true), y = overlap_prop)) +
@@
     Term = case_when(
-      Term == "B_implied_trueFimpliedT" ~ "implication type",
+      Term == "B_implied_trueFimpliedT" ~ "Statement Truth",
+      Term == "familiarity" ~ "Exposure",
+      Term == "implication" ~ "Implication",
       # grepl("Intercept\\[", Term) ~ gsub("Intercept\\[(.*)\\]","$\\\\alpha_{\\1}$", Term),
@@
     Term = case_when(
-      Term == "B_implied_trueFimpliedT" ~ "implication type",
+      Term == "B_implied_trueFimpliedT" ~ "Statement Truth",
+      grepl("familiarity", Term) ~ "Exposure",
+      Term == "implication" ~ "Implication",
       # grepl("Intercept\\[", Term) ~ gsub("Intercept\\[(.*)\\]","$\\\\alpha_{\\1}$", Term),
-      grepl("familiarity", Term) ~ "familiarity",
       # Term == "item_sd(Intercept)" ~ "$\\sigma_{item}$",
@@
     Term = case_when(
-      Term == "B_implied_trueFimpliedT" ~ "implication type",
+      Term == "B_implied_trueFimpliedT" ~ "Statement Truth",
+      Term == "familiarity" ~ "Exposure",
+      Term == "implication" ~ "Implication",
       # grepl("Intercept\\[", Term) ~ gsub("Intercept\\[(.*)\\]","$\\\\alpha_{\\1}$", Term),
-      grepl("familiarity", Term) ~ "familiarity",
+      grepl("familiarity", Term) ~ "Exposure",
       # Term == "item_sd(Intercept)" ~ "$\\sigma_{item}$",
@@
 We developed materials involving more distant inferences that do not afford direct one-to-one mappings between shared concepts in the premise and implication statements.
+Thus, any reduction in truth ratings for the contradicted statements following exposure to the premise statements are not likely to be attributable to disfluency [cf.
+@silva.etal2017].
 For instance, one statement pair was the premise statement, “No U.S.
-astronauts have died since the Challenger explosion in 1986” paired with the implied statement, “The space shuttle Columbia disintegrated over Texas in 2003” (true, but implied to be false by the "premise").
-In contrast with prior materials using "contradictory" statements, there is no direct one-to-one mapping of concepts between the two statements.
+astronauts have died since the Challenger explosion in 1986” paired with the implied statement, “The space shuttle Columbia disintegrated over Texas in 2003” (true, but contradicted by the "premise").
+In contrast with prior work examining conflicting statements, there is no direct one-to-one mapping of concepts between the two statements.
 Instead, a chain of inferences must developed for a conclusion to be drawn: that the space shuttle Columbia contained astronauts, that its disintegration would have killed them, and that 2003 is later than 1986.
@@
     exposed = if_else(exposed==1, "Exposed", "Unexposed"),
-    B_implied_true = as.character(B_implied_true)
+    B_implied_true = if_else( B_implied_true == "T implied F", "Contradicted", "Entailed") #as.character(B_implied_true)
     ) %>% 
@@
 We created 24 pairs of statements presenting claims about history.
-Each pair consisted of a "premise" statement and a "implied" statement.
-Each “premise” statement was a falsehood that implied either the truth or falsity of the “implied” statement.
-Of the pairs, 12 had premise statements implying that a true statement was false (true-implied-false) and the other 12 had premise statements implying another false statement was true (false-implied-true).
-For instance, one "true-implied-false" pair was the premise statement, “No U.S.
-astronauts have died since the Challenger explosion in 1986” paired with the implied statement, “The space shuttle Columbia disintegrated over Texas in 2003” (true, but implied to be false by the "premise").
-An example of a "false-implied-true" pair was the premise statement, "The Tour de France has been held every year since its inception in 1903" paired with the implied statement, "The Tour de France was still held during WWI and WWII" (false, but implied true by the "premise").
+Each pair consisted of a "premise" statement and a "implication" statement.
+Each “premise” statement was a falsehood that either entailed or contradicted the “implication” statement.
+Of the pairs, 12 had premise statements implying that a true statement was false ("contradicted") and the other 12 had premise statements implying another false statement was true ("entailed").
+For instance, one "contradicted" pair was the premise statement, “No U.S.
+astronauts have died since the Challenger explosion in 1986” paired with the implied statement, “The space shuttle Columbia disintegrated over Texas in 2003” (true, but contradicted by the "premise").
+An example of an "entailed" pair was the premise statement, "The Tour de France has been held every year since its inception in 1903" paired with the implied statement, "The Tour de France was still held during WWI and WWII" (false, but entailed by the "premise").
 
@@
 Participants were quizzed on the 8 previously-exposed premise statements, as well as their corresponding "implied" statements.
-For comparison, they were also quizzed on 8 new premise statements and 8 unrelated implied statements.
+For comparison within subjects, they were also quizzed on 8 new premise statements and 8 unrelated implication statements.
 Participants were randomly assigned into three counterbalancing conditions that varied which of the statement-pairs were assigned to each exposure/test combination.
@@
 Similarly, we tested for the illusory implication effect by comparing truth ratings for implication statements whose corresponding premise statements were and were not presented during the exposure phase.
-If mere exposure to "premise" statements affects participants' judgments of the "implied" statements, this would be evidence for an illusory implication effect.
+If mere exposure to "premise" statements affects participants' judgments of the "implication" statements, this would be evidence for an illusory implication effect.
 
@@
 Figure 1 shows participants’ average truth ratings for the implication statements in the quiz condition.
-Participants gave higher truth ratings following exposure for false statements implied to be true, and somewhat lower truth ratings following exposure for true statements implied to be false.
+Participants gave higher truth ratings following exposure for entailed statements, and somewhat lower truth ratings following exposure for contradicted statements.
 
@@
 
-An increase in truth ratings for the false-implied-true statements could potentially be explained by increased familiarity.
+An increase in truth ratings for the entailed statements could potentially be explained by familiarity or fluency.
 This would be consistent with findings from Arkes and colleagues [-@arkes.etal1991], who found evidence for illusory truth for novel statements that were on the same topic as previously-seen statements.
-However, familiarity cannot explain the decrease in truth ratings for the true-implied-false items, as familiarity should encourage uniformly higher, not lower, truth ratings.
-
-A regression model was used to tease apart the effects of familiarity and logical implication.
-The model includes a) a binary variable indicating the type of implication statement (true-implied-false or false-implied-true), b) a binary predictor for prior exposure to the related premise coded zero when the premise was not seen and 1 when it was seen (capturing potential effects of familiarity), c) and a variable capturing the effect of implications, coded 1 for implied truth, -1 for implied falsehood, and zero when the related premise statement was not seen at exposure.
-Thus the "implication" predictor accounts for the effect of exposure to the premise on truth judgments for the implied statements (either positive or negative), while the "familiarity" predictor accounts for any positive effect of familiarity.
+However, simple metacognitive cues cannot explain the decrease in truth ratings for the contradicted items, as familiarity or fluency should encourage uniformly higher, not lower, truth ratings.
+
+A regression model was used to tease apart the possible metacognitive effects of mere exposure of and logical implication.
+The model includes a) a binary variable indicating the type of implication statement's actual truth value (true or false), b) a binary predictor for prior exposure to the related premise coded zero when the premise was not seen and 1 when it was seen (capturing potential uni-directional metacognitive effects of exposure), c) and a variable capturing the effect of implications, coded 1 for implied truth, -1 for implied falsehood, and zero when the related premise statement was not seen at exposure.
+Thus the "implication" predictor accounts for the effect of exposure to the premise on truth judgments for the implied statements (either positive or negative), while the "exposure" predictor accounts for any positive effect from metacognitive cues.
 We also incorporated "maximal" random intercepts and slopes for all terms varying by subject and item [@barr.etal2013].
@@
 \text{response} \sim 
- \text{item type} &+ \text{familiarity} + \text{implication} \\
-&+ (1 + \text{familiarity} + \text{implication}|\text{subject}) \\
-&+ (1 + \text{familiarity} + \text{implication}|\text{item})
+ \text{statement truth} &+ \text{exposure} + \text{implication} \\
+&+ (1 + \text{exposure} + \text{implication}|\text{subject}) \\
+&+ (1 + \text{exposure} + \text{implication}|\text{item})
 \end{align*}
@@
   mutate(
-    exposed = if_else(exposed, "Exposed", "Unexposed")
+    exposed = if_else(exposed, "Exposed", "Unexposed"),
+    B_implied_true = if_else(B_implied_true == "T implied F", "Contradicted", "Entailed") 
   ) %>% 
@@
 At exposure, participants were told that they would see a set of statements, some true and some false, and they were instructed to rate how interesting each statement was.
-This change was made to avoid forcing participants to make a true/false judgment for the exposed premise-statements, which could have pushed them to try to make coherent or consistent responses to the related implication-statements.
+This change was made to avoid forcing participants to make a true/false judgment for the exposed premise statements, which could have pushed them to try to make coherent or consistent responses to the related implication statements.
 Importantly, as with a true/false quiz, the presentation of statements in this context should not warrant any inference as to the truth of those statements.
@@
 The illusory implication effect was again observed.
-As shown in the figure, truth ratings were clearly affected by exposure: participants gave higher truth ratings following exposure for false statements implied to be true, and lower truth ratings following exposure for true statements implied to be false.
+As shown in the figure, truth ratings were clearly affected by exposure: participants gave higher truth ratings following exposure for entailed statements, and lower truth ratings following exposure for contradicted statements.
 Table 2 shows the posterior population-level estimates for an identical regression as was conducted for Experiment 1.
@@
   ) %>% 
-  mutate(exposed = if_else(exposed, "Exposed", "Unexposed"))
+  mutate(exposed = if_else(exposed, "Exposed", "Unexposed")) %>% 
+  mutate(B_implied_true = if_else(B_implied_true == "T implied F", "Contradicted", "Entailed") )
 
@@
   ) %>% 
-  mutate(B_implied_true = ordered(B_implied_true, levels = c("F implied T", "T implied F"))) %>% 
+  # mutate(B_implied_true = ordered(B_implied_true, levels = c("F implied T", "T implied F"))) %>% 
   ggplot(
@@
 
-In Experiments 1 and 2, the effect of exposure was weaker than the measured implication effect, so that average truth ratings were higher after exposure for false-implied-true items and lower for true-implied-false items (Figures 1 and 2).
-In Experiment 3 however, the estimated effect of exposure and implication were similar, and truth ratings were similar for exposed and unexposed true-implied-false items (Figure 3).
-Although the estimate for the implication effect in our statistical model was clearly greater than zero, this may nevertheless allow a possible alternate explanation of our findings: that the true-implied-false items' premise statements simply do not engender as much familiarity or fluency for their implications as do the false-implied-true items, so that the effects differ between these groups.
+In Experiments 1 and 2, the effect of exposure was weaker than the measured implication effect, so that average truth ratings were higher after exposure for entailed items and lower for contradicted items (Figures 1 and 2).
+In Experiment 3 however, the estimated effect of exposure and implication were similar, and truth ratings were similar for exposed and unexposed contradicted items (Figure 3).
+Although the estimate for the implication effect in our statistical model was clearly greater than zero, this may nevertheless allow a possible alternate explanation of our findings: that the contradicted items' premise statements simply do not engender as much familiarity or fluency for their implications as do the entailed items, so that the effects differ between these groups.
 Put another way, the identification of the "implication" effect coefficient rests on the assumption that familiarity and fluency effects are not accidentally confounded with item groups.

### 2024-04-26 paper/paper-papaja.Rmd
updated wordcount
typo fixes

@@
 However, genuine beliefs do not only affect truth judgments about individual statements, they also imply other beliefs and drive decision-making.
-Here, we consider whether exposure to “premise” statements affects people's truth ratings for novel “implied” statements, a pattern of findings a pattern of findings we call the "illusory implication" effect.
+Here, we consider whether exposure to “premise” statements affects people's truth ratings for novel “implied” statements, a pattern of findings we call the "illusory implication" effect.
 We argue these effects would constitute evidence for genuine belief change from mere exposure and  identify a handful of existing findings that offer preliminary support for this claim.
@@
 keywords          : "Illusory truth; Metacognition; Cognitive Psychology; Misinformation"
-wordcount         : "5636"
+wordcount         : "5660"
 
@@
 In contrast with prior work examining conflicting statements, there is no direct one-to-one mapping of concepts between the two statements.
-Instead, a chain of inferences must developed for a conclusion to be drawn: that the space shuttle Columbia contained astronauts, that its disintegration would have killed them, and that 2003 is later than 1986.
+Instead, a chain of inferences must be developed for a conclusion to be drawn: that the space shuttle Columbia contained astronauts, that its disintegration would have killed them, and that 2003 is later than 1986.
 

### 2024-06-28 paper/paper-papaja-revising.Rmd
revisions 6-24

--- 2024-04-26 paper/paper-papaja.Rmd
@@
 ---
-title             : "Illusory Implications: Mere exposure to ideas can induce beliefs"
+title             : "Illusory Implications: Incidental exposure to ideas can induce beliefs"
 shorttitle        : "Illusory implications"
@@
 abstract: |
-  Numerous psychological findings have shown that mere exposure to ideas makes those ideas seem more true, a finding commonly referred to as the “illusory truth” effect.
+  Numerous psychological findings have shown that incidental exposure to ideas makes those ideas seem more true, a finding commonly referred to as the “illusory truth” effect.
 Under predominant accounts of the illusory truth effect, initial exposure to a statement provides a metacognitive feeling of “fluency” or familiarity that, upon subsequent exposure, leads people to infer the statement is more likely to be true.
@@
 Here, we consider whether exposure to “premise” statements affects people's truth ratings for novel “implied” statements, a pattern of findings we call the "illusory implication" effect.
-We argue these effects would constitute evidence for genuine belief change from mere exposure and  identify a handful of existing findings that offer preliminary support for this claim.
+We argue these effects would constitute evidence for genuine belief change from incidental exposure and  identify a handful of existing findings that offer preliminary support for this claim.
 Building upon these, we conduct three new pre-registered experiments to further test this hypothesis, finding additional evidence that exposure to “premise” statements affected participants’ truth ratings for novel “implied” statements, including for considerably more distant implications than those previously explored.
-Our findings suggest that the effects of mere exposure reach further than previously thought, with potentially consequential implications for concerns around mis- and disinformation.
+Our findings suggest that the effects of incidental exposure reach further than previously thought, with potentially consequential implications for concerns around mis- and disinformation.
 
@@
 library(textstem)
+
+source("fig-code.R")
 
@@
   ) %>% 
-  make_table("Population coefficients of Bayesian regression model for Implication effects in Experiment 1.")
+  make_table("Population coefficients and 95\\% credible intervals of Bayesian regression model for Implication effects in Experiment 1.")
 
@@
   ) %>% 
-  make_table("Population coefficients of Bayesian regression model for Implication effects in Experiment 2.")
+  make_table("Population coefficients and 95\\% credible intervals of Bayesian regression model for Implication effects in Experiment 2.")
 
@@
   ) %>% 
-  make_table("Population coefficients of Bayesian regression model for Implication effects in Experiment 3.")
+  make_table("Population coefficients and 95\\% credible intervals of Bayesian regression model for Implication effects in Experiment 3.")
 
@@
 Perhaps not.
-Numerous psychological findings indicate that mere exposure to ideas makes those ideas appear more true, a finding commonly referred to as the “illusory truth” effect [@dekeersmaecker.etal2020; @dechene.etal2010; @fazio.etal2015; @hasher.etal1977; @pennycook.etal2018].
+Numerous psychological findings indicate that incidental exposure to ideas makes those ideas appear more true, a finding commonly referred to as the “illusory truth” effect [@dekeersmaecker.etal2020; @dechene.etal2010; @fazio.etal2015; @hasher.etal1977; @pennycook.etal2018].
 This effect suggests that we cannot exist in an environment of mis- and disinformation without being affected by it---without exposure to these ideas distorting our sense of what is true.
@@
 
-The effect is one of "illusory" truth because it occurs following "mere exposure" to statements.
+The effect is one of "illusory" truth because it occurs following "incidental exposure" to statements.
 That is, it occurs when the statements are seen or heard in a non-communicative context, such as when they are read during a true/false quiz.
@@
 
-Why should mere exposure affect later ratings of truth?
+Why should incidental exposure affect later ratings of truth?
 Gilbert [-@gilbert1991] proposed that any idea entertained is by default first believed and must later be "unbelieved." However, predominant theoretical accounts attribute illusory truth effects not to belief but rather to metacognitive experiences of fluency or familiarity [@dechene.etal2010].
@@
 When thinking is easy and fluent, people tend to judge the targets of that thinking more favorably than if thinking is challenging and disfluent [@wanke.etal1997].
-A number of findings have shown that processing fluency contributes to the illusory truth effect: Mere prior exposure to statements makes later processing of those statements more fluent, and this metacognitive experience of fluency leads people to rate those statements as more true [@difonzo.etal2016; @fazio.etal2015; @unkelbach2007; @wang.etal2016].
+A number of findings have shown that processing fluency contributes to the illusory truth effect: Incidental prior exposure to statements makes later processing of those statements more fluent, and this metacognitive experience of fluency leads people to rate those statements as more true [@difonzo.etal2016; @fazio.etal2015; @unkelbach2007; @wang.etal2016].
 
@@
 
-## Mere Exposure and Belief
+## Incidental Exposure and Belief
 
@@
 Even more worrisome, there could be important societal implications if merely being exposed to an idea causes people to adopt it as a belief: Pennycook and colleagues [-@pennycook.etal2018] argue that the illusory truth effect, combined with an environment of pervasive misinformation, has important consequences for the functioning of democratic society [also see @UnkelbachEtAl2019Truth].
-The illusory truth effect suggests that mere exposure to misinformation could have impacts that cannot be stopped by fact-checking labels [@pennycook.etal2018], nor effectively curbed by retractions or corrective information [c.f.
+The illusory truth effect suggests that incidental exposure to misinformation could have impacts that cannot be stopped by fact-checking labels [@pennycook.etal2018], nor effectively curbed by retractions or corrective information [c.f.
 @ecker.etal2011; @lewandowsky.etal2012].
@@
 However, there are more to _beliefs_ than ratings of truth.
-Though a person’s assent to a proposition (whether they agree with or judge it to be truthful) is an important marker and convenient measure of belief, other features of belief are just as essential.
+Though a person’s assent to a proposition (whether they agree with or judge statements expressing it to be truthful) is an important marker and convenient measure of belief, other features of belief are just as essential.
 For one, beliefs imply other beliefs: for instance, believing that “it is sunny out” implies the belief that “it is daytime”.
@@
 
-The extent to which the illusory truth effect operates entirely through metacognitive processes stands to determine the extent of its real-world consequences.
-Generally, it is the wider consequences of belief (beyond truth ratings for a specific proposition) that should be most concerning.
-If illusory truth effects operate only through metacognitive influences, then the impact of prior exposure depends on a subsequent re-exposure to a now fluent or familiar situation or statement (which might include one activating similar concepts).
-To be sure, this might have important consequences, such as increasing the tendency to share information online after having been exposed previously [@vellani.etal2023].
+<!-- It is important to distinguish between statements and the propositions or ideas they express.
+Two different statements might express the same proposition: the pair of statements "I went to the store the day before yesterday" and "I went to the store two days ago" is one example.
+When incidental exposure produces metacognitive experiences like fluency, this might lead to increased endorsement of a specific statement, without truly affecting a person's attitude toward the proposition it represents.
+Instead, the effects of prior exposure would depend on features of subsequently-seen statements (e.g.
+lexical similarity: sharing words; or conceptual similarity: sharing concepts), rather than the propositions they express.
+-->
+
+```{r theory-fig, echo = FALSE, message=FALSE, fig.align='center', fig.cap='Diagram depicting the theorized causal relations between latent beliefs, metacogntive experiences, and truth ratings for different types of statements.
+We highlight the paths from observable behavior which can only be traced back to exposure through beliefs in red.', out.width='0.75\\linewidth', fig.pos='h'}
+knitr::include_graphics("theory-fig.pdf")
+```
+
+Figure \@ref(fig:theory-fig) depicts our hypotheses concerning the causal relationships between latent beliefs, metacognitive experiences, and overt behaviors such truth ratings and actions.
+As shown in the figure, experiences or manipulations that affect metacognitive experiences, such as incidental exposure, might produce increased truth ratings for statements expressing $P$ (either through lexical or conceptual similarity).
+However, beliefs should be expected to produce additional consequences.
+Here, we focus on inferences: a manipulation affecting belief in $P$ should increase truth ratings for statements expressing some proposition $Q$ where $P$ implies $Q$, and decrease belief in $R$ where $P$ implies $\neg R$.
+Figure \@ref(fig:theory-fig) also depicts a possible relation between metacognitive experiences and statements expressing $Q$.
+Where a statement $S_Q$ expressing $Q$ invokes a similar set of concepts as $S_P$, conceptual fluency might produce increased truth ratings.
+
+Crucially, our theory draws no connection between metacognitive experiences and reduced truth judgments for statements expressing $R$ where $P$ implies $\neg R$.
+Such a connection would require a cogent theory of how conceptual "disfluency" might be engendered to reduce truth judgments without the process of inference that would implicate changes in belief.
+Current accounts have not developed such a theory [e.g.
+c.f.
+@AlterOppenheimer2009Uniting; @VogelEtAl2020Truth], and it appears unlikely to be workable: people frequently encounter chains of statements invoking different concepts, yet they do so seemingly without issue, as considering one topic does not necessarily make it more difficult to consider another unrelated topic.
+
+We argue that where effects of incidental exposure on truth judgments are attributable to metacognitive processes, these effects do not generally constitute evidence for a change in beliefs.
+Applying the theoretical framework sketched in Figure \@ref(fig:theory-fig), cases where incidental exposure to a proposition $P$ increases truth ratings for statements $S_P$ expressing that proposition do not generally constitute evidence that incidental exposure has affected beliefs, since there is a causal pathway between exposure and this result that does not pass through belief.
+Likewise, evidence of exposure to $S_P$ increasing truth ratings for $S_Q$ where $S_Q$ shares a similar set of concepts with $S_P$ do not constitute evidence for effects on beliefs.
+However, evidence that exposure decreases truth ratings for $S_R$ or increases truth ratings for $S'_Q$ where $S'_Q$ does not directly share concepts with $S_P$ would constitute evidence for effects on beliefs.
+
+Practically speaking, the extent to which effects of incidental exposure operate entirely through metacognitive processes stands to determine the extent of their real-world consequences.
+Generally, it is the wider consequences of belief (beyond truth ratings for a specific statement) that should be most concerning.
+If effects of incidental exposure operate only through metacognitive influences, then the impact of prior exposure depends on a subsequent re-exposure to a now fluent or familiar situation or statement (which might include one activating similar concepts).
+To be sure, metacognitive effects might have important consequences even if they do not affect beliefs, such as increasing the tendency to share information online after having been exposed previously [@vellani.etal2023].
 Still, the impact of metacognitive effects seem less likely to drive inferences or decision-making in more novel situations.
@@
 A crucial feature of belief is that beliefs are connected with other related beliefs.
-Thus, one way to test whether mere exposure affects beliefs is to examine whether mere exposure to statements can affect truth ratings for other related or implied statements.
-We refer to this hypothetical impact of mere exposure on implied statements as the "Illusory Implication" Effect.
-
-We believe a number of existing findings provide at least preliminary evidence for this illusory implication effect.
-A number of researchers have found that mere exposure affects truth ratings for paraphrases and direct implications or contradictions [@arkes.etal1991; @bacon1979; @garcia-marques.etal2015; @silva.etal2017; @unkelbach.rom2017].
-A crucial question is how distantly inferences are drawn following mere exposure, and whether or not metacognitive accounts can explain these findings.
-
-Silva and colleagues [-@silva.etal2017] found that mere exposure to statements increased truth ratings for paraphrases of those statements that expressed the same idea while sharing only minimal surface-level linguistic features.
-Conversely, exposure to statements *decreased* truth ratings for rephrased contradictions of those statements.
-Silva and colleagues [-@silva.etal2017] explain their findings as resulting from conceptual fluency and disfluency produced by the initial mere exposure to statements: prior exposure "activates" the relevant concepts, easing or interfering with their retrieval and use in processing subsequent statements and thereby producing different metacognitive experiences.
-
-However, findings from Unkelbach and Rom [-@unkelbach.rom2017] cast doubt on a purely fluency-based account of these types of findings.
-These researchers also found that exposure to statements decreased truth judgments for contradictions of those statements.
-Crucially however, they also found that responses were faster for contradictions of previously-seen statements than for novel statements, indicating that processing of these statements was *more* rather than less fluent.
-Thus "conceptual disfluency" is not likely to be an appropriate account of these findings [@unkelbach.rom2017].
-Instead, they and other authors theorize that people tend to believe statements so long as they are generally coherent with respect to existing knowledge, even if they are presented in a context of mere exposure [@garcia-marques.etal2015; @gilbert1991; @unkelbach.rom2017].
+Thus, one way to test whether incidental exposure affects beliefs is to examine whether incidental exposure to statements can affect truth ratings for other related or implied statements in ways that cannot be attributed to metacognitive effects.
+We refer to this hypothetical impact of incidental exposure on implied statements as the "Illusory Implication" Effect.
+
+A key test of the illusory implication effect as opposed to other illusory truth effects is whether prior incidental exposure can _reduce_ judgments of truth.
+The experience of fluency is inherently relative: a metacognitive experience is only fluent or disfluent with respect to some point of comparison.
+For instance, utterances can be made relatively fluent by rhyming (McGlone & Tofighbakhsh, 2000) or they can be made relatively disfluent with the addition of speech markers like "um" and "uh" Brennan & Williams (1995).
+However, in the case of repetition-induced fluency (whether conceptual, lexical, or otherwise), repetition can produce fluency, but it is less clear how a such a manipulation might produce disfluency.
+Thus, it is difficult to see how effects of incidental exposure that reduce truth ratings can be explained by metacognitive factors.
+
+A handful of existing findings provide some preliminary evidence for our hypothesized illusory implication effect.
+A number of researchers have found that incidental exposure affects truth ratings for novel statements contradicting previously exposed statements [@garcia-marques.etal2015; @silva.etal2017; @unkelbach.rom2017].
+To our knowledge, three experiments have found evidence for this pattern of influence: Garcia-Marques (n = 58) [@garcia-marques.etal2015], Silva et al.
+[-@silva.etal2017] (Exp.
+2, n = 45), and Unkelbach and Rom [-@unkelbach.rom2017] (Exp.
+3; self-referent condition; n = 27).
+Such a pattern is not easily attributable to purely metacogntiive processes.
+Instead, several of these researchers [e.g., @garcia-marques.etal2015; @silva.etal2017] attribute these lowered truth ratings to the "meaning of the statements" rather than to conceptual fluency.
+
+Garcia-marques and colleagues [-@garcia-marques.etal2015] induced incidental exposure by asking participants to rate how interesting statements were, warning them that half were true and half false.
+After this initial task inducing incidental exposure, a subsequent quiz featured a mix of novel statements, repeated statements, and statements contradicting those previously seen either immediately or after a 1-week delay.
+In the immediate-test condition, contradictory statements were rated as less true than novel statements.[^bacon]
+
+[^bacon]: These researchers used materials following those used by Bacon [-@bacon1979, Exp.
+2], who induced incidental exposure through a series of true-or-false quizzes.
+In this study, contradictory statements were rated as less true following a two-week delay, but only when participants were warned that the statements had been changed.
+This warning provides an avenue by which metacognitive cues could be used to inform truth judgments about contradictions without inducing beliefs.
+
+Silva and colleagues [-@silva.etal2017] found that exposure to statements increased truth ratings for paraphrases of those statements that expressed the same proposition while sharing only minimal surface-level linguistic features.
+These researchers explained their findings as resulting from conceptual fluency produced by the initial exposure to statements: prior exposure "activates" the relevant concepts, easing with their retrieval and use in processing subsequent statements and thereby producing different metacognitive experiences.
+Silva and colleagues [-@silva.etal2017] also found that exposure to statements *decreased* truth ratings for rephrased contradictions of those statements.
+However, these findings must be qualified by the exposure context used here: Silva and colleagues asked participants to rate their interest in statements, but did not warn them that the statements could be true or false.
+Thus the pragmatics of the task are unclear, and participants might have believed these statements because they understood them to be presented as facts.
+
+Unkelbach and Rom [-@unkelbach.rom2017] examined both implied and contradictory statements.
+They found that exposure to implied statements increased truth ratings (Exp 4.) and that statements saw a significant decrease in truth ratings following incidental exposure to their contradictions in a "self-relevant" processing condition (Exp.
+4).
+Crucially, the results of this experiment appear to preclude any purely fluency-based account of reduced truth ratings following incidental exposure to contradictory statements: These researchers found that responses were faster for contradictions of previously-seen statements than for novel statements, indicating that processing of these statements was *more* rather than less fluent.
+Thus even if an appropriate notion of "conceptual disfluency" could be developed, it is unlikely it could be an appropriate account of these findings [@unkelbach.rom2017].
+Instead, they and other authors theorize that people tend to believe statements so long as they are generally coherent with respect to existing knowledge, even if they are presented in a context of incidental exposure [@garcia-marques.etal2015; @gilbert1991; @silva.etal2017; @unkelbach.rom2017].
+
+The presently-hypothesized illusory inference effect is a key theoretical link demonstrating the connection between incidental exposure and _beliefs_, and  all of the attendant downstream consequences associated with beliefs.
+A handful of prior findings suggest the existence of such an effect, yet this existing evidence seems far from dispositive, being based on only a handful of small studies with sometimes inconsistent methods.
 
@@
 
-Prior work tested the effect of mere exposure on implied statements using only direct contradictions of the original statement (or paraphrases of direct contradictions).
-Here, we sought to further test for an "illusory implication effect" by testing the effect of mere exposure to statements on their more distant logical implications.
-To do so, we designed a series of studies to examine whether mere exposure to one statement (the “premise”) could affect truth ratings for another statement this premise would logically entail to be either true or false (the “implication”).
-
-We developed materials involving more distant inferences that do not afford direct one-to-one mappings between shared concepts in the premise and implication statements.
-Thus, any reduction in truth ratings for the contradicted statements following exposure to the premise statements are not likely to be attributable to disfluency [cf.
-@silva.etal2017].
+Prior work tested the effect of incidental exposure on implied statements using only direct contradictions of the original statement (or paraphrases of direct contradictions).
+Here, we sought to further test for an "illusory implication effect" by testing the effect of incidental exposure to statements on their more distant logical implications.
+To do so, we designed a series of studies to examine whether incidental exposure to one statement (the “premise”) could affect truth ratings for another statement this premise would logically entail to be either true or false (the “implication”).
+
+In our studies, we sought to eliminate as much as possible the role of perceptual and conceptual fluency in affecting truth judgments.
+These efforts are aimed at demonstrating as clearly as possible a causal pathway  between incidental exposure and effects on truth ratings through changes in belief.
+That is, establishing a pathway from exposure to increased truth ratings for implications $Q$ and decreased truth ratings for contradictions $R$.
+Compared with prior work, we developed materials involving more distant inferences that do not afford direct one-to-one mappings between shared concepts in the premise and implication statements.
+Thus, any reduction in truth ratings for the contradicted statements following exposure to the premise statements are not reasonably attributable to disfluency (closing the alternate pathway depicted in Figure X).
 For instance, one statement pair was the premise statement, “No U.S.
@@
 
-In addition, we make several additional contributions toward definitively demonstrating the illusory implication effect beyond existing studies by 1) testing the effect with large samples of participants (present studies $N = 1020$), 2) utilizing modern hierarchical statistical modeling methods, and 3) taking great pains to ensure that participants' truth ratings are not influenced by task demands (Experiments 2 and 3).
-Across our studies, we find further compelling evidence for this “illusory implication” effect, indicating that mere exposure has genuine impacts on beliefs extending beyond those attributable to metacognitive cues.
+Additionally, we make several further contributions toward definitively demonstrating the illusory implication effect beyond existing studies by 1) testing the effect with large samples of participants (present studies $N = 1020$), 2) utilizing modern hierarchical statistical modeling methods to generalize beyond the exact items used, and 3) taking great pains to ensure that participants' truth ratings are not influenced by task demands (Experiments 2 and 3).
+Across our studies, we find further compelling evidence for this “illusory implication” effect, indicating that incidental exposure has genuine impacts on beliefs extending beyond those attributable to metacognitive cues.
 
@@
 All experiments were preregistered prior to data collection and all materials, data, and analysis code have been made publicly available on the Open Science Foundation website (https://osf.io/znq3y/).
-We report our approach to determining sample sizes, all data exclusions, all manipulations, and all measures in each study.
+In these preregistrations, we report our approach to determining sample sizes, all data exclusions, all manipulations, and all measures in each study.
+The analytic plan for Experiments 1 and 2 was preregistered in terms of its general designs, and a more detailed analytic plan was preregistred for Experiment 3.
 Where appropriate, we explicitly differentiate between preregistered analyses and non-preregistered analyses in our results.
@@
 
-```{r fig1, fig.env="figure", fig.pos = "h", fig.align = "center", fig.cap="Average truth ratings for Implication statements in Experiment 1, broken down by exposure (exposed vs.
-unexposed), instructions (fact vs quiz conditions), implication type (false-implied-true vs true-implied-false).
-For visualization purposes, means were calculated by translating ordinal responses onto a scale from -2.5 to 2.5.
-Error bars indicate standard errors.
-"}
-fig1 <- df %>% 
-  filter(phase=="test") %>% 
-  filter(item_type == "B") %>% 
-  group_by(exposed, exposure_cond, B_implied_true) %>% 
-  summarize(
-    M = mean(resp_num),
-    se = sd(resp_num)/sqrt(n()),
-    ll = M - se,
-    ul = M + se
-  ) %>% 
-  mutate(
-    exposure_cond = if_else(exposure_cond=="fact", "Fact condition", "Quiz condition"),
-    exposed = if_else(exposed==1, "Exposed", "Unexposed"),
-    B_implied_true = if_else( B_implied_true == "T implied F", "Contradicted", "Entailed") #as.character(B_implied_true)
-    ) %>% 
-  ggplot(
-    aes(x=B_implied_true, color=factor(exposed), shape=factor(exposed), y = M, ymin = ll, ymax=ul)
-  ) +
-  geom_pointrange(position=position_dodge(width=.5)) +
-  # coord_flip() +
-  scale_color_manual(values = c("#da1e28", "#0f62fe")) +
-  labs(y = "Truth rating", x = "Implication type", color="Exposure", shape="Exposure") +
-  facet_wrap(~exposure_cond) +
-  theme_bw(base_size=9.5) +
-  theme(panel.grid=element_blank(), legend.position="bottom") +
-  theme(axis.text.x = element_text(angle = 10, vjust = 0.5, hjust=.5))
-
-fig1
+```{r fig-exp1, fig.env="figure", fig.pos = "h", fig.width=6, fig.height = 3, fig.align = "center", fig.cap="Left: Test-phase responses and model estimates for Implication statements in Experiment 3, broken down by exposure (exposed vs.
+unexposed) and implication type (contradicted and entailed).
+Bars represent observed proportions.
+Points and errorbars indicate model predictions with 95% credible intervals.
+Right: Difference in proportions of responses between exposed and unexposed conditions at test, aggregated across true and false response options.
+Bars indicate observed differences, with points and errobars indicating model predictions and 50% credible intervals."}
+library(tidybayes)
+
+make_exp_figure(df_reg %>% filter(exposure_cond=="quiz"), fit_imp_quiz_int)
+
 ```
@@
 These participants rated how confident they were that each of the presented statements were true or false.
-The purpose of the quiz condition was to provide "mere exposure" to these statements, to test for illusory truth and illusory implication effects.
+The purpose of the quiz condition was to provide "incidental exposure" to these statements, to test for illusory truth and illusory implication effects.
 
@@
 Similarly, we tested for the illusory implication effect by comparing truth ratings for implication statements whose corresponding premise statements were and were not presented during the exposure phase.
-If mere exposure to "premise" statements affects participants' judgments of the "implication" statements, this would be evidence for an illusory implication effect.
+If incidental exposure to "premise" statements affects participants' judgments of the "implication" statements, this would be evidence for an illusory implication effect.
 
@@
 Four MCMC chains were run for each model, with 2000 samples (1000 burn-in) drawn from each.
-Chains were assessed for convergence with $\hat{R}$ and the total estimated effective sample size was verified to be greater than 1000 for all parameters [@gelman.etal2014].
+MCMC diagnostics indicate good convergence: chains were assessed for convergence with $\hat{R}$ (all parameters $\hat{R} = 1 \pm .01$) and the total estimated effective sample size was verified to be greater than 1000 for all parameters [@gelman.etal2014].
+All models were fit without divergences and without warnings for the Estimated Bayesian Fraction of Missing Information (eBFMI).
 
@@
 
-Figure 1 shows participants’ average truth ratings for the implication statements in the quiz condition.
-Participants gave higher truth ratings following exposure for entailed statements, and somewhat lower truth ratings following exposure for contradicted statements.
+Figure \@ref(fig:fig-exp1) shows participants’ truth judgments for the implication statements in the quiz condition of Experiment 1.
+The left panel of Figure \@ref(fig:fig-exp1) shows a detailed look at the pattern of responding across experimental conditions.
+As shown, a larger proportion of responses for "Maybe ...", "Probably ...", and "Definitely True" and a smaller proportion of "Maybe ...", "Probably ...", and "Definitely False" responses were observed when participants were exposed to the entailing premise statements.
+Crucially, this pattern was reversed for the contradicted statements: fewer participants chose responses indicating truth when they were exposed to the contradicting premises, and instead more chose responses indicating falsehood.
+The right panel of Figure 1 shows this comparison more clearly by aggregating across the different "true" and "false" responses and computing a difference in proportions between statements where participants were exposed to premise statements or not.
+Again, contradictory premises increased false responses and decreased true responses, and the opposite pattern was observed for entailing statements.
 
@@
 
-A regression model was used to tease apart the possible metacognitive effects of mere exposure of and logical implication.
-The model includes a) a binary variable indicating the type of implication statement's actual truth value (true or false), b) a binary predictor for prior exposure to the related premise coded zero when the premise was not seen and 1 when it was seen (capturing potential uni-directional metacognitive effects of exposure), c) and a variable capturing the effect of implications, coded 1 for implied truth, -1 for implied falsehood, and zero when the related premise statement was not seen at exposure.
+A regression model was used to tease apart the possible metacognitive effects of incidental exposure of and logical implication.
+The model includes 1) a binary variable indicating the type of implication statement's actual truth value (true or false), 2) a binary predictor for prior exposure to the related premise coded zero when the premise was not seen and 1 when it was seen (capturing potential uni-directional metacognitive effects of exposure), 3) and a variable capturing the effect of implications, coded 1 for implied truth, -1 for implied falsehood, and zero when the related premise statement was not seen at exposure.
 Thus the "implication" predictor accounts for the effect of exposure to the premise on truth judgments for the implied statements (either positive or negative), while the "exposure" predictor accounts for any positive effect from metacognitive cues.
@@
 The results indicate that both familiarity and the implication of the premise statements affected participants’ truth ratings.
-The implication effect is credibly greater than zero.
+Figure 1 also visualizes the relevant posterior predictive distributions for this model's fixed effects, ignoring the random effects structure.
+The model does well in predicting the proportions of responses, with approximately accurate coverage as expected for the different sets of intervals.
+It is worth noting that these model estimates cannot be expected to perfectly correspond to the observed proportions, because the effects of individual items and participants have been (at least partially) accounted for in the random effects of the models.
 
@@
 Our primary preregistered test compared participants’ truth ratings for the exposed premise statements at test against their ratings for another set of 8 unexposed premise statements.
-This analysis did not find any evidence of an illusory truth effect, with a posterior estimate for the exposure parameter that credibly included zero.
-However, a secondary preregistered analysis comparing participants’ initial truth judgments for the premise-statements to their later truth judgments for those same statements at test did find a credible increase in endorsements, $\beta =$ `r report_reg_coef(fit_premise_within, "phasetest")`.
-
-Although the "quiz" presentation used in Experiment 1 is traditionally seen as providing "mere exposure" to statements, asking participants to judge the truth of these items may have lead them to feel pressured to maintain consistency in their responses.
+This analysis did not find evidence of any meaningful illusory truth effect, with a posterior estimate for the exposure parameter near zero.
+However, a secondary preregistered analysis comparing participants’ initial truth judgments for the premise-statements to their later truth judgments for those same statements at test did find an increase in endorsements, $\beta =$ `r report_reg_coef(fit_premise_within, "phasetest")`.
+
+Although the "quiz" presentation used in Experiment 1 is traditionally seen as providing "incidental exposure" to statements, asking participants to judge the truth of these items may have lead them to feel pressured to maintain consistency in their responses.
 Participants showed a general bias toward judging statements “true” during the exposure phase, responding with some variation of “true” for 71.8% of responses.
@@
 
-```{r fig2, fig.env="figure", fig.pos = "t", fig.align = "center", fig.cap="Average truth ratings for Implication statements in Experiment 2, broken down by exposure (exposed vs.
-unexposed) and implication type (false-implied-true vs true-implied-false).
-For visualization purposes, means were calculated by translating ordinal responses onto a scale from -2.5 to 2.5.
-Error bars indicate standard errors."}
-
-fig2 <- df2 %>% 
-  filter(phase=="test") %>% 
-  mutate(resp_num = recode(response,
-                           `Definitely True` = 2.5,
-                           `Probably True` = 1.5,
-                           `Maybe True` = .5,
-                           `Maybe False` = -.5,
-                           `Probably False` = -1.5,
-                           `Definitely False` = -2.5)
-  ) %>% 
-  mutate(
-    B_implied_true = ifelse(item_pair %in% c("war", "whitehouse", "bezos", "jackson",
-                                          "indonesia", "eiffel", "newyork", "presidents",
-                                          "tour", "africa", "disney", "avengers"), "F implied T", "T implied F"),
-      acc_num = ifelse(B_implied_true == "F implied T", -1*resp_num, resp_num)
-  ) %>% 
-  filter(item_type == "B") %>% 
-  mutate(
-    exposed = if_else(exposed, "Exposed", "Unexposed"),
-    B_implied_true = if_else(B_implied_true == "T implied F", "Contradicted", "Entailed") 
-  ) %>% 
-  group_by(exposed, exposure_cond, B_implied_true) %>% 
-  summarize(
-    M = mean(resp_num),
-    se = sd(resp_num)/sqrt(n()),
-    ll = M - se,
-    ul = M + se
-  ) %>% 
-  ggplot(
-    aes(x=B_implied_true, color=factor(exposed), shape=factor(exposed), y = M, ymin = ll, ymax=ul)
-  ) +
-  geom_pointrange(position=position_dodge(width=.5)) +
-  scale_color_manual(values = c("#da1e28", "#0f62fe")) +
-  theme_bw(base_size=10) +
-  theme(legend.position="bottom", panel.grid=element_blank()) +
-  # coord_flip() +
-  labs(y = "Truth rating", x = "Implication type", color="Exposure", shape="Exposure")
-
-fig2
+```{r fig-exp2, fig.env="figure", fig.pos = "t", fig.width=6, fig.height = 3, fig.align = "center", fig.cap="Left: Test-phase responses and model estimates for Implication statements in Experiment 2, broken down by exposure (exposed vs.
+unexposed) and implication type (contradicted and entailed).
+Bars represent observed proportions.
+Points and errorbars indicate model predictions with 95% credible intervals.
+Right: Difference in proportions of responses between exposed and unexposed conditions at test, aggregated across true and false response options.
+Bars indicate observed differences, with points and errobars indicating model predictions and 50% credible intervals."}
+
+make_exp_figure(df_reg2, fit_imp_exp2)
 ```
@@
 
-Figure 2 shows participants’ average truth ratings for the implication statements with and without exposure to their corresponding premise statements in Experiment 2.
+Figure \@ref(fig:fig-exp2) shows participants’ average truth ratings for the implication statements with and without exposure to their corresponding premise statements in Experiment 2.
 The illusory implication effect was again observed.
@@
 
-```{r fig3, fig.env="figure*", fig.pos = "!hbt", fig.align = "center", fig.width=6.5, fig.height=3, fig.cap="Average truth ratings for individual premise and implication statements in Experiment 2, broken down by exposure (exposed vs.
-unexposed) and implication type (false-implied-true vs true-implied-false).
-For visualization purposes, means were calculated by translating ordinal responses onto a scale from -2.5 to 2.5.
-Error bars indicate standard errors.", num.cols.cap=2}
-
-df2_plotting <- df2 %>% 
-  filter(phase=="test") %>% 
-  mutate(resp_num = recode(response,
-                           `Definitely True` = 2.5,
-                           `Probably True` = 1.5,
-                           `Maybe True` = .5,
-                           `Maybe False` = -.5,
-                           `Probably False` = -1.5,
-                           `Definitely False` = -2.5)
-  ) %>% 
-  mutate(
-    B_implied_true = ifelse(item_pair %in% c("war", "whitehouse", "bezos", "jackson",
-                                          "indonesia", "eiffel", "newyork", "presidents",
-                                          "tour", "africa", "disney", "avengers"), "F implied T", "T implied F")
-  ) %>% 
-  mutate(exposed = if_else(exposed, "Exposed", "Unexposed")) %>% 
-  mutate(B_implied_true = if_else(B_implied_true == "T implied F", "Contradicted", "Entailed") )
-
-df2_extra <- df2_plotting %>% 
-  filter(phase=="test") %>% 
-  filter(item_type == "A") %>% 
-  group_by(exposed, exposure_cond, item_pair, B_implied_true) %>% 
-  summarize(
-    M = mean(resp_num),
-    # se = sd(resp_num)/sqrt(n()),
-    # ll = M - se,
-    # ul = M + se
-  ) %>% 
-  spread(exposed, M) %>% 
-  mutate(diff = abs(Exposed-Unexposed)) %>% 
-  select(-Exposed, -Unexposed)
-
-plt_A <- df2_plotting %>% 
-  filter(phase=="test") %>% 
-  filter(item_type == "A") %>% 
-  group_by(exposed, exposure_cond, item_pair, B_implied_true) %>% 
-  summarize(
-    M = mean(resp_num),
-    se = sd(resp_num)/sqrt(n()),
-    ll = M - se,
-    ul = M + se
-  ) %>% 
-  left_join(df2_extra) %>% 
-  ggplot(
-    aes(x=reorder(item_pair, diff), color=factor(exposed), shape=factor(exposed), y = M, ymin = ll, ymax=ul)
-  ) +
-  geom_pointrange(position=position_dodge(width=.5)) +
-  scale_color_manual(values = c("#da1e28", "#0f62fe")) +
-  coord_flip() +
-  labs(title = "Premise-statements", y = "Truth rating", x= "Item", color="Exposure", shape="Exposure") +
-  theme_bw(base_size=9) +
-  theme(legend.position="right", legend.direction="vertical",panel.grid=element_blank()) +
-  facet_wrap(~B_implied_true, scales="free", ncol=1)
-
-# plt_A
-
-plt_B <- df2_plotting %>%  
-  filter(phase=="test") %>% 
-  filter(item_type == "B") %>% 
-  group_by(exposed, exposure_cond, item_pair, B_implied_true) %>% 
-  summarize(
-    M = mean(resp_num),
-    se = sd(resp_num)/sqrt(n()),
-    ll = M - se,
-    ul = M + se
-  ) %>% 
-  left_join(df2_extra) %>% 
-  ggplot(
-    aes(x=reorder(item_pair, diff), color=factor(exposed), shape=factor(exposed),  y = M, ymin = ll, ymax=ul)
-  ) +
-  geom_pointrange(position=position_dodge(width=.5)) +
-  scale_color_manual(values = c("#da1e28", "#0f62fe")) +
-  coord_flip() +
-  labs(title = "Implication-statements", y = "Truth rating", x= "Item", color="Exposure", shape="Exposure") +
-  theme_bw(base_size=9) +
-  theme(legend.position="right", legend.direction="vertical", panel.grid=element_blank()) +
-  facet_wrap(~B_implied_true, scales="free", ncol=1)
-
-fig3 <- (plt_A + plt_B + guide_area()) + plot_layout(widths=c(4,4,1.5), heights=5, guides="collect")
-
-# fig3 # fix legend
-```
-
 In addition, participant’s truth ratings for the premise statements in Experiment 2 revealed the classic illusory truth effect: premise statements were rated as more true when participants had been exposed to them previously $\beta =$ `r report_reg_coef(fit_prem_exp2, "exposedTRUE")`.
@@
 Participants were instructed to carefully read two statements and select the one that they thought would be more believable to the average person.
-This task should provide a pragmatic context for mere exposure where it was explicitly and pragmatically clear that the statements could be either true or false, but without inducing the consistency pressure that might come with a true or false “quiz”.
-To further ensure this task provided mere exposure, participants were reminded that statements could be true or false each time they judged their believability and comprehension check questions were included before and after the exposure phase to ensure participants remembered this aspect of the task.
+This task should provide a pragmatic context for incidental exposure where it was explicitly and pragmatically clear that the statements could be either true or false, but without inducing the consistency pressure that might come with a true or false “quiz”.
+To further ensure this task provided incidental exposure, participants were reminded that statements could be true or false each time they judged their believability and comprehension check questions were included before and after the exposure phase to ensure participants remembered this aspect of the task.
 
@@
 
-```{r fig-exp3, fig.align='center', fig.cap="Average truth ratings for Implication statements in Experiment 3, broken down by exposure (exposed vs.
-unexposed) and implication type (false-implied-true vs true-implied-false).
-For visualization purposes, means were calculated by translating ordinal responses onto a scale from -2.5 to 2.5.
-Error bars indicate standard errors."}
-
-df3 %>% 
-  filter(phase=="test") %>% 
-  filter(item_type == "B") %>% 
-  mutate(
-    exposed = if_else(exposed, "Exposed", "Unexposed"),
-    B_implied_true = if_else(B_implied_true == "T implied F", "Contradicted", "Entailed") 
-  ) %>% 
-  group_by(exposed, B_implied_true) %>% 
-  summarize(
-    M = mean(resp_num),
-    se = sd(resp_num)/sqrt(n()),
-    ll = M - se,
-    ul = M + se
-  ) %>% 
-  # mutate(B_implied_true = ordered(B_implied_true, levels = c("F implied T", "T implied F"))) %>% 
-  ggplot(
-    aes(x=B_implied_true, color=factor(exposed), shape=factor(exposed), y = M, ymin = ll, ymax=ul)
-  ) +
-  geom_pointrange(position=position_dodge(width=.5))  +
-  scale_color_manual(values = c("#da1e28", "#0f62fe")) +
-  theme_bw(base_size=10) +
-  theme(legend.position="bottom", panel.grid=element_blank()) +
-  # coord_flip() +
-  labs(y = "Truth rating", x = "Implication type", color="Exposure", shape="Exposure")
-```
+```{r fig-exp3, fig.align='center', fig.width=6, fig.height = 3, fig.cap="Left: Test-phase responses and model estimates for Implication statements in Experiment 3, broken down by exposure (exposed vs.
+unexposed) and implication type (contradicted and entailed).
+Bars represent observed proportions.
+Points and errorbars indicate model predictions with 95% credible intervals.
+Right: Difference in proportions of responses between exposed and unexposed conditions at test, aggregated across true and false response options Bars indicate observed differences, with points and errorbars indicating model predictions and 50% credible intervals.
+Predictions were generated from a model fit excluding participants who in any way indicated suspicion about the purpose of the study."}
+
+make_exp_figure(df_reg3, fit_imp_maximal3)
+
+```
+
+Figure \@ref(fig:fig-exp3) presents the results from Experiment 3 as for prior experiments.
+The effects in Experiment 3 are notably subtler than in prior experiments, yet the key findings were again replicated.
+As shown in the right panel, contradictory premises increased false responses and decreased true responses, and the opposite pattern was observed for entailing statements.
 
@@
 
-In Experiments 1 and 2, the effect of exposure was weaker than the measured implication effect, so that average truth ratings were higher after exposure for entailed items and lower for contradicted items (Figures 1 and 2).
-In Experiment 3 however, the estimated effect of exposure and implication were similar, and truth ratings were similar for exposed and unexposed contradicted items (Figure 3).
-Although the estimate for the implication effect in our statistical model was clearly greater than zero, this may nevertheless allow a possible alternate explanation of our findings: that the contradicted items' premise statements simply do not engender as much familiarity or fluency for their implications as do the entailed items, so that the effects differ between these groups.
-Put another way, the identification of the "implication" effect coefficient rests on the assumption that familiarity and fluency effects are not accidentally confounded with item groups.
-
+In the regression models used in Experiments 1 through 3, a coefficient for prior exposure to a related statement is used to represent potential positive effects on truth ratings owing to metacognitive influences, such as fluency and familiarity.
+Consequently, identification of the "implication" effect coefficient rests on the assumption that the directional metacognitive effects of prior exposure are not accidentally confounded with item groups (contradicted and entailed statements).
 To assess the potential for confounding in the familiarity or fluency that exposure to the premise statements would likely engender for the implication statements, we examined the degree of surface-level overlap within each item pair.
 We compared the proportion of shared words (after lemmatization) between premise and implications within each item pair.
-As shown in Figure 4, there is no evidence for confounding between surface-level similarity and item types (median overlap = .19 in false-implied-true items, .21 in true-implied-false items) that would suggest differential effects of fluency or familiarity across types.
-
-Taken in concert with findings from Experiments 1 and 2, where the effect of exposure differed in sign across the item groups, we interpret our regressions as identifying and estimating an implication effect.
+As shown in Figure \@ref(fig:item-fluency-print-fig), there is no evidence for confounding between surface-level similarity and item types (median overlap = .19 in false-implied-true items, .21 in true-implied-false items) that would suggest differential effects of fluency or familiarity across types.
 
@@
 
-We observed an “illusory implication” effect across three preregistered experiments: mere exposure to premises (e.g.
+We observed an “illusory implication” effect across three preregistered experiments: incidental exposure to premises (e.g.
 that “The Tour de France has been held every year since its inception in 1903”) influenced participant’s truth ratings for examples of their logical implications (e.g.
 that “The Tour de France was still held during World War I and World War II”).
-Unlike the illusory truth effect, this novel effect cannot be easily explained by the effects of familiarity or fluency on the judgment process [cf.
+Unlike the illusory truth effect, this effect cannot be easily explained by the effects of familiarity or fluency on the judgment process [cf.
 @arkes.etal1991; @fazio.etal2015; @unkelbach2007; @wang.etal2016]: whereas these metacognitive cues would be expected to uniformly increase endorsements, participants’ truth ratings for implication statements were both increased and decreased according to the implications of the corresponding premises.
-Thus, it appears that mere exposure has a genuine effect on people’s beliefs beyond simply creating a sense of familiarity.
+Thus, it appears that incidental exposure has a genuine effect on people’s beliefs beyond simply creating a sense of fluency.
 
 What cognitive mechanism could explain these findings?
-One potential explanation is that mere exposure to the premise statements leads participants to at least partially adopt them as beliefs.
+One potential explanation is that incidental exposure to the premise statements leads participants to at least partially adopt them as beliefs.
 Simply entertaining the idea may create such an impression despite awareness of its ambiguous truth-value [@gilbert.etal1990; @gilbert1991; @garcia-marques.etal2015; @unkelbach.rom2017].
@@
 A key feature of these mental models is their coherence: updating them is expected to have wide-ranging and coherent implications [@gerstenberg.tenenbaum2017; @powell.etal2023; @thagard1989].
-In the context of mere exposure effects, Unkelbach and Rom [-@unkelbach.rom2017] have proposed an account based on parallel constraint satisfaction within semantic networks, which might plausibly be elaborated to account for the present findings.
-In any case, we take the coherence of inferences following exposure to be evidence for a meaningful influence on people's mental models, and the key to our claim that mere exposure influences beliefs.
+In the context of incidental exposure effects, Unkelbach and Rom [-@unkelbach.rom2017] have proposed an account based on parallel constraint satisfaction within semantic networks, which might plausibly be elaborated to account for the present findings.
+In any case, we take the coherence of inferences following exposure to be evidence for a meaningful influence on people's mental models, and the key to our claim that incidental exposure influences beliefs.
 
@@
 The spread of misinformation and fake news appears to outpace the spread of factual news [@vosoughi.etal2018] and creates massive challenges for online platforms seeking to remove, label, and correct misinformation [@sharma.etal2019].
-These challenges may be further compounded by apparently basic features of human cognition: that mere exposure to false statements can lead them to be _believed_.
+These challenges may be further compounded by apparently basic features of human cognition: that incidental exposure to false statements can lead them to be _believed_.
 

### 2024-06-28 paper/theory-fig.tex
revisions 6-24

\documentclass{standalone}
\usepackage{tikz}
\usepackage{amsmath}
\usetikzlibrary{arrows.meta, positioning}
\begin{document}

\begin{tikzpicture}[
    squarednode/.style={rectangle, draw=black, thick, minimum height=1cm, minimum width=2.5cm},
    ->, thick, >=Stealth
]

% Nodes
\node[squarednode]      (exposure)                                  {Exposure to $S_P$};
\node[squarednode]      (meta)           [below left=of exposure]   {Metacognitive processes};
\node[squarednode]      (belief)         [below right=of exposure]        {Belief in $P$};
\node[squarednode]      (truthSP)        [below left=of belief]  {$\uparrow$ Truth of $S_P$};
% \node[squarednode]      (concepts)       [below=of meta]            {If $S_P$ shares concepts with $S_Q$};
\node[squarednode]      (inference)      [below=of belief]          {Inference};
\node[squarednode]      (actions)        [below right=of belief]    {Actions};
\node[squarednode]      (truthSQ)        [below left=of inference, yshift=-1cm]          {$\uparrow$ Truth of $S_Q$};
\node[squarednode]      (truthSQprime)   [below=of inference, yshift=-1cm]         {$\uparrow$ Truth of $S'_Q$};
\node[squarednode]      (truthSR)        [below right=of inference, yshift=-1cm]       {$\downarrow$ Truth of $S_R$};
\node[draw=none,fill=none]      (concepts)   [below=of meta, yshift=-1cm]  {$S_P$ shares concepts with $S_Q$};
\node[draw=none, fill=none, text width=2.54cm,align=center]      (Noconcepts)   [below=of truthSQprime, yshift=1cm]  {$S_P$ shares no concepts with $S'_Q$};

% Arrows
\draw[->] (exposure) -- (meta);
% \draw[->] (exposure.south) -- (truthSP.north);
\draw[->] (exposure) -- (belief);
\draw[red, very thick] (exposure) -- (belief);
\draw[->] (belief) -- (actions);
\draw[->] (meta) -- (truthSP);
% \draw[->] (meta.south) -- (concepts.north);
% \draw[dashed, ->] (concepts.east) -- (truthSQ.west);
\draw[dashed, -] (meta) -- (concepts);
% \draw[dashed, |->] (concepts) -- (truthSQ.west);
\draw[dashed, ->] (concepts.south) |- (truthSQ.west);
\draw[->] (belief) -- (truthSP);
\draw[->] (belief) -- (inference);

\draw[red, very thick] (belief) -- (inference);
% \draw[->] (inference) -- (truthSR);
\draw[->] (inference)  -- (truthSR) node [midway, above, sloped] (TextNode) {$P \implies \neg R$};
\draw[red, very thick] (inference)  -- (truthSR);
\draw[->] (inference) -- (truthSQ) node [midway, above, sloped] (TextNode) {$P \implies Q$};
\draw[->] (inference) -- (truthSQprime) node [midway, above, sloped] (TextNode) {$P \implies Q$};
\draw[red, very thick] (inference) -- (truthSQprime);

\end{tikzpicture}

\end{document}

### 2024-06-28 paper/fig-code.R

@@ -1,0 +1,145 @@
+make_exp_figure <- function(df_plotting, fit_model){
+  
+  
+  obs_diffs <- df_plotting %>% 
+    filter(phase=="test") %>% 
+    filter(item_type == "B") %>% 
+    mutate(
+      exposed = if_else((exposed == TRUE | exposed==1), "Exposed", "Unexposed"),
+      B_implied_true = if_else(B_implied_true == "T implied F", "Contradicted", "Entailed") 
+    ) %>% 
+    group_by(exposed, B_implied_true, response) %>% 
+    summarize(
+      N = n()
+    ) %>%
+    group_by(B_implied_true, exposed) %>%
+    mutate(P = N/sum(N)) %>%
+    ungroup() %>% 
+    mutate(exposed = if_else((exposed == TRUE | exposed=="Exposed"), "Exposed", "Unexposed")) %>% 
+    
+    mutate(response = if_else(grepl("True", response), "True", "False")) %>%
+    group_by(B_implied_true, response, exposed) %>%
+    summarize(
+      P = sum(P),
+      SE = sqrt( P*(1-P) / sum(N))
+    ) %>% 
+    ungroup() %>%
+    pivot_wider(names_from = exposed, values_from = c(P, SE)) %>%
+    mutate(
+      diff = P_Exposed - P_Unexposed,
+      diff_se = sqrt(SE_Exposed^2 + SE_Unexposed^2)
+    ) %>%
+    rename(Implication = B_implied_true)
+  
+  
+  schema_preds  <-  df_plotting %>%
+    filter(phase=="test") %>%
+    filter(item_type == "B") %>%
+    # filter(implication != 0) %>%
+    distinct(B_implied_true, familiarity, implication, .keep_all = TRUE) %>%
+    select(B_implied_true, familiarity, implication) %>% 
+    mutate(item = "new_item", subj_id = 9999) %>% 
+    add_epred_draws(fit_model, allow_new_levels = TRUE, re_formula = NA) %>% 
+    ungroup() %>% 
+    mutate(
+      exposed = if_else((familiarity == TRUE | familiarity==1), "Exposed", "Unexposed"),
+      B_implied_true = if_else(B_implied_true == "T implied F", "Contradicted", "Entailed") 
+    ) %>%
+    select(-implication, -familiarity, -.row, -.chain, -.iteration) %>%
+    ungroup() %>% 
+    rename(response = .category) %>% 
[…]

### 2024-07-05 paper/paper-papaja-revising.Rmd
resub

@@
 keywords          : "Illusory truth; Metacognition; Cognitive Psychology; Misinformation"
-wordcount         : "5660"
+wordcount         : "6791"
 
@@
 c.f.
-@AlterOppenheimer2009Uniting; @VogelEtAl2020Truth], and it appears unlikely to be workable: people frequently encounter chains of statements invoking different concepts, yet they do so seemingly without issue, as considering one topic does not necessarily make it more difficult to consider another unrelated topic.
+@alter.oppenheimer2009; @VogelEtAl2020Truth], and it appears unlikely to be workable: people frequently encounter chains of statements invoking different concepts, yet they do so seemingly without issue, as considering one topic does not necessarily make it more difficult to consider another unrelated topic.
 

### 2024-09-25 paper/paper-papaja-revising.Rmd
additional revisions for resubmission

@@
     corresponding : yes    # Define only one corresponding author
-    address       : "Arizona State University - West Campus, 4701 W. Thunderbird Rd., Glendale, AZ 85306"
-    email         : "jmikell@asu.edu"
+    address       : "University of California, Irvine"
+    email         : "jmikell@uci.edu"
     role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
@@
   - name          : "Derek Powell"
-    affiliation   : "1"
+    affiliation   : "2"
+    address       : "Arizona State University - West Valley"
     role:
@@
   Numerous psychological findings have shown that incidental exposure to ideas makes those ideas seem more true, a finding commonly referred to as the “illusory truth” effect.
-Under predominant accounts of the illusory truth effect, initial exposure to a statement provides a metacognitive feeling of “fluency” or familiarity that, upon subsequent exposure, leads people to infer the statement is more likely to be true.
+Under many accounts of the illusory truth effect, initial exposure to a statement provides a metacognitive feeling of “fluency” or familiarity that, upon subsequent exposure, leads people to infer the statement is more likely to be true.
 However, genuine beliefs do not only affect truth judgments about individual statements, they also imply other beliefs and drive decision-making.
@@
 keywords          : "Illusory truth; Metacognition; Cognitive Psychology; Misinformation"
-wordcount         : "6791"
+wordcount         : "7531"
 
@@
 library(textstem)
+library(tidybayes)
 
@@
 
+fit_imp_quiz_int_dist <- brm(
+  bf(response ~ familiarity + implication + B_implied_true + 
+    (1 + familiarity + implication|subj_id) + 
+    (1 + familiarity + implication|item_pair)) + lf(disc ~ 0 + familiarity + implication + B_implied_true),  # familiarity + implication + grouping (optional)
+  data = df_reg %>% filter(exposure_cond=="quiz"),
+  family = cumulative(),
+  chains = 4,
+  cores = 4,
+  file="../local/fit_imp_quiz_int_max_dist"
+)
+
 fit_imp_quiz_int <- brm(
-  response ~ familiarity + implication + B_implied_true + 
+  bf(response ~ familiarity + implication + B_implied_true + 
     (1 + familiarity + implication|subj_id) + 
-    (1 + familiarity + implication|item_pair),  # familiarity + implication + grouping (optional)
+    (1 + familiarity + implication|item_pair)),
   data = df_reg %>% filter(exposure_cond=="quiz"),
@@
   file="../local/fit_imp_fact_int_max"
+)
+
+fit_imp_quiz_int_bin <- brm(
+  response_bin ~ familiarity + implication + B_implied_true + 
+    (1 + familiarity + implication|subj_id) + 
+    (1 + familiarity + implication|item_pair),  # familiarity + implication + grouping (optional)
+  data = df_reg %>% filter(exposure_cond=="quiz") %>% mutate(response_bin = if_else(grepl("True", response), 1, 0)),
+  family = bernoulli(),
+  chains = 4,
+  cores = 4,
+  file="../local/fit_imp_quiz_int_max_bin"
 )
@@
 
+fit_imp_exp2_bin <- brm(
+  response_bin ~ familiarity + implication + B_implied_true + 
+    (1 + familiarity + implication|subj_id) + 
+    (1 + familiarity + implication|item_pair),  # familiarity + implication + grouping (optional)
+  data = df_reg2 %>% mutate(response_bin = if_else(grepl("True", response), 1, 0)),
+  family = bernoulli(),
+  prior = set_prior("lkj(3)", class = "cor"),
+  chains = 4,
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta=.9),
+  file="../local/fit_imp_exp2_max_bin"
+)
+
 # Premises
@@
 
+fit_imp_maximal3_bin <- brm(
+  response_bin ~ familiarity + implication + B_implied_true + 
+    (1 + familiarity + implication|subj_id) + 
+    (1 + familiarity + implication|item),  # familiarity + implication + grouping (optional)
+  data = df_reg3_noflag %>% mutate(response_bin = if_else(grepl("True", response), 1, 0)),
+  family = bernoulli(),
+  prior = set_prior("lkj(3)", class = "cor"),
+  chains = 4,
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta=.9),
+  file="../local/fit_imp_maximal3_bin"
+)
+
 ```
@@
 Why should incidental exposure affect later ratings of truth?
-Gilbert [-@gilbert1991] proposed that any idea entertained is by default first believed and must later be "unbelieved." However, predominant theoretical accounts attribute illusory truth effects not to belief but rather to metacognitive experiences of fluency or familiarity [@dechene.etal2010].
+Gilbert [-@gilbert1991] proposed that any idea entertained is by default first believed and must later be "unbelieved." This perspective offers one explanation for the illusory truth effect: people might simply believe what they are told, even when that exposure is incidental, or the pragmatics of the situation indicate that they should withhold judgment [@garcia-marques.etal2015; @silva.etal2017; @unkelbach.rom2017].
+
+In contrast, rather than imagine people are truly so credulous, another set of theoretical accounts attribute illusory truth effects to metacognitive experiences of fluency or familiarity [@dechene.etal2010], rather than genuine influences on beliefs.
 Research on metacognition has shown that people's behavior and judgments are not only influenced by the content of their thoughts, but also by the phenomenological experience of processing those thoughts [@alter.oppenheimer2009; @schwarz2015].
-
 Metacognitive processing "fluency" is the experienced ease of processing or thinking about information [@alter.oppenheimer2009; @wanke.hansen2015].
@@
 
-Both theoretical accounts provide a similar overall view of the illusory truth effect: Roughly, when asked whether something is true or false, people search their memory for knowledge pertaining to belief in the statement, but they also rely on their metacognitive experiences of familiarity and fluency.
+Both metacognitive accounts provide a similar overall view of the illusory truth effect: Roughly, when asked whether something is true or false, people search their memory for knowledge pertaining to belief in the statement, but they also rely on their metacognitive experiences of familiarity and fluency.
 Reliance on metacognitive experiences may be, at least to some extent, adaptive: Familiarity and fluency are both signs that we have “heard this somewhere before.” According to foundational theories of communication, a general assumption that communicators strive to make true utterances is a prerequisite for successful communication and social functioning [@grice1989].
 Thus, if having “heard something before” suggests that someone said it, then ecologically this is a reasonable cue to truth.
-Psychological studies that elicit the illusory truth effect hijack these metacognitive heuristics: They provide this sense of familiarity or fluency, but from a communicative context that lacks any reasonable assumption of truth.
+On this account, psychological studies that elicit the illusory truth effect hijack these metacognitive heuristics: They provide this sense of familiarity or fluency, but from a communicative context that lacks any reasonable assumption of truth.
 
@@
 
-However, there are more to _beliefs_ than ratings of truth.
+The extent of these impacts depends upon the source of the illusory truth effect, whether it is a product of metacognitive processes or results from the adoption of incidentally experienced information as beliefs.
+Beliefs inform actions and decision-making: When truly believed, misinformation can have serious consequences.
+For instance, a recent study found that people who marked just one piece of vaccine misinformation as accurate were nearly twice as likely to be unvaccinated against COVID-19 compared to those who did not endorse any vaccine misinformation [@ognyanova.etal2021].
+In addition, beliefs imply other beliefs: for instance, believing that “it is sunny out” implies the belief that “it is daytime”.
+And believing misinformation, such as “Dominion voting machines were rigged in the 2020 U.S.
+election” might imply the belief that “Donald Trump actually won the 2020 U.S.
+election.”
+
+<!-- To be clear, there are more to _beliefs_ than ratings of truth.
 Though a person’s assent to a proposition (whether they agree with or judge statements expressing it to be truthful) is an important marker and convenient measure of belief, other features of belief are just as essential.
@@
 For instance, a recent study found that people who marked just one piece of vaccine misinformation as accurate were nearly twice as likely to be unvaccinated against COVID-19 compared to those who did not endorse any vaccine misinformation [@ognyanova.etal2021].
+-->
 
@@
 
-Figure \@ref(fig:theory-fig) depicts our hypotheses concerning the causal relationships between latent beliefs, metacognitive experiences, and overt behaviors such truth ratings and actions.
+Some researchers have argued that the illusory truth effect itself demonstrates an influence of prior exposure on beliefs [e.g.
+@garcia-marques.etal2015].
+Yet as these points illustrate, there is more to _belief_ than ratings of truth.
+Though a person’s assent to a proposition (whether they agree with or judge statements expressing it to be truthful) is an important marker and convenient measure of belief, these other features are just as essential.
+Moreover, we argue that where effects of incidental exposure on truth judgments are plausibly attributable to well-established metacognitive processes, these effects do not generally constitute evidence for a change in beliefs.
+
+Nevertheless, these wider consequences can help to determine whether incidental exposure to information modifies beliefs or influences judgments solely through metacognitive processes.
+Figure \@ref(fig:theory-fig) depicts a set of hypotheses concerning the causal relationships between latent beliefs, metacognitive experiences, and overt behaviors such truth ratings and actions.
 As shown in the figure, experiences or manipulations that affect metacognitive experiences, such as incidental exposure, might produce increased truth ratings for statements expressing $P$ (either through lexical or conceptual similarity).
@@
 
-Crucially, our theory draws no connection between metacognitive experiences and reduced truth judgments for statements expressing $R$ where $P$ implies $\neg R$.
+Crucially, our diagram draws no connection between metacognitive experiences and reduced truth judgments for statements expressing $R$ where $P$ implies $\neg R$.
 Such a connection would require a cogent theory of how conceptual "disfluency" might be engendered to reduce truth judgments without the process of inference that would implicate changes in belief.
@@
 
-We argue that where effects of incidental exposure on truth judgments are attributable to metacognitive processes, these effects do not generally constitute evidence for a change in beliefs.
-Applying the theoretical framework sketched in Figure \@ref(fig:theory-fig), cases where incidental exposure to a proposition $P$ increases truth ratings for statements $S_P$ expressing that proposition do not generally constitute evidence that incidental exposure has affected beliefs, since there is a causal pathway between exposure and this result that does not pass through belief.
+Applying the causal model sketched in Figure \@ref(fig:theory-fig), cases where incidental exposure to a proposition $P$ increases truth ratings for statements $S_P$ expressing that proposition do not, in our view, constitute evidence that incidental exposure has affected beliefs, since there is a potential causal pathway between exposure and this result that does not pass through belief.
 Likewise, evidence of exposure to $S_P$ increasing truth ratings for $S_Q$ where $S_Q$ shares a similar set of concepts with $S_P$ do not constitute evidence for effects on beliefs.
@@
 A handful of existing findings provide some preliminary evidence for our hypothesized illusory implication effect.
-A number of researchers have found that incidental exposure affects truth ratings for novel statements contradicting previously exposed statements [@garcia-marques.etal2015; @silva.etal2017; @unkelbach.rom2017].
-To our knowledge, three experiments have found evidence for this pattern of influence: Garcia-Marques (n = 58) [@garcia-marques.etal2015], Silva et al.
+A number of researchers have found that incidental exposure affects truth ratings for novel statements contradicting previously-seen statements [@garcia-marques.etal2015; @silva.etal2017; @unkelbach.rom2017].
+To our knowledge, three studies have found evidence for this specific pattern of influence: Garcia-Marques ($n = 58$) [@garcia-marques.etal2015], Silva et al.
 [-@silva.etal2017] (Exp.
-2, n = 45), and Unkelbach and Rom [-@unkelbach.rom2017] (Exp.
-3; self-referent condition; n = 27).
+2, $n = 45$), and Unkelbach and Rom [-@unkelbach.rom2017] (Exp.
+3; $n = 89$).
 Such a pattern is not easily attributable to purely metacogntiive processes.
-Instead, several of these researchers [e.g., @garcia-marques.etal2015; @silva.etal2017] attribute these lowered truth ratings to the "meaning of the statements" rather than to conceptual fluency.
+<!-- Instead, these researchers [e.g., @garcia-marques.etal2015; @silva.etal2017] attribute these lowered truth ratings to the "meaning of the statements" rather than to conceptual fluency. -->
 
@@
 These researchers explained their findings as resulting from conceptual fluency produced by the initial exposure to statements: prior exposure "activates" the relevant concepts, easing with their retrieval and use in processing subsequent statements and thereby producing different metacognitive experiences.
-Silva and colleagues [-@silva.etal2017] also found that exposure to statements *decreased* truth ratings for rephrased contradictions of those statements.
-However, these findings must be qualified by the exposure context used here: Silva and colleagues asked participants to rate their interest in statements, but did not warn them that the statements could be true or false.
-Thus the pragmatics of the task are unclear, and participants might have believed these statements because they understood them to be presented as facts.
-
-Unkelbach and Rom [-@unkelbach.rom2017] examined both implied and contradictory statements.
-They found that exposure to implied statements increased truth ratings (Exp 4.) and that statements saw a significant decrease in truth ratings following incidental exposure to their contradictions in a "self-relevant" processing condition (Exp.
-4).
+However, Silva and colleagues [-@silva.etal2017] also found that exposure to statements *decreased* truth ratings for rephrased contradictions of those statements, which we identify as a key test of an "illusory implication" effect.
+
+Similarly, Unkelbach and Rom [-@unkelbach.rom2017] examined both implied and contradictory statements.
+They found that exposure to implied statements increased truth ratings (Exp 4) and that statements saw a significant decrease in truth ratings following incidental exposure to their contradictions (Exp.
+3).
 Crucially, the results of this experiment appear to preclude any purely fluency-based account of reduced truth ratings following incidental exposure to contradictory statements: These researchers found that responses were faster for contradictions of previously-seen statements than for novel statements, indicating that processing of these statements was *more* rather than less fluent.
-Thus even if an appropriate notion of "conceptual disfluency" could be developed, it is unlikely it could be an appropriate account of these findings [@unkelbach.rom2017].
+Thus, even if an appropriate notion of "conceptual disfluency" could be developed, it is unlikely it could be an appropriate account of these findings [@unkelbach.rom2017].
 Instead, they and other authors theorize that people tend to believe statements so long as they are generally coherent with respect to existing knowledge, even if they are presented in a context of incidental exposure [@garcia-marques.etal2015; @gilbert1991; @silva.etal2017; @unkelbach.rom2017].
 
+However, these last two sets of findings must be qualified by the exposure context used in these studies: These researchers asked participants to make incidental judgments about these statements but did not warn them that the statements could be true or false [@silva.etal2017; @unkelbach.rom2017].
+Thus, the pragmatics of the task are unclear, and participants might have believed these statements because they understood them to be presented as facts.[^urnote]
+
+[^urnote]: Unkelbach and Rom [-@unkelbach.rom2017] did test a "warning" condition in another experiment of a similar design (Exp.
+1), but here did not find a significant reduction in truth judgments for contradictions of previously-seen statements.
+
 The presently-hypothesized illusory inference effect is a key theoretical link demonstrating the connection between incidental exposure and _beliefs_, and  all of the attendant downstream consequences associated with beliefs.
-A handful of prior findings suggest the existence of such an effect, yet this existing evidence seems far from dispositive, being based on only a handful of small studies with sometimes inconsistent methods.
+A handful of prior findings suggest the existence of such an effect, yet we do not take this existing evidence to be dispositive.
 
@@
 
-Prior work tested the effect of incidental exposure on implied statements using only direct contradictions of the original statement (or paraphrases of direct contradictions).
+Existing research examining the effects of incidental exposure on implied statements has two key limitations.
+First, it has often been conducted in ambiguous contexts where participants might not understand that statements might be true or false, and that therefore do not always clearly induce incidental exposure.
+We address this through study designs inducing clear cases of incidental exposure.
+Second, prior work has only examined effects on direct contradictions of originally-seen statements (or paraphrases of direct contradictions).
 Here, we sought to further test for an "illusory implication effect" by testing the effect of incidental exposure to statements on their more distant logical implications.
@@
 Compared with prior work, we developed materials involving more distant inferences that do not afford direct one-to-one mappings between shared concepts in the premise and implication statements.
-Thus, any reduction in truth ratings for the contradicted statements following exposure to the premise statements are not reasonably attributable to disfluency (closing the alternate pathway depicted in Figure X).
+Thus, any reduction in truth ratings for the contradicted statements following exposure to the premise statements are not reasonably attributable to disfluency (closing the alternate pathway depicted in Figure \@ref(fig:theory-fig)).
 For instance, one statement pair was the premise statement, “No U.S.
@@
 Right: Difference in proportions of responses between exposed and unexposed conditions at test, aggregated across true and false response options.
-Bars indicate observed differences, with points and errobars indicating model predictions and 50% credible intervals."}
+Bars indicate observed differences, with points and errobars indicating binary logistic regression model predictions and 50% credible intervals."}
 library(tidybayes)
 
-make_exp_figure(df_reg %>% filter(exposure_cond=="quiz"), fit_imp_quiz_int)
+make_exp_figure_two_models(df_reg %>% filter(exposure_cond=="quiz"), fit_imp_quiz_int, fit_imp_quiz_int_bin)
 
@@
 This analysis did not find evidence of any meaningful illusory truth effect, with a posterior estimate for the exposure parameter near zero.
-However, a secondary preregistered analysis comparing participants’ initial truth judgments for the premise-statements to their later truth judgments for those same statements at test did find an increase in endorsements, $\beta =$ `r report_reg_coef(fit_premise_within, "phasetest")`.
+However, an additional preregistered analysis comparing participants’ initial truth judgments for the premise-statements to their later truth judgments for those same statements at test did find an increase in endorsements, $\beta =$ `r report_reg_coef(fit_premise_within, "phasetest")`.
 
@@
 Attempting to maintain consistency with these prior responses could have thus influenced their responses to the implication statements and thus the observed effect may be an artifact of the experimental context.
+
+### Secondary and exploratory analyses
+
+Examination of posterior predictive plots across predictor levels indicate some modest misspecification of the model.
+To corroborate the primary preregistered analysis, we conducted a secondary logistic regression analysis after binarizing responses into "true" and "false" categories.
+The estimate of the implication effect in this model was consistent with the primary preregistered analysis $\beta =$ `r report_reg_coef(fit_imp_quiz_int_bin, "implication")`.
+This model's predictions are plotted in Figure \@ref(fig:fig-exp1), right.
 
@@
 Right: Difference in proportions of responses between exposed and unexposed conditions at test, aggregated across true and false response options.
-Bars indicate observed differences, with points and errobars indicating model predictions and 50% credible intervals."}
-
-make_exp_figure(df_reg2, fit_imp_exp2)
+Bars indicate observed differences, with points and errobars indicating binary logistic regression model predictions and 50% credible intervals."}
+
+make_exp_figure_two_models(df_reg2, fit_imp_exp2, fit_imp_exp2_bin)
 ```
@@
 
+### Secondary and exploratory analyses
+
+Examination of posterior predictive plots across predictor levels indicate a good fit of the ordinal regression model to participants' observed responses.
+Nevertheless, for consistency of presentation, we conducted a secondary logistic regression analysis after binarizing responses into "true" and "false" categories.
+The estimate of the implication effect in this model was consistent with the primary preregistered analysis $\beta =$ `r report_reg_coef(fit_imp_exp2_bin, "implication")`.
+This model's predictions are plotted in Figure \@ref(fig:fig-exp2), right.
+
 # Experiment 3
@@
 
-```{r fig-exp3, fig.align='center', fig.width=6, fig.height = 3, fig.cap="Left: Test-phase responses and model estimates for Implication statements in Experiment 3, broken down by exposure (exposed vs.
+Figure \@ref(fig:fig-exp3) presents the results from Experiment 3 as for prior experiments.
+The effects in Experiment 3 are notably subtler than in prior experiments, yet the key findings were again replicated.
+As shown in the right panel, contradictory premises increased false responses and decreased true responses, and the opposite pattern was observed for entailing statements.
+
+### Preregistered analyses
+
+First, in accordance with our preregistered analysis plan, we replicate the primary "maximal" random-effects regression analysis of Experiments 1 and 2.
+Table 3 presents the results of this regression.
+The illusory implication effect was replicated, with a magnitude of the implication coefficient very similar to that observed in Experiment 1.
+
+Next, to explore the possibility that these results might be influenced by task demands, we also carried out a parallel analysis excluding participants whose responses may have been impacted by their suspicions as to the true purpose of the study [@coles.frank2023].
+We excluded both participants who explicitly stated they were suspicious when asked directly, and those who, when forced to guess in an open-ended response, guessed that the true purpose had to do with the first phase somehow affecting their responses in the second.
+Removing these participants did not substantially change the estimate of the implication effect, $\beta =$ `r report_reg_coef(fit_imp_maximal3, "implication")`.
+
+```{r fig-exp3, fig.pos = "H", fig.align='center', fig.width=6, fig.height = 3, fig.cap="Left: Test-phase responses and model estimates for Implication statements in Experiment 3, broken down by exposure (exposed vs.
 unexposed) and implication type (contradicted and entailed).
@@
 Points and errorbars indicate model predictions with 95% credible intervals.
-Right: Difference in proportions of responses between exposed and unexposed conditions at test, aggregated across true and false response options Bars indicate observed differences, with points and errorbars indicating model predictions and 50% credible intervals.
-Predictions were generated from a model fit excluding participants who in any way indicated suspicion about the purpose of the study."}
-
-make_exp_figure(df_reg3, fit_imp_maximal3)
-
-```
-
-Figure \@ref(fig:fig-exp3) presents the results from Experiment 3 as for prior experiments.
-The effects in Experiment 3 are notably subtler than in prior experiments, yet the key findings were again replicated.
-As shown in the right panel, contradictory premises increased false responses and decreased true responses, and the opposite pattern was observed for entailing statements.
-
-### Preregistered analyses
-
-First, in accordance with our preregistered analysis plan, we replicate the primary "maximal" random-effects regression analysis of Experiments 1 and 2.
-Table 3 presents the results of this regression.
-The illusory implication effect was replicated, with a magnitude of the implication coefficient very similar to that observed in Experiment 1.
-
-Next, to explore the possibility that these results might be influenced by task demands, we also carried out a parallel analysis excluding participants whose responses may have been impacted by their suspicions as to the true purpose of the study [@coles.frank2023].
-We excluded both participants who explicitly stated they were suspicious when asked directly, and those who, when forced to guess in an open-ended response, guessed that the true purpose had to do with the first phase somehow affecting their responses in the second.
-Removing these participants did not substantially change the estimate of the implication effect, $\beta =$ `r report_reg_coef(fit_imp_maximal3, "implication")`.
+Right: Difference in proportions of responses between exposed and unexposed conditions at test, aggregated across true and false response options.
+Bars indicate observed differences, with points and errobars indicating binary logistic regression model predictions and 50% credible intervals.
+Plotted responses and model predictions exclude participants who in any way indicated suspicion about the purpose of the study.
+Note that, as before, these model predictions exclude random effects of items and subject and so will not necessarily align precisely with the observed data."}
+
+make_exp_figure_two_models(df_reg3_noflag, fit_imp_maximal3, fit_imp_maximal3_bin)
+
+```
 
@@
 ### Secondary and exploratory analyses
+
+As with Experiment 1, examination of posterior predictive plots across predictor levels indicate some modest misspecification of the model.
+To corroborate the primary preregistered analysis, we conducted a secondary logistic regression analysis after binarizing responses into "true" and "false" categories.
+The estimate of the implication effect in this model was generally consistent with the primary preregistered analysis $\beta =$ `r report_reg_coef(fit_imp_maximal3_bin, "implication")`.
+This model's predictions are plotted in Figure \@ref(fig:fig-exp3), right.
 
@@
 
-```{r, item-fluency-print-fig, fig.align = "center", fig.cap = "Word-level overlap between premise and implication statements (after lemmatization)."}
+```{r, item-fluency-print-fig, fig.pos = "H", fig.align = "center", fig.cap = "Word-level overlap between premise and implication statements (after lemmatization)."}
 item_fluency_fig
@@
 @arkes.etal1991; @fazio.etal2015; @unkelbach2007; @wang.etal2016]: whereas these metacognitive cues would be expected to uniformly increase endorsements, participants’ truth ratings for implication statements were both increased and decreased according to the implications of the corresponding premises.
-Thus, it appears that incidental exposure has a genuine effect on people’s beliefs beyond simply creating a sense of fluency.
-
-What cognitive mechanism could explain these findings?
-One potential explanation is that incidental exposure to the premise statements leads participants to at least partially adopt them as beliefs.
-Simply entertaining the idea may create such an impression despite awareness of its ambiguous truth-value [@gilbert.etal1990; @gilbert1991; @garcia-marques.etal2015; @unkelbach.rom2017].
-If so, this apparently occurs despite participant's awareness that the statements may be false, as ensured by constant reminders and comprehension checks in Experiment 3.
-
-Modern theorists do not conceive of beliefs as lists of propositiosn, but instead as subserved by structured mental models or intuitive theories of how the world works [e.g.
-@gelman.legare2011; @gerstenberg.tenenbaum2017; @powell.etal2023; @johnson-laird2001; @thagard1989].
-Thus, coming to believe a statement entails updating one's mental models with the new information conveyed by that statement.
-A key feature of these mental models is their coherence: updating them is expected to have wide-ranging and coherent implications [@gerstenberg.tenenbaum2017; @powell.etal2023; @thagard1989].
-In the context of incidental exposure effects, Unkelbach and Rom [-@unkelbach.rom2017] have proposed an account based on parallel constraint satisfaction within semantic networks, which might plausibly be elaborated to account for the present findings.
-In any case, we take the coherence of inferences following exposure to be evidence for a meaningful influence on people's mental models, and the key to our claim that incidental exposure influences beliefs.
+Thus, it appears that incidental exposure has a genuine effect on people’s beliefs beyond simply creating a sense of fluency or familiarity.
+
+What cognitive mechanisms could explain these findings?
+Broadly, our findings are most consistent with the perspective advocated by Gilbert [-@gilbert1991; @gilbert.etal1990], who argued that people adopt testimony as belief by default.
+In line with this, a number of researchers have sought to explain how exposure to a related statement can influence truth judgments for novel statements by concluding that statements are simply believed, even following incidental exposure [e.g., @garcia-marques.etal2015; @unkelbach.rom2017].
+We concur with these claims, concluding that our best explanation is that incidental exposure to the premise statements leads participants to at least partially adopt them as beliefs that then influence judgments about the implication statements via a process of inference.
+If so, this apparently occurs despite participant's awareness that the statements may be false, as was carefully ensured by constant reminders and comprehension checks in Experiment 3.
+Simply entertaining an idea, even knowing it could be false, can apparently induce some measure of belief.
+
+Following from this same basic assumption, Unkelbach and Rom [-@unkelbach.rom2017] have proposed a more detailed process-level account of repetition-induced truth effects.
+According to their "Referential Theory", reading or hearing a statement activates the concepts to which it refers, where knowledge of these concepts is represented as a network of nodes connected by excitatory or inhibitory edges.
+A judgment of truth is made based on the overall coherence of the network, given the activation of concepts and their connections (e.g.
+an excitatory connection between two activated concepts would be coherent, but an inhibitory connection between two activated concepts would be incoherent).
+In a spirit similar to Gilbert [-@gilbert1991], the theory explains how repetition can increase judgments of truth by positing that exposure to a statement modifies these links in memory: "If information is new ...
+it will be instigated within memory, and the respective links will form" [@unkelbach.rom2017, p.113].
+At test, repeated statements therefore have "on average more corresponding references which are coherently linked due to the prior presentation" [@unkelbach.rom2017, p.113].
+
+In contrast to the referential theory, we interpret our findings by directly appealing to beliefs and inferences, rather than any more specific representation, like coherentist networks.
+We take beliefs and attendant inferential processes to be core elements in the basic ontology of cognitive science, and so prefer an explanation that avoids positing additional intermediary mechanisms to explain this specific finding, and that allows for a plurality of potential representations.
+
+Nevertheless, as emphasized by the referential theory [@unkelbach.rom2017], interrelations among beliefs and a broader notion of "coherence" are key to understanding these findings.
+Modern theorists generally do not conceive of beliefs as lists of propositions, but instead as subserved by structured mental models or intuitive theories of how the world works [@gelman.legare2011; @gerstenberg.tenenbaum2017; @powell.etal2023; @johnson-laird2001; @thagard1989].
+These mental models might be represented in a variety of formats, such as simulations [e.g.
+@gerstenbergCounterfactualSimulationModel2020]; formal logics, programs, or grammars [e.g.
+@BramleyEtAlLocal; @piantadosiLogicalPrimitivesThought2016; @OverlanEtAl2017Learning]; directed graphs [e.g., @GriffithsTenenbaum2005Structure; @powell.etal2023]; or, indeed, undirected graphs such as coherentist networks [e.g.,  @simonReduxCognitiveConsistency2004; @spellmanCoherenceModelCognitive1993; @thagardExplanatoryCoherence1989].
+On our view, coming to believe a statement entails updating one's relevant mental models, whatever their formats, with the new information conveyed by that statement.
+Importantly, coherence is a key feature of all of these mental models, such that updating them is expected to have wide-ranging and coherent implications [@gerstenberg.tenenbaum2017; @powell.etal2023; @thagard1989].
+We take the coherence of inferences following exposure to be evidence for a meaningful influence on people's mental models, and the key to our claim that incidental exposure influences beliefs.
 

### 2024-09-25 paper/pp_checks.Rmd
additional revisions for resubmission

--- 2024-07-05 paper/paper-papaja-revising.Rmd
@@
 ---
-title             : "Illusory Implications: Incidental exposure to ideas can induce beliefs"
-shorttitle        : "Illusory implications"
-
-author: 
-  - name          : "Justin Mikell"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Arizona State University - West Campus, 4701 W. Thunderbird Rd., Glendale, AZ 85306"
-    email         : "jmikell@asu.edu"
-    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-      - "Investigation"
-      - "Methodology"
-      - "Visualization"
-      - "Writing - Original Draft Preparation"
-      - "Writing - Review & Editing"
-  - name          : "Derek Powell"
-    affiliation   : "1"
-    role:
-      - "Conceptualization"
-      - "Formal Analysis"
-      - "Methodology"
-      - "Writing - Original Draft Preparation"
-      - "Writing - Review & Editing"
-      - "Validation"
-      - "Supervision"
-
-affiliation:
-  - id            : "1"
-    institution   : "Arizona State University"
-
-authornote: |
-  Arizona State University, School of Social and Behavioral Sciences
-
-  Materials, data, and analysis code for all experiments can be found at https://osf.io/znq3y/.
-Preregistration information for Experiments can be found at https://osf.io/c4w8s/ (Exp.
-1), https://osf.io/czkdh/ (Exp.
-2), and https://osf.io/5ewa6 (Exp.
-3).
-Experiments 1 and 2 were first reported in a six-page paper presented at the _43rd Annual Meeting of the Cognitive Science Society_ and appearing in the conference's _Proceedings_.
-This non-archival conference paper included similar narrative motivations for the studies as presented here, with a more limited literature review.
-The narrative interpretations and discussion have changed in light of the findings of the novel third experiment and additional analyses reported here.
-
-  **Ethics Statement**: This research was conducted under the approval of the Institutional Review Board at Arizona State University, IRB protocol ID STUDY00013322.
-Informed consent was obtained from all participants.
-
-abstract: |
-  Numerous psychological findings have shown that incidental exposure to ideas makes those ideas seem more true, a finding commonly referred to as the “illusory truth” effect.
-Under predominant accounts of the illusory truth effect, initial exposure to a statement provides a metacognitive feeling of “fluency” or familiarity that, upon subsequent exposure, leads people to infer the statement is more likely to be true.
-However, genuine beliefs do not only affect truth judgments about individual statements, they also imply other beliefs and drive decision-making.
-Here, we consider whether exposure to “premise” statements affects people's truth ratings for novel “implied” statements, a pattern of findings we call the "illusory implication" effect.
-We argue these effects would constitute evidence for genuine belief change from incidental exposure and  identify a handful of existing findings that offer preliminary support for this claim.
-Building upon these, we conduct three new pre-registered experiments to further test this hypothesis, finding additional evidence that exposure to “premise” statements affected participants’ truth ratings for novel “implied” statements, including for considerably more distant implications than those previously explored.
-Our findings suggest that the effects of incidental exposure reach further than previously thought, with potentially consequential implications for concerns around mis- and disinformation.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "Illusory truth; Metacognition; Cognitive Psychology; Misinformation"
-wordcount         : "6791"
-
-bibliography      : "references.bib"
-
-floatsintext      : yes
-linenumbers       : yes
-draft             : no
-mask              : no
-
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-
-classoption       : "man"
-output            : papaja::apa6_pdf
+title: "Posterior Retrodictive Checks"
+output: pdf_document
+# date: "2024-09-24"
 ---
 
-```{r setup, include = FALSE}
-library("papaja")
-r_refs("r-references.bib")
-```
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE)
 
-```{r analysis-preferences}
-# Seed for random number generation
-set.seed(42)
-# knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
-```
-
-```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
-                      fig.pos = "tb", fig.path='figs/',
-                      echo=F, warning=F, cache=F, 
-                      message=F, sanitize = T)
 ```
@@
   file="../local/fit_imp_exp2_max"
+)
+
+fit_imp_exp2_bin <- brm(
+  response_bin ~ familiarity + implication + B_implied_true + 
+    (1 + familiarity + implication|subj_id) + 
+    (1 + familiarity + implication|item_pair),  # familiarity + implication + grouping (optional)
+  data = df_reg2 %>% mutate(response_bin = if_else(grepl("True", response), 1, 0)),
+  family = bernoulli(),
+  prior = set_prior("lkj(3)", class = "cor"),
+  chains = 4,
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta=.9),
+  file="../local/fit_imp_exp2_max_bin"
 )
@@
   file="../local/fit_imp_rep_maximal3"
+)
+
+fit_imp_maximal3_bin <- brm(
+  response_bin ~ familiarity + implication + B_implied_true + 
+    (1 + familiarity + implication|subj_id) + 
+    (1 + familiarity + implication|item),  # familiarity + implication + grouping (optional)
+  data = df_reg3_noflag %>% mutate(response_bin = if_else(grepl("True", response), 1, 0)),
+  family = bernoulli(),
+  prior = set_prior("lkj(3)", class = "cor"),
+  chains = 4,
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta=.9),
+  file="../local/fit_imp_maximal3_bin"
 )
@@
 
-Every day, people are faced with a barrage of unsupported claims, from pestering ad campaigns to blatant disinformation on social media.
-Altogether, we live within an information environment that is more connected and more saturated than ever before in human history [@bak-coleman.etal2021].
-If we are sufficiently critical consumers of media, can we benefit from this rich access to information without being exploited by advertisers or misled by misinformation?
-Perhaps not.
-Numerous psychological findings indicate that incidental exposure to ideas makes those ideas appear more true, a finding commonly referred to as the “illusory truth” effect [@dekeersmaecker.etal2020; @dechene.etal2010; @fazio.etal2015; @hasher.etal1977; @pennycook.etal2018].
-This effect suggests that we cannot exist in an environment of mis- and disinformation without being affected by it---without exposure to these ideas distorting our sense of what is true.
+## Experiment 1
 
-Studies examining the illusory truth effect typically proceed in at least two phases.
-At exposure, participants are introduced to a set of false statements.
-Typically, the statements are part of a true/false quiz, or a cover story explains they are part of some other innocuous judgment task.
-Then, after some intervening time ranging from minutes to weeks, participants are asked to judge whether these statements are true.
-On average, participants rate the statements as more “true” when they have been exposed to them previously---an illusory truth effect.
-
-Decades of research have demonstrated the consistency and robustness of the illusory truth effect [e.g.
-see @dechene.etal2010].
-The effect has been demonstrated for frivolous trivia questions [e.g.
-@hasher.etal1977; @lacassagne.etal2021; @unkelbach2007; @wang.etal2016] as well as consequential fake news headlines [@pennycook.etal2018].
-The illusory truth effect has also been shown to be robust across people with different levels of cognitive ability, need for cognitive closure, and cognitive styles [@dekeersmaecker.etal2020].
-
-The effect is one of "illusory" truth because it occurs following "incidental exposure" to statements.
-That is, it occurs when the statements are seen or heard in a non-communicative context, such as when they are read during a true/false quiz.
-Generally, being told something---even by a source of unknown trustworthiness---is prima facie reason for believing it [@grice1989; also see @corneille.etal2020].
-But reading a statement on a true/false quiz is not reason for believing it; the statement is just as likely to be false as to be true.
-
-## Theories of the Illusory Truth Effect
-
-Why should incidental exposure affect later ratings of truth?
-Gilbert [-@gilbert1991] proposed that any idea entertained is by default first believed and must later be "unbelieved." However, predominant theoretical accounts attribute illusory truth effects not to belief but rather to metacognitive experiences of fluency or familiarity [@dechene.etal2010].
-Research on metacognition has shown that people's behavior and judgments are not only influenced by the content of their thoughts, but also by the phenomenological experience of processing those thoughts [@alter.oppenheimer2009; @schwarz2015].
-
-Metacognitive processing "fluency" is the experienced ease of processing or thinking about information [@alter.oppenheimer2009; @wanke.hansen2015].
-When thinking is easy and fluent, people tend to judge the targets of that thinking more favorably than if thinking is challenging and disfluent [@wanke.etal1997].
-A number of findings have shown that processing fluency contributes to the illusory truth effect: Incidental prior exposure to statements makes later processing of those statements more fluent, and this metacognitive experience of fluency leads people to rate those statements as more true [@difonzo.etal2016; @fazio.etal2015; @unkelbach2007; @wang.etal2016].
-
-Metacognitive fluency can derive from different sources, and several different types have been shown to influence ratings of truth.
-Fluency can be *perceptual*: statements printed more clearly or in easier-to-read fonts are rated as more true [@ReberSchwarz1999Effects].
-Fluency can be influenced by *memory retrieval*, such that statements that are easier to retreive from memory are rated as more true [@BeggEtAl1992dissociation].
-And fluency can also be semantic or *conceptual* in nature: concepts that are semantically primed are seen as better answers to trivia questions [@KelleyLindsay1993Remembering], and statements evoking previously-activated concepts are rated as more true [@silva.etal2017].
-
-Other researchers have attributed the illusory truth effect to familiarity [e.g.
-@arkes.etal1991; @mitchell.etal2006].
-Prior exposure to statements engenders a feeling of familiarity upon re-exposure to those statements.
-This metacognitive feeling of familiarity provides a cue that leads people to infer the statement is more likely to be true.
-
-Both theoretical accounts provide a similar overall view of the illusory truth effect: Roughly, when asked whether something is true or false, people search their memory for knowledge pertaining to belief in the statement, but they also rely on their metacognitive experiences of familiarity and fluency.
-Reliance on metacognitive experiences may be, at least to some extent, adaptive: Familiarity and fluency are both signs that we have “heard this somewhere before.” According to foundational theories of communication, a general assumption that communicators strive to make true utterances is a prerequisite for successful communication and social functioning [@grice1989].
-Thus, if having “heard something before” suggests that someone said it, then ecologically this is a reasonable cue to truth.
-Psychological studies that elicit the illusory truth effect hijack these metacognitive heuristics: They provide this sense of familiarity or fluency, but from a communicative context that lacks any reasonable assumption of truth.
-
-## Incidental Exposure and Belief
-
-There is something unsettling about the illusory truth effect and the lack of agency seems to imply about our own thinking.
-Even more worrisome, there could be important societal implications if merely being exposed to an idea causes people to adopt it as a belief: Pennycook and colleagues [-@pennycook.etal2018] argue that the illusory truth effect, combined with an environment of pervasive misinformation, has important consequences for the functioning of democratic society [also see @UnkelbachEtAl2019Truth].
-The illusory truth effect suggests that incidental exposure to misinformation could have impacts that cannot be stopped by fact-checking labels [@pennycook.etal2018], nor effectively curbed by retractions or corrective information [c.f.
-@ecker.etal2011; @lewandowsky.etal2012].
-
-However, there are more to _beliefs_ than ratings of truth.
-Though a person’s assent to a proposition (whether they agree with or judge statements expressing it to be truthful) is an important marker and convenient measure of belief, other features of belief are just as essential.
-For one, beliefs imply other beliefs: for instance, believing that “it is sunny out” implies the belief that “it is daytime”.
-And believing misinformation, such as “Dominion voting machines were rigged in the 2020 U.S.
-election” might imply the belief that “Donald Trump actually won the 2020 U.S.
-election.” Beliefs also inform action and decision-making, and whether or not someone holds a belief can be judged (at least partly) from their decisions.
-When truly believed, misinformation can have serious consequences.
-For instance, a recent study found that people who marked just one piece of vaccine misinformation as accurate were nearly twice as likely to be unvaccinated against COVID-19 compared to those who did not endorse any vaccine misinformation [@ognyanova.etal2021].
-
-<!-- It is important to distinguish between statements and the propositions or ideas they express.
-Two different statements might express the same proposition: the pair of statements "I went to the store the day before yesterday" and "I went to the store two days ago" is one example.
-When incidental exposure produces metacognitive experiences like fluency, this might lead to increased endorsement of a specific statement, without truly affecting a person's attitude toward the proposition it represents.
-Instead, the effects of prior exposure would depend on features of subsequently-seen statements (e.g.
-lexical similarity: sharing words; or conceptual similarity: sharing concepts), rather than the propositions they express.
--->
-
-```{r theory-fig, echo = FALSE, message=FALSE, fig.align='center', fig.cap='Diagram depicting the theorized causal relations between latent beliefs, metacogntive experiences, and truth ratings for different types of statements.
-We highlight the paths from observable behavior which can only be traced back to exposure through beliefs in red.', out.width='0.75\\linewidth', fig.pos='h'}
-knitr::include_graphics("theory-fig.pdf")
-```
-
-Figure \@ref(fig:theory-fig) depicts our hypotheses concerning the causal relationships between latent beliefs, metacognitive experiences, and overt behaviors such truth ratings and actions.
-As shown in the figure, experiences or manipulations that affect metacognitive experiences, such as incidental exposure, might produce increased truth ratings for statements expressing $P$ (either through lexical or conceptual similarity).
-However, beliefs should be expected to produce additional consequences.
-Here, we focus on inferences: a manipulation affecting belief in $P$ should increase truth ratings for statements expressing some proposition $Q$ where $P$ implies $Q$, and decrease belief in $R$ where $P$ implies $\neg R$.
-Figure \@ref(fig:theory-fig) also depicts a possible relation between metacognitive experiences and statements expressing $Q$.
-Where a statement $S_Q$ expressing $Q$ invokes a similar set of concepts as $S_P$, conceptual fluency might produce increased truth ratings.
-
-Crucially, our theory draws no connection between metacognitive experiences and reduced truth judgments for statements expressing $R$ where $P$ implies $\neg R$.
-Such a connection would require a cogent theory of how conceptual "disfluency" might be engendered to reduce truth judgments without the process of inference that would implicate changes in belief.
-Current accounts have not developed such a theory [e.g.
-c.f.
-@alter.oppenheimer2009; @VogelEtAl2020Truth], and it appears unlikely to be workable: people frequently encounter chains of statements invoking different concepts, yet they do so seemingly without issue, as considering one topic does not necessarily make it more difficult to consider another unrelated topic.
-
-We argue that where effects of incidental exposure on truth judgments are attributable to metacognitive processes, these effects do not generally constitute evidence for a change in beliefs.
-Applying the theoretical framework sketched in Figure \@ref(fig:theory-fig), cases where incidental exposure to a proposition $P$ increases truth ratings for statements $S_P$ expressing that proposition do not generally constitute evidence that incidental exposure has affected beliefs, since there is a causal pathway between exposure and this result that does not pass through belief.
-Likewise, evidence of exposure to $S_P$ increasing truth ratings for $S_Q$ where $S_Q$ shares a similar set of concepts with $S_P$ do not constitute evidence for effects on beliefs.
-However, evidence that exposure decreases truth ratings for $S_R$ or increases truth ratings for $S'_Q$ where $S'_Q$ does not directly share concepts with $S_P$ would constitute evidence for effects on beliefs.
-
-Practically speaking, the extent to which effects of incidental exposure operate entirely through metacognitive processes stands to determine the extent of their real-world consequences.
-Generally, it is the wider consequences of belief (beyond truth ratings for a specific statement) that should be most concerning.
-If effects of incidental exposure operate only through metacognitive influences, then the impact of prior exposure depends on a subsequent re-exposure to a now fluent or familiar situation or statement (which might include one activating similar concepts).
-To be sure, metacognitive effects might have important consequences even if they do not affect beliefs, such as increasing the tendency to share information online after having been exposed previously [@vellani.etal2023].
-Still, the impact of metacognitive effects seem less likely to drive inferences or decision-making in more novel situations.
-Instead, their influences would be constrained to operating within contexts similar to the initial exposure.
-
-## Is there an "Illusory Implication" Effect?
-
-A crucial feature of belief is that beliefs are connected with other related beliefs.
-Thus, one way to test whether incidental exposure affects beliefs is to examine whether incidental exposure to statements can affect truth ratings for other related or implied statements in ways that cannot be attributed to metacognitive effects.
-We refer to this hypothetical impact of incidental exposure on implied statements as the "Illusory Implication" Effect.
-
-A key test of the illusory implication effect as opposed to other illusory truth effects is whether prior incidental exposure can _reduce_ judgments of truth.
-The experience of fluency is inherently relative: a metacognitive experience is only fluent or disfluent with respect to some point of comparison.
-For instance, utterances can be made relatively fluent by rhyming (McGlone & Tofighbakhsh, 2000) or they can be made relatively disfluent with the addition of speech markers like "um" and "uh" Brennan & Williams (1995).
-However, in the case of repetition-induced fluency (whether conceptual, lexical, or otherwise), repetition can produce fluency, but it is less clear how a such a manipulation might produce disfluency.
-Thus, it is difficult to see how effects of incidental exposure that reduce truth ratings can be explained by metacognitive factors.
-
-A handful of existing findings provide some preliminary evidence for our hypothesized illusory implication effect.
-A number of researchers have found that incidental exposure affects truth ratings for novel statements contradicting previously exposed statements [@garcia-marques.etal2015; @silva.etal2017; @unkelbach.rom2017].
-To our knowledge, three experiments have found evidence for this pattern of influence: Garcia-Marques (n = 58) [@garcia-marques.etal2015], Silva et al.
-[-@silva.etal2017] (Exp.
-2, n = 45), and Unkelbach and Rom [-@unkelbach.rom2017] (Exp.
-3; self-referent condition; n = 27).
-Such a pattern is not easily attributable to purely metacogntiive processes.
-Instead, several of these researchers [e.g., @garcia-marques.etal2015; @silva.etal2017] attribute these lowered truth ratings to the "meaning of the statements" rather than to conceptual fluency.
-
-Garcia-marques and colleagues [-@garcia-marques.etal2015] induced incidental exposure by asking participants to rate how interesting statements were, warning them that half were true and half false.
-After this initial task inducing incidental exposure, a subsequent quiz featured a mix of novel statements, repeated statements, and statements contradicting those previously seen either immediately or after a 1-week delay.
-In the immediate-test condition, contradictory statements were rated as less true than novel statements.[^bacon]
-
-[^bacon]: These researchers used materials following those used by Bacon [-@bacon1979, Exp.
-2], who induced incidental exposure through a series of true-or-false quizzes.
-In this study, contradictory statements were rated as less true following a two-week delay, but only when participants were warned that the statements had been changed.
-This warning provides an avenue by which metacognitive cues could be used to inform truth judgments about contradictions without inducing beliefs.
-
-Silva and colleagues [-@silva.etal2017] found that exposure to statements increased truth ratings for paraphrases of those statements that expressed the same proposition while sharing only minimal surface-level linguistic features.
-These researchers explained their findings as resulting from conceptual fluency produced by the initial exposure to statements: prior exposure "activates" the relevant concepts, easing with their retrieval and use in processing subsequent statements and thereby producing different metacognitive experiences.
-Silva and colleagues [-@silva.etal2017] also found that exposure to statements *decreased* truth ratings for rephrased contradictions of those statements.
-However, these findings must be qualified by the exposure context used here: Silva and colleagues asked participants to rate their interest in statements, but did not warn them that the statements could be true or false.
-Thus the pragmatics of the task are unclear, and participants might have believed these statements because they understood them to be presented as facts.
-
-Unkelbach and Rom [-@unkelbach.rom2017] examined both implied and contradictory statements.
-They found that exposure to implied statements increased truth ratings (Exp 4.) and that statements saw a significant decrease in truth ratings following incidental exposure to their contradictions in a "self-relevant" processing condition (Exp.
-4).
-Crucially, the results of this experiment appear to preclude any purely fluency-based account of reduced truth ratings following incidental exposure to contradictory statements: These researchers found that responses were faster for contradictions of previously-seen statements than for novel statements, indicating that processing of these statements was *more* rather than less fluent.
-Thus even if an appropriate notion of "conceptual disfluency" could be developed, it is unlikely it could be an appropriate account of these findings [@unkelbach.rom2017].
-Instead, they and other authors theorize that people tend to believe statements so long as they are generally coherent with respect to existing knowledge, even if they are presented in a context of incidental exposure [@garcia-marques.etal2015; @gilbert1991; @silva.etal2017; @unkelbach.rom2017].
-
-The presently-hypothesized illusory inference effect is a key theoretical link demonstrating the connection between incidental exposure and _beliefs_, and  all of the attendant downstream consequences associated with beliefs.
-A handful of prior findings suggest the existence of such an effect, yet this existing evidence seems far from dispositive, being based on only a handful of small studies with sometimes inconsistent methods.
-
-## The present studies
-
-Prior work tested the effect of incidental exposure on implied statements using only direct contradictions of the original statement (or paraphrases of direct contradictions).
-Here, we sought to further test for an "illusory implication effect" by testing the effect of incidental exposure to statements on their more distant logical implications.
-To do so, we designed a series of studies to examine whether incidental exposure to one statement (the “premise”) could affect truth ratings for another statement this premise would logically entail to be either true or false (the “implication”).
-
-In our studies, we sought to eliminate as much as possible the role of perceptual and conceptual fluency in affecting truth judgments.
-These efforts are aimed at demonstrating as clearly as possible a causal pathway  between incidental exposure and effects on truth ratings through changes in belief.
-That is, establishing a pathway from exposure to increased truth ratings for implications $Q$ and decreased truth ratings for contradictions $R$.
-Compared with prior work, we developed materials involving more distant inferences that do not afford direct one-to-one mappings between shared concepts in the premise and implication statements.
-Thus, any reduction in truth ratings for the contradicted statements following exposure to the premise statements are not reasonably attributable to disfluency (closing the alternate pathway depicted in Figure X).
-For instance, one statement pair was the premise statement, “No U.S.
-astronauts have died since the Challenger explosion in 1986” paired with the implied statement, “The space shuttle Columbia disintegrated over Texas in 2003” (true, but contradicted by the "premise").
-In contrast with prior work examining conflicting statements, there is no direct one-to-one mapping of concepts between the two statements.
-Instead, a chain of inferences must be developed for a conclusion to be drawn: that the space shuttle Columbia contained astronauts, that its disintegration would have killed them, and that 2003 is later than 1986.
-
-Additionally, we make several further contributions toward definitively demonstrating the illusory implication effect beyond existing studies by 1) testing the effect with large samples of participants (present studies $N = 1020$), 2) utilizing modern hierarchical statistical modeling methods to generalize beyond the exact items used, and 3) taking great pains to ensure that participants' truth ratings are not influenced by task demands (Experiments 2 and 3).
-Across our studies, we find further compelling evidence for this “illusory implication” effect, indicating that incidental exposure has genuine impacts on beliefs extending beyond those attributable to metacognitive cues.
-
-# Transparency and Openness
-
-All experiments were preregistered prior to data collection and all materials, data, and analysis code have been made publicly available on the Open Science Foundation website (https://osf.io/znq3y/).
-In these preregistrations, we report our approach to determining sample sizes, all data exclusions, all manipulations, and all measures in each study.
-The analytic plan for Experiments 1 and 2 was preregistered in terms of its general designs, and a more detailed analytic plan was preregistred for Experiment 3.
-Where appropriate, we explicitly differentiate between preregistered analyses and non-preregistered analyses in our results.
-All statistical models were fit using the `brms` R package [@burkner2017].
-Our study was reviewed and approved by the Arizona State University Institutional Review Board under project identifier `STUDY00013322`.
-Participants provided informed consent before beginning the study.
-
-# Experiment 1
-
-## Methods
-
-### Participants
-
-A total of 400 Participants were recruited through CloudResearch.
-Basic attention check questions were used to screen inattentive participants.
-These questions asked participants to simply give a particular response.
-Participants who failed any attention checks or who indicated during debrief that they had searched online for question answers were excluded from analysis.
-This left a final sample of `r nrow(df_demo)` (`r exp1_n_f` female, median age `r exp1_age` years-old).
-
-```{r fig-exp1, fig.env="figure", fig.pos = "h", fig.width=6, fig.height = 3, fig.align = "center", fig.cap="Left: Test-phase responses and model estimates for Implication statements in Experiment 3, broken down by exposure (exposed vs.
-unexposed) and implication type (contradicted and entailed).
-Bars represent observed proportions.
-Points and errorbars indicate model predictions with 95% credible intervals.
-Right: Difference in proportions of responses between exposed and unexposed conditions at test, aggregated across true and false response options.
-Bars indicate observed differences, with points and errobars indicating model predictions and 50% credible intervals."}
-library(tidybayes)
-
-make_exp_figure(df_reg %>% filter(exposure_cond=="quiz"), fit_imp_quiz_int)
+```{r cars}
+pp_check(fit_imp_quiz_int, type = 'bars_grouped', group = 'implication', ndraws = 500)
+pp_check(fit_imp_quiz_int, type = 'bars_grouped', group = 'familiarity', ndraws = 500)
+pp_check(fit_imp_quiz_int, type = 'bars_grouped', group = 'B_implied_true', ndraws = 500)
 
@@
 
-### Materials
+## Experiment 2
 
-We created 24 pairs of statements presenting claims about history.
-Each pair consisted of a "premise" statement and a "implication" statement.
-Each “premise” statement was a falsehood that either entailed or contradicted the “implication” statement.
-Of the pairs, 12 had premise statements implying that a true statement was false ("contradicted") and the other 12 had premise statements implying another false statement was true ("entailed").
-For instance, one "contradicted" pair was the premise statement, “No U.S.
-astronauts have died since the Challenger explosion in 1986” paired with the implied statement, “The space shuttle Columbia disintegrated over Texas in 2003” (true, but contradicted by the "premise").
-An example of an "entailed" pair was the premise statement, "The Tour de France has been held every year since its inception in 1903" paired with the implied statement, "The Tour de France was still held during WWI and WWII" (false, but entailed by the "premise").
-
-### Procedures
-
-The study proceeded in three phases: 1) an exposure phase, 2) a distraction phase, and 3) a final testing phase.
-
-At the beginning of the study, participants were randomly assigned to either the “fact” (_n_ = 100) or “quiz” (_n_ = 300) exposure condition and to one of three counterbalancing conditions.
-At the exposure phase, participants were presented with a subset of the "premise" statements (8 of the 24 total, counterbalanaced across participants) as well a set of control statements (12 true, 4 false).
-Participants assigned to the quiz and fact conditions received different instructions and performed different tasks in the exposure phase.
-
-Participants in the fact condition were told the study’s main purpose was to learn more about how people learn and apply new knowledge.
-They were informed that the initial set of statements they would see were all facts, and were asked to rate how surprising each “fact” was to them.
-The true purpose of this condition was to test whether participants would successfully draw inferences from the "premise" to the "implied" statements when told that the premises were true.
-
-Participants in the quiz condition were told that they were to be given a true/false quiz, so that some of the statements would be true and some false.
-These participants rated how confident they were that each of the presented statements were true or false.
-The purpose of the quiz condition was to provide "incidental exposure" to these statements, to test for illusory truth and illusory implication effects.
-
-After their initial exposure to the statements, participants provided basic demographic information and were then presented with the expanded 7 question version of the Cognitive Reflection Task [CRT, @frederick2005; @toplak.etal2014].
-These tasks served to provide a period of distraction between exposure and test.
-
-Finally, the testing phase for all participants consisted of a true-or-false quiz, where participants were asked to rate their confidence that each statement was true or false.
-These ratings were made on a 6-point scale from "Definitely False" to "Definitely True".
-The items tested included both "premise" and "implied" statements from the 24 item-pairs.
-Participants were quizzed on the 8 previously-exposed premise statements, as well as their corresponding "implied" statements.
-For comparison within subjects, they were also quizzed on 8 new premise statements and 8 unrelated implication statements.
-Participants were randomly assigned into three counterbalancing conditions that varied which of the statement-pairs were assigned to each exposure/test combination.
-
-We tested for the illusory truth effect by comparing truth ratings for premise statements seen during exposure to those that had not been seen.
-Similarly, we tested for the illusory implication effect by comparing truth ratings for implication statements whose corresponding premise statements were and were not presented during the exposure phase.
-If incidental exposure to "premise" statements affects participants' judgments of the "implication" statements, this would be evidence for an illusory implication effect.
-
-At the end of the study, participants were debriefed and presented with a list of the false statements they had seen.
-
-## Results
-
-Participants made their truth ratings in the test phase on a 6-point scale from “Definitely false” to “Definitely true”.
-To properly treat these Likert-style responses, the truth ratings were analyzed using multilevel Bayesian cumulative ordinal regression models [@burkner.vuorre2019] with random intercepts and slopes for participants and items.
-Cumulative ordinal regression models assume that participants’ discrete responses are driven by a continuous latent variable and a set of $k-1$ thresholds determining the range of the continuous variable corresponding to each of the ordinal response options.
-This model helps to account for potential differences in scale usage as well as the bounded nature of the response scale.
-
-All models were fit using the `brms` R package [@burkner2017], with model posteriors estimated using the No-U-Turn Markov Chain Monte Carlo (MCMC) sampler implemented in Stan.
-Four MCMC chains were run for each model, with 2000 samples (1000 burn-in) drawn from each.
-MCMC diagnostics indicate good convergence: chains were assessed for convergence with $\hat{R}$ (all parameters $\hat{R} = 1 \pm .01$) and the total estimated effective sample size was verified to be greater than 1000 for all parameters [@gelman.etal2014].
-All models were fit without divergences and without warnings for the Estimated Bayesian Fraction of Missing Information (eBFMI).
-
-### Fact condition
-
-First, we examined truth ratings for the premise and implications statements among participants who were told that the false statements were true at exposure (“fact” condition).
-As expected, exposure to the premise statements when presented as “facts” increased endorsement at test, $\beta =$ `r report_reg_coef(fit_premise_fact, "exposed")`.
-In addition, participants successfully drew inferences from these “facts”, resulting in decreased accuracy for the implication statements at test, $\beta =$ `r report_reg_coef(fit_imp_fact, "exposed")`.
-
-### Quiz condition
-
-Figure \@ref(fig:fig-exp1) shows participants’ truth judgments for the implication statements in the quiz condition of Experiment 1.
-The left panel of Figure \@ref(fig:fig-exp1) shows a detailed look at the pattern of responding across experimental conditions.
-As shown, a larger proportion of responses for "Maybe ...", "Probably ...", and "Definitely True" and a smaller proportion of "Maybe ...", "Probably ...", and "Definitely False" responses were observed when participants were exposed to the entailing premise statements.
-Crucially, this pattern was reversed for the contradicted statements: fewer participants chose responses indicating truth when they were exposed to the contradicting premises, and instead more chose responses indicating falsehood.
-The right panel of Figure 1 shows this comparison more clearly by aggregating across the different "true" and "false" responses and computing a difference in proportions between statements where participants were exposed to premise statements or not.
-Again, contradictory premises increased false responses and decreased true responses, and the opposite pattern was observed for entailing statements.
-
-```{r, showtable1, echo=F, message=F}
-mytable1
-```
-
-```{r, showtable2, echo=F, message=F}
-mytable2
-```
-
-An increase in truth ratings for the entailed statements could potentially be explained by familiarity or fluency.
-This would be consistent with findings from Arkes and colleagues [-@arkes.etal1991], who found evidence for illusory truth for novel statements that were on the same topic as previously-seen statements.
-However, simple metacognitive cues cannot explain the decrease in truth ratings for the contradicted items, as familiarity or fluency should encourage uniformly higher, not lower, truth ratings.
-
-A regression model was used to tease apart the possible metacognitive effects of incidental exposure of and logical implication.
-The model includes 1) a binary variable indicating the type of implication statement's actual truth value (true or false), 2) a binary predictor for prior exposure to the related premise coded zero when the premise was not seen and 1 when it was seen (capturing potential uni-directional metacognitive effects of exposure), 3) and a variable capturing the effect of implications, coded 1 for implied truth, -1 for implied falsehood, and zero when the related premise statement was not seen at exposure.
-Thus the "implication" predictor accounts for the effect of exposure to the premise on truth judgments for the implied statements (either positive or negative), while the "exposure" predictor accounts for any positive effect from metacognitive cues.
-We also incorporated "maximal" random intercepts and slopes for all terms varying by subject and item [@barr.etal2013].
-Expressed in the common “lme4” syntax [@bates.etal2015], the regression model was:
-
-\begin{align*}
-\text{response} \sim 
- \text{statement truth} &+ \text{exposure} + \text{implication} \\
-&+ (1 + \text{exposure} + \text{implication}|\text{subject}) \\
-&+ (1 + \text{exposure} + \text{implication}|\text{item})
-\end{align*}
-
-Table 1 presents a summary of the posterior distribution estimated for the population-level coefficients in this model.
-The results indicate that both familiarity and the implication of the premise statements affected participants’ truth ratings.
-Figure 1 also visualizes the relevant posterior predictive distributions for this model's fixed effects, ignoring the random effects structure.
-The model does well in predicting the proportions of responses, with approximately accurate coverage as expected for the different sets of intervals.
-It is worth noting that these model estimates cannot be expected to perfectly correspond to the observed proportions, because the effects of individual items and participants have been (at least partially) accounted for in the random effects of the models.
-
-Somewhat surprisingly, the effects of exposure on participants’ truth ratings for the premise statements were very subtle.
-Our primary preregistered test compared participants’ truth ratings for the exposed premise statements at test against their ratings for another set of 8 unexposed premise statements.
-This analysis did not find evidence of any meaningful illusory truth effect, with a posterior estimate for the exposure parameter near zero.
-However, a secondary preregistered analysis comparing participants’ initial truth judgments for the premise-statements to their later truth judgments for those same statements at test did find an increase in endorsements, $\beta =$ `r report_reg_coef(fit_premise_within, "phasetest")`.
-
-Although the "quiz" presentation used in Experiment 1 is traditionally seen as providing "incidental exposure" to statements, asking participants to judge the truth of these items may have lead them to feel pressured to maintain consistency in their responses.
-Participants showed a general bias toward judging statements “true” during the exposure phase, responding with some variation of “true” for 71.8% of responses.
-Attempting to maintain consistency with these prior responses could have thus influenced their responses to the implication statements and thus the observed effect may be an artifact of the experimental context.
-
-# Experiment 2
-
-Experiment 2 was conducted to address the possibility that a pressure for internal consistency produced the effects on the implication-statement truth ratings observed in Experiment 1.
-
-```{r fig-exp2, fig.env="figure", fig.pos = "t", fig.width=6, fig.height = 3, fig.align = "center", fig.cap="Left: Test-phase responses and model estimates for Implication statements in Experiment 2, broken down by exposure (exposed vs.
-unexposed) and implication type (contradicted and entailed).
-Bars represent observed proportions.
-Points and errorbars indicate model predictions with 95% credible intervals.
-Right: Difference in proportions of responses between exposed and unexposed conditions at test, aggregated across true and false response options.
-Bars indicate observed differences, with points and errobars indicating model predictions and 50% credible intervals."}
-
-make_exp_figure(df_reg2, fit_imp_exp2)
-```
-
-## Methods
-
-### Participants
-
-A total of 300 participants were recruited from CloudResearch using procedures identical to Experiment 1.
-As before, participants who failed attention check questions or who indicated they had looked up answers were excluded from analyses, leaving a final sample of `r nrow(df2_demo)` (`r exp2_n_f` female, median age `r exp2_age` years-old).
-
-### Materials and procedures
-
-All items and procedures were identical to Experiment 1 except for three changes.
-
-First, all participants were assigned to a new “interest” condition.
-At exposure, participants were told that they would see a set of statements, some true and some false, and they were instructed to rate how interesting each statement was.
-This change was made to avoid forcing participants to make a true/false judgment for the exposed premise statements, which could have pushed them to try to make coherent or consistent responses to the related implication statements.
-Importantly, as with a true/false quiz, the presentation of statements in this context should not warrant any inference as to the truth of those statements.
-
-Secondly, to provide a longer period of distraction between exposure and test, the ordering of the items in the test phase was rearranged.
-Participants first judged the truth of 24 control items (12 true and 12 false) to extend the time between viewing the main test items.
-
-Third, another change in test presentation order further prevented any consistency-pressure: Participants judged the truth of all of the implication statements before judging the premise statements.
-
-## Results
-
-Figure \@ref(fig:fig-exp2) shows participants’ average truth ratings for the implication statements with and without exposure to their corresponding premise statements in Experiment 2.
-The illusory implication effect was again observed.
-As shown in the figure, truth ratings were clearly affected by exposure: participants gave higher truth ratings following exposure for entailed statements, and lower truth ratings following exposure for contradicted statements.
-Table 2 shows the posterior population-level estimates for an identical regression as was conducted for Experiment 1.
-Experiment 2 replicated the effects of Experiment 1 in a revised design that eliminated any consistency pressure or demands on participants.
-In fact, the magnitude of the effect for the implication statements was somewhat larger in Experiment 2 than in Experiment 1.
-
-In addition, participant’s truth ratings for the premise statements in Experiment 2 revealed the classic illusory truth effect: premise statements were rated as more true when participants had been exposed to them previously $\beta =$ `r report_reg_coef(fit_prem_exp2, "exposedTRUE")`.
-
-The illusory implication effect for the implication statements was generally similar in magnitude to the illusory truth effect observed among the premise statements.
-
-# Experiment 3
-
-Finally, we sought to replicate these findings once again and to rule out the possibility that the illusory implication effect is a product of experimental idiosyncrasies, such as misapprehension of instructions or potential task demands.
-In Experiment 3, we ensured participants were attending carefully and understood the instructions through additional attention and comprehension checks.
-In addition, we evaluated the potential for unanticipated task demands by excluding from analysis any participants who had suspicions about the study or who guessed its true purpose.
-
-## Method
-
-### Participants
-
-We recruited 540 participants from CloudResearch's Connect service and excluded participants who failed attention or instruction comprehension checks, leaving a final sample of `r nrow(df3_demo)` (`r exp3_n_f` female, median age `r exp3_age` years-old).
-In a secondary set of pre-registered analyses, we also excluded participants who suspected that the study had an undisclosed purpose (`r exp3_susp_count`) and additionally those who guessed the true purpose of the study in an open-ended response (`r exp3_flagged_count`), leaving `r nrow(df3_demo) - exp3_susp_count - exp3_flagged_count` participants for these analyses.
-
-### Materials and procedures
-
-All item pairs were identical to Experiment 2.
-However, several changes were made to the study's procedures to limit the potential that the illusory implication effect could be produced by idiosyncratic features of the experimental task.
-
-First, we sought to further ensure that participants would not mistake the exposure context for a communicative context, while also not inducing any consistency pressure at test.
-To this end, we created a new "believability" task for the exposure phase.
-Participants were instructed to carefully read two statements and select the one that they thought would be more believable to the average person.
-This task should provide a pragmatic context for incidental exposure where it was explicitly and pragmatically clear that the statements could be either true or false, but without inducing the consistency pressure that might come with a true or false “quiz”.
-To further ensure this task provided incidental exposure, participants were reminded that statements could be true or false each time they judged their believability and comprehension check questions were included before and after the exposure phase to ensure participants remembered this aspect of the task.
-
-To improve our ability to measure the implication effect, in Experiment 3 participants were exposed to 12 of the premise statements at exposure rather than eight.
-As a consequence, we did not examine illusory truth effects.
-The 24 premise statements were randomly sorted into 12 comparison questions, and each participant saw 6 of these pairs at exposure, along with five control pairs.
-
-Following the test phase, two new questions assessed the potential influence of task demands.
-Participants were asked if they ever suspected the study had an undisclosed purpose (yes/no), and then regardless of their answer asked to explain their suspicions in an open-ended response.
-Responses were coded to identify participants who guessed that the true purpose concerned how their responses to the test phase would be affected by the exposure phase or by seeing misinformation.
-These participants were excluded from certain analyses following our pre-registered analysis plan.
-
-Finally, six attention check questions were more tightly integrated into the task to better ensure attendant responding.
-Following the "infrequency-items" approach of [@zorowitz.etal2023], exposure phase attention checks consisted of a very well-known fact and one wholly unbelievable fiction.
-Test phase attention checks consisted of very easy trivia items that all participants should know.
-Participants who answered any of the six attention checks incorrectly were excluded from analyses.
-
-## Results
-
-```{r fig-exp3, fig.align='center', fig.width=6, fig.height = 3, fig.cap="Left: Test-phase responses and model estimates for Implication statements in Experiment 3, broken down by exposure (exposed vs.
-unexposed) and implication type (contradicted and entailed).
-Bars represent observed proportions.
-Points and errorbars indicate model predictions with 95% credible intervals.
-Right: Difference in proportions of responses between exposed and unexposed conditions at test, aggregated across true and false response options Bars indicate observed differences, with points and errorbars indicating model predictions and 50% credible intervals.
-Predictions were generated from a model fit excluding participants who in any way indicated suspicion about the purpose of the study."}
-
-make_exp_figure(df_reg3, fit_imp_maximal3)
+```{r pressure}
+pp_check(fit_imp_exp2, type = 'bars_grouped', group = 'implication', ndraws = 500)
+pp_check(fit_imp_exp2, type = 'bars_grouped', group = 'familiarity', ndraws = 500)
+pp_check(fit_imp_exp2, type = 'bars_grouped', group = 'B_implied_true', ndraws = 500)
 
@@
 
-Figure \@ref(fig:fig-exp3) presents the results from Experiment 3 as for prior experiments.
-The effects in Experiment 3 are notably subtler than in prior experiments, yet the key findings were again replicated.
-As shown in the right panel, contradictory premises increased false responses and decreased true responses, and the opposite pattern was observed for entailing statements.
+## Experiment 3
 
-### Preregistered analyses
+```{r}
 
-First, in accordance with our preregistered analysis plan, we replicate the primary "maximal" random-effects regression analysis of Experiments 1 and 2.
-Table 3 presents the results of this regression.
-The illusory implication effect was replicated, with a magnitude of the implication coefficient very similar to that observed in Experiment 1.
+pp_check(fit_imp_maximal3, type = 'bars_grouped', group = 'implication', ndraws = 500)
+pp_check(fit_imp_maximal3, type = 'bars_grouped', group = 'familiarity', ndraws = 500)
+pp_check(fit_imp_maximal3, type = 'bars_grouped', group = 'B_implied_true', ndraws = 500)
 
-Next, to explore the possibility that these results might be influenced by task demands, we also carried out a parallel analysis excluding participants whose responses may have been impacted by their suspicions as to the true purpose of the study [@coles.frank2023].
-We excluded both participants who explicitly stated they were suspicious when asked directly, and those who, when forced to guess in an open-ended response, guessed that the true purpose had to do with the first phase somehow affecting their responses in the second.
-Removing these participants did not substantially change the estimate of the implication effect, $\beta =$ `r report_reg_coef(fit_imp_maximal3, "implication")`.
-
-```{r, showtable3, echo=F, message=F}
-mytable3
 ```
-
-### Secondary and exploratory analyses
-
-In the regression models used in Experiments 1 through 3, a coefficient for prior exposure to a related statement is used to represent potential positive effects on truth ratings owing to metacognitive influences, such as fluency and familiarity.
-Consequently, identification of the "implication" effect coefficient rests on the assumption that the directional metacognitive effects of prior exposure are not accidentally confounded with item groups (contradicted and entailed statements).
-To assess the potential for confounding in the familiarity or fluency that exposure to the premise statements would likely engender for the implication statements, we examined the degree of surface-level overlap within each item pair.
-We compared the proportion of shared words (after lemmatization) between premise and implications within each item pair.
-As shown in Figure \@ref(fig:item-fluency-print-fig), there is no evidence for confounding between surface-level similarity and item types (median overlap = .19 in false-implied-true items, .21 in true-implied-false items) that would suggest differential effects of fluency or familiarity across types.
-
-```{r, item-fluency-print-fig, fig.align = "center", fig.cap = "Word-level overlap between premise and implication statements (after lemmatization)."}
-item_fluency_fig
-```
-
-# Discussion
-
-We observed an “illusory implication” effect across three preregistered experiments: incidental exposure to premises (e.g.
-that “The Tour de France has been held every year since its inception in 1903”) influenced participant’s truth ratings for examples of their logical implications (e.g.
-that “The Tour de France was still held during World War I and World War II”).
-Unlike the illusory truth effect, this effect cannot be easily explained by the effects of familiarity or fluency on the judgment process [cf.
-@arkes.etal1991; @fazio.etal2015; @unkelbach2007; @wang.etal2016]: whereas these metacognitive cues would be expected to uniformly increase endorsements, participants’ truth ratings for implication statements were both increased and decreased according to the implications of the corresponding premises.
-Thus, it appears that incidental exposure has a genuine effect on people’s beliefs beyond simply creating a sense of fluency.
-
-What cognitive mechanism could explain these findings?
-One potential explanation is that incidental exposure to the premise statements leads participants to at least partially adopt them as beliefs.
-Simply entertaining the idea may create such an impression despite awareness of its ambiguous truth-value [@gilbert.etal1990; @gilbert1991; @garcia-marques.etal2015; @unkelbach.rom2017].
-If so, this apparently occurs despite participant's awareness that the statements may be false, as ensured by constant reminders and comprehension checks in Experiment 3.
-
-Modern theorists do not conceive of beliefs as lists of propositiosn, but instead as subserved by structured mental models or intuitive theories of how the world works [e.g.
-@gelman.legare2011; @gerstenberg.tenenbaum2017; @powell.etal2023; @johnson-laird2001; @thagard1989].
-Thus, coming to believe a statement entails updating one's mental models with the new information conveyed by that statement.
-A key feature of these mental models is their coherence: updating them is expected to have wide-ranging and coherent implications [@gerstenberg.tenenbaum2017; @powell.etal2023; @thagard1989].
-In the context of incidental exposure effects, Unkelbach and Rom [-@unkelbach.rom2017] have proposed an account based on parallel constraint satisfaction within semantic networks, which might plausibly be elaborated to account for the present findings.
-In any case, we take the coherence of inferences following exposure to be evidence for a meaningful influence on people's mental models, and the key to our claim that incidental exposure influences beliefs.
-
-Finally, it is worth clarifying that our findings and wider claims about the illusory implication effect are broadly consistent with existing metacognitive explanations of the illusory truth effect: we hypothesize that the illusory implication effect is a distinct and somewhat orthogonal effect, based in appreciation of the meaning of statements rather than solely in the metacognitive experiences they produce.
-Nevertheless, it remains clear that fluency plays a crucial role in the classic illusory truth effect, especially as evidenced by neuroscientific findings [@wang.etal2016] and by the findings indicating the effect can be reversed when the pragmatics of the exposure context are changed [@corneille.etal2020].
-
-## Conclusions
-
-Whatever the cognitive mechanism behind the illusory implication effect we observed, these findings are further cause for concern about the spread and influence of misinformation.
-The spread of misinformation and fake news appears to outpace the spread of factual news [@vosoughi.etal2018] and creates massive challenges for online platforms seeking to remove, label, and correct misinformation [@sharma.etal2019].
-These challenges may be further compounded by apparently basic features of human cognition: that incidental exposure to false statements can lead them to be _believed_.
-
-\newpage
-
-# References
-
-::: {#refs custom-style="Bibliography"}
-:::

### 2024-09-25 paper/fig-code.R

@@ -1,3 +1,4 @@
+library(tidybayes)
 make_exp_figure <- function(df_plotting, fit_model){
   
   
@@ -143,3 +144,182 @@ make_exp_figure <- function(df_plotting, fit_model){
   plt_raw + plt_binned + plot_layout(widths = c(3,1), guides = "collect")
   
 }
+
+
+make_exp_figure_two_models <- function(df_plotting, fit_model1, fit_model2){
+  
+  
+  obs_diffs <- df_plotting %>% 
+    filter(phase=="test") %>% 
+    filter(item_type == "B") %>% 
+    mutate(
+      exposed = if_else((exposed == TRUE | exposed==1), "Exposed", "Unexposed"),
+      B_implied_true = if_else(B_implied_true == "T implied F", "Contradicted", "Entailed") 
+    ) %>% 
+    group_by(exposed, B_implied_true, response) %>% 
+    summarize(
+      N = n()
+    ) %>%
+    group_by(B_implied_true, exposed) %>%
+    mutate(P = N/sum(N)) %>%
+    ungroup() %>% 
+    mutate(exposed = if_else((exposed == TRUE | exposed=="Exposed"), "Exposed", "Unexposed")) %>% 
+    
+    mutate(response = if_else(grepl("True", response), "True", "False")) %>%
+    group_by(B_implied_true, response, exposed) %>%
+    summarize(
+      P = sum(P),
+      SE = sqrt( P*(1-P) / sum(N))
+    ) %>% 
+    ungroup() %>%
+    pivot_wider(names_from = exposed, values_from = c(P, SE)) %>%
+    mutate(
+      diff = P_Exposed - P_Unexposed,
+      diff_se = sqrt(SE_Exposed^2 + SE_Unexposed^2)
+    ) %>%
+    rename(Implication = B_implied_true)
+  
+  
+  schema_preds  <-  df_reg %>% filter(exposure_cond=="quiz") %>%
+    filter(phase=="test") %>%
+    filter(item_type == "B") %>%
+    # filter(implication != 0) %>%
+    distinct(B_implied_true, familiarity, implication, .keep_all = TRUE) %>%
+    select(B_implied_true, familiarity, implication) %>% 
+    mutate(item = "new_item", subj_id = 9999) %>% 
+    add_epred_draws(fit_imp_quiz_int_bin, allow_new_levels = TRUE, re_formula = NA) %>% 
+    ungroup() %>% 
+    mutate(
[…]

### 2024-09-25 paper/updated-figs.R

@@ -1,0 +1,375 @@
+library(tidybayes)
+
+# fig1
+plot_df <- df %>% 
+  filter(phase=="test") %>% 
+  filter(item_type == "B") %>% 
+  group_by(exposed, exposure_cond, B_implied_true, response) %>% 
+  # summarize(
+  #   M = mean(resp_num),
+  #   se = sd(resp_num)/sqrt(n()),
+  #   ll = M - se,
+  #   ul = M + se
+  # ) %>% 
+  summarize(
+    N = n()
+  ) %>%
+  group_by(exposure_cond, B_implied_true, exposed) %>%
+  mutate(N = N/sum(N)) %>%
+  mutate(
+    exposure_cond = if_else(exposure_cond=="fact", "Fact condition", "Quiz condition"),
+    exposed = if_else(exposed==1, "Exposed", "Unexposed"),
+    B_implied_true = if_else( B_implied_true == "T implied F", "Contradicted", "Entailed") #as.character(B_implied_true)
+  ) %>% 
+  spread(exposed, N) %>% 
+  mutate(diff = Exposed - Unexposed)
+
+plot_df %>% 
+ggplot(
+    aes(x=B_implied_true,  y = response, fill = diff)
+  ) +
+  geom_tile() +
+  scale_fill_gradient2( low = '#b2182b', mid = '#f7f7f7', high = '#2166ac', midpoint = 0) +
+  facet_wrap(~exposure_cond) +
+  theme_bw(base_size=9.5) +
+  theme(panel.grid=element_blank(), legend.position="bottom") +
+  theme(axis.text.x = element_text(angle = 10, vjust = 0.5, hjust=.5))
+
+
+df %>% 
+  filter(phase=="test") %>% 
+  filter(item_type == "B") %>% 
+  group_by(exposed, exposure_cond, B_implied_true, response) %>% 
+  # summarize(
+  #   M = mean(resp_num),
+  #   se = sd(resp_num)/sqrt(n()),
+  #   ll = M - se,
+  #   ul = M + se
+  # ) %>% 
+  summarize(
+    N = n()
+  ) %>%
+  group_by(exposure_cond, B_implied_true, exposed) %>%
+  mutate(N = N/sum(N)) %>%
+  mutate(
+    exposure_cond = if_else(exposure_cond=="fact", "Fact condition", "Quiz condition"),
+    exposed = if_else(exposed==1, "Exposed", "Unexposed"),
[…]

### 2024-11-15 paper/paper-papaja-revising.Rmd
resubmission

@@
   - id            : "1"
+    institution   : "University of California, Irvine"
+  - id            : "2"
     institution   : "Arizona State University"
@@
 keywords          : "Illusory truth; Metacognition; Cognitive Psychology; Misinformation"
-wordcount         : "7531"
+wordcount         : "7658"
 
@@
 Similarly, Unkelbach and Rom [-@unkelbach.rom2017] examined both implied and contradictory statements.
-They found that exposure to implied statements increased truth ratings (Exp 4) and that statements saw a significant decrease in truth ratings following incidental exposure to their contradictions (Exp.
+Consistent with the hypothesized Illusory Implication effect, they found that exposure to statements increased truth ratings for novel implied statements (Exp.
+4) and that statements saw a significant decrease in truth ratings following incidental exposure to their contradictions (Exp.
 3).
-Crucially, the results of this experiment appear to preclude any purely fluency-based account of reduced truth ratings following incidental exposure to contradictory statements: These researchers found that responses were faster for contradictions of previously-seen statements than for novel statements, indicating that processing of these statements was *more* rather than less fluent.
+Crucially, the results of this experiment appear to preclude any purely fluency-based account of reduced truth ratings following exposure to contradictory statements: These researchers found that responses were faster for contradictions of previously-seen statements than for novel statements, indicating that processing of these statements was *more* rather than less fluent.
 Thus, even if an appropriate notion of "conceptual disfluency" could be developed, it is unlikely it could be an appropriate account of these findings [@unkelbach.rom2017].
@@
 
-However, these last two sets of findings must be qualified by the exposure context used in these studies: These researchers asked participants to make incidental judgments about these statements but did not warn them that the statements could be true or false [@silva.etal2017; @unkelbach.rom2017].
-Thus, the pragmatics of the task are unclear, and participants might have believed these statements because they understood them to be presented as facts.[^urnote]
-
-[^urnote]: Unkelbach and Rom [-@unkelbach.rom2017] did test a "warning" condition in another experiment of a similar design (Exp.
-1), but here did not find a significant reduction in truth judgments for contradictions of previously-seen statements.
+However, these last two sets of findings must be qualified by the exposure context used in these studies: These researchers exposed participants to statements by asking them to make innocuous judgments about these statements (e.g., "how interesting is this statement?") but did not warn them prior to this exposure that the statements could be true or false [@silva.etal2017; @unkelbach.rom2017].
+The pragmatics of this task are therefore ambiguous: participants might have imagined they were being asked to rate the interesting-ness of a set of entirely true statements, and might have then believed those statements for that reason.
+In the absence of strong and clear task pragmatics (e.g.
+a true or false quiz), explicit warnings are key to establishing incidental exposure as we have defined it here.
 
@@
 Existing research examining the effects of incidental exposure on implied statements has two key limitations.
-First, it has often been conducted in ambiguous contexts where participants might not understand that statements might be true or false, and that therefore do not always clearly induce incidental exposure.
-We address this through study designs inducing clear cases of incidental exposure.
-Second, prior work has only examined effects on direct contradictions of originally-seen statements (or paraphrases of direct contradictions).
+First, it has often been conducted in ambiguous contexts where participants might not understand that statements might be true or false, and that therefore do not always clearly induce incidental exposure, as we have defined it here.
+We address this in our study designs, by paying careful attention to the pragmatics of the exposure tasks to ensure they constitute clear cases of incidental exposure.
+Second, prior work has primarily examined effects on direct contradictions of originally-seen statements (or paraphrases of direct contradictions) [but see @unkelbach.rom2017].
 Here, we sought to further test for an "illusory implication effect" by testing the effect of incidental exposure to statements on their more distant logical implications.

### 2024-11-15 paper/pp_checks.R

@@ -1,0 +1,12 @@
+pp_check(fit_imp_quiz_int, type = 'bars_grouped', group = 'implication', ndraws = 500)
+pp_check(fit_imp_quiz_int, type = 'bars_grouped', group = 'familiarity', ndraws = 500)
+pp_check(fit_imp_quiz_int, type = 'bars_grouped', group = 'B_implied_true', ndraws = 500)
+
+
+pp_check(fit_imp_exp2, type = 'bars_grouped', group = 'implication', ndraws = 500)
+pp_check(fit_imp_exp2, type = 'bars_grouped', group = 'familiarity', ndraws = 500)
+pp_check(fit_imp_exp2, type = 'bars_grouped', group = 'B_implied_true', ndraws = 500)
+
+pp_check(fit_imp_maximal3, type = 'bars_grouped', group = 'implication', ndraws = 500)
+pp_check(fit_imp_maximal3, type = 'bars_grouped', group = 'familiarity', ndraws = 500)
+pp_check(fit_imp_maximal3, type = 'bars_grouped', group = 'B_implied_true', ndraws = 500)
